### 2018-02-23 Modeling.Rmd
First commit

---
title: "Modeling"
author: "Michael Bostwick"
date: "2/7/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Modeling

```{r , echo=FALSE, include=FALSE}
setwd("/Users/michaelbostwick/Documents/Galapagos")

library(tidyverse)
library(ggplot2)
library(readxl)
library(glmnet)
library(mpath)
library(reshape2)
library(gridExtra)
library(leaps)

load("BASES_CENSO_UPA_spss/clean_data.RData")
vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")

```

## Variable pre-selection

Check for highly correlated variables. Variables that have perfect or very high correlation will have one of the pair removed (marked as a "C" in the Variables spreadsheet).

```{r, echo=FALSE, warning=FALSE}
vars_include <- vars[which(is.na(vars$`Surface area with invasive species`)),1]$`Variable Name`
x_include <- subset(reduced_data, select =vars_include)

numeric_data <- model.matrix(~., x_include)[,-1]
corrs <-cor(numeric_data)
corrs[upper.tri(corrs)] <- NA
diag(corrs) <- NA

corrs_reshape <- melt(corrs, na.rm = TRUE)
(high_corrs <- corrs_reshape[abs(corrs_reshape$value) > 0.9,])

```

## UPA Production

```{r, echo=FALSE}
zero_prod <- reduced_data[reduced_data$productivity ==0,]
summary(zero_prod$netincome)
par(mfrow = c(2,1))
raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity))
log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity)))
lower_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) +
  xlim(0, 10000) + xlab("productivity (0-10,000)")
grid.arrange(raw_prod_hist, lower_prod_hist, log_prod_hist, nrow = 1, top = "Histogram of Productivity Response Variable")

```

```{r, echo=FALSE}

production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
production_x <- model.matrix(~., production_x.df)[,-1]

log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])

set.seed(1)
foldid=sample(1:10,size=length(log_productivity),replace=TRUE)
cv1=cv.glmnet(production_x,log_productivity,foldid=foldid,alpha=1)
cv.75=cv.glmnet(production_x,log_productivity,foldid=foldid,alpha=0.75)
cv.5=cv.glmnet(production_x,log_productivity,foldid=foldid,alpha=.5)
cv.25=cv.glmnet(production_x,log_productivity,foldid=foldid,alpha=0.25)
cv0=cv.glmnet(production_x,log_productivity,foldid=foldid,alpha=0)

cv1.mse <- cv1$cvm[cv1$lambda == cv1$lambda.1se]
cv.75.mse <- cv.75$cvm[cv.75$lambda == cv.75$lambda.1se]
cv.5.mse <- cv.5$cvm[cv.5$lambda == cv.5$lambda.1se]
cv.25.mse <- cv.25$cvm[cv.25$lambda == cv.25$lambda.1se]
cv0.mse <- cv0$cvm[cv0$lambda == cv0$lambda.1se]

par(mfrow=c(2,2))
plot(cv1,main="Alpha = 1");plot(cv.75,main="Alpha = 0.75");plot(cv.5,main="Alpha = 0.5");plot(cv0,main="Alpha = 0");
plot(log(cv1$lambda),cv1$cvm,pch=19,col="red",xlab="log(Lambda)",ylab=cv1$name,  main="Alpha = 1")
points(log(cv.5$lambda),cv.5$cvm,pch=19,col="grey", main="Alpha = 0.5")
points(log(cv0$lambda),cv0$cvm,pch=19,col="blue", main="Alpha = 0")
par(mfrow=c(1,1))

par(mfrow=c(2,1))
plot(cv1$glmnet.fit, s = cv1$lambda.1se, xvar = "dev", label = FALSE)
plot(cv.5$glmnet.fit, s = cv.5$lambda.1se, xvar = "dev", label = FALSE)

lasso.coef <- predict(cv1,type="coefficients",s=cv1$lambda.1se)[1:ncol(production_x),] 
lasso.coef[lasso.coef!=0]

cv.5.predicts <- predict(cv.5, newx = production_x,  type = "response", s = cv.5$lambda.1se)
cv.5.resids <- (log_productivity - cv.5.predicts)
plot(cv.5.predicts, cv.5.resids)

```

Alpha [0.25 - 1] all seem to exhibit similar predictive performance, alpha = 0.5 is chosen. Alpha = 0 has slightly higher error.

```{r, echo=FALSE}
set.seed(1)
bs_coefs <- data.frame(matrix(ncol = ncol(production_x)+1, nrow = 0))
colnames(bs_coefs) <- c("(Intercept)", colnames(production_x))

for (i in 1:1000){
  bootstrap_samples <- sample(1:nrow(production_x), size = nrow(production_x), replace = TRUE)
  weights <- table(factor(bootstrap_samples, levels = 1:nrow(production_x)))
  bs.elasticnet = glmnet(production_x,log_productivity,weights = weights,alpha=.5, lambda =  cv.5$lambda.1se)

  bs.coef <- t(data.frame(predict(bs.elasticnet,type="coefficients",s=bs.elasticnet$lambda.1se)[1:(ncol(production_x)+1),]))
  bs_coefs <- rbind(bs_coefs, bs.coef)
}

row.names(bs_coefs) <- c(seq(1:1000))

boxplot(bs_coefs[, abs(colSums(bs_coefs)) > 0.1])

pct_zeroes <- data.frame(colSums(bs_coefs != 0)/nrow(bs_coefs))
pct_zeroes <- rownames_to_column(pct_zeroes)
colnames(pct_zeroes) <- c("variable", "pct_zero")

ggplot(pct_zeroes[pct_zeroes$pct_zero < .01,]) + geom_bar(stat='identity', aes(x=variable, y=pct_zero)) + coord_flip()

ggplot(data = melt(bs_coefs[,51:100]), aes(x=variable, y=value)) + geom_boxplot() + coord_flip()
```

Limit boxplot to coefficients that are not zero

```{r, echo=FALSE}
best.fwd <- regsubsets(log_productivity~., data=production_x.df ,nvmax=200,method = "forward")
best.fwd.summary <- summary(best.fwd)

best.bwd <- regsubsets(log_productivity~., data=production_x.df ,nvmax=200,method = "backward")
best.bwd.summary <- summary(best.bwd)

plot_results <- function(select.summary1, select.summary2){
  par(mfrow=c(2,2))
  plot(select.summary1$cp ,xlab="Number of Variables ",ylab="Cp", type='l')
  min.cp <- which.min(select.summary1$cp)
  points(min.cp,select.summary1$cp[min.cp],col="red",cex=2,pch=20)

  plot(select.summary1$bic ,xlab="Number of Variables ",ylab="BIC", type='l')
  min.bic <- which.min(select.summary1$bic)
  points(min.bic,select.summary1$bic[min.bic],col="red",cex=2,pch=20)

  plot(select.summary2$cp ,xlab="Number of Variables ",ylab="Cp", type='l')
  min.cp <- which.min(select.summary2$cp)
  points(min.cp,select.summary2$cp[min.cp],col="blue",cex=2,pch=20)

  plot(select.summary2$bic ,xlab="Number of Variables ",ylab="BIC", type='l')
  min.bic <- which.min(select.summary2$bic)
  points(min.bic,select.summary2$bic[min.bic],col="blue",cex=2,pch=20)

  par(mfrow=c(1,1))
}

plot_results(best.fwd.summary, best.bwd.summary)

fwd.coefs <- coef(best.fwd, which.min(best.fwd.summary$bic))
incl_vars <- names(fwd.coefs)[-1]

incl_x <- production_x[,incl_vars]

fwd.predicts <- incl_x %*% fwd.coefs[-1] + fwd.coefs[1]
fwd.resids <- (log_productivity - fwd.predicts)

plot(fwd.predicts, fwd.resids)
```

### 2018-02-23 Paper/paper_draft.tex
First commit

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2016
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2016}

\usepackage[final, nonatbib]{nips_2016}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts, amssymb, amsmath}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}
\usepackage{algpseudocode}
\usepackage{algorithmicx}
\usepackage{csvsimple}

\newcommand{\bX}{\mathbf{X}}
\newcommand{\bY}{\mathbf{Y}}
\newcommand{\bbeta}{\mathbf{\beta}}

\graphicspath{ {images/} }

\title{Survival of the Fittest: Variable Selection on Galapagos Agricultural Data}

\author{
  Michael Bostwick\\
  Department of Statistics and Operations Research\\
  University of North Carolina at Chapel Hill\\
  \texttt{mgb2188@live.unc.edu} \\
}

\begin{filecontents*}{scientists.csv}
name,surname,age
Albert,Einstein,133
Marie,Curie,145
Thomas,Edison,165
\end{filecontents*}

\begin{document}

\maketitle
\begin{abstract}
  This paper...

  \end{abstract}

\section{Introduction}

\subsection{Background}

The Galapagos Islands make for a feasible and significant case study of complex systems. Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Galapagos Islands than other systems. Yet the Galapagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment. Prior work has created agent-based models of the Galapagos, but with limited interaction parameters between agents, particularly in regards to farm success walsh. In order to create a more detailed, and perhaps more accurate, simulation, the relationships between different factors on the island must be better understood. This work aims to sort through a large number of possible relationships and clarify the empirically most significant ones for future study and incorporation into simulations.

\subsection{Data}

The data available to study the dynamics between agricultural measures and related factors primarily 
comes from the Census of Agricultural Production Units (UPA) of Galapagos (Censo de las Unidades 
de Producci\'on Agropecuaria (upa) de Gal\'apagos). This is a self-reported survey with data from 755 
farms detailing the production and sale of crops and livestock, agricultural expenses, and land use 
decisions.

In addition to the census, data is also available from ... including information on water, energy and road access.

In total, under the direction of the client 240 variables were selected for consideration in modeling 
relationships between predictors and the outcome variables of interest.

\section{Modeling}
\subsection{Challenges to address}

The primary challenge in this analysis is the vast amount of potential predictor variables. This challenge is twofold; a.) when 
the number of predictors is large the calculation of a reliable model is difficult and b.) interpreting the coefficients of many 
predictors simultaneously is not an easy task for humans (and will make resulting simulations overly complicated). For this 
reason, the analysis focuses on the use of several variable selection techniques and a comparison of their performance. 
These variable selection techniques aim to build a linear model with a subset of the available variables that still maintains 
a strong predictive performance.

Secondarily, when performing standard linear regression the outcome variable is assumed to be roughly normally 
distributed. When this is not true, as is the case for several outcome variables in this study, a poorly fitting model will be 
found with unreliable coefficients. In order to address this issue transformations to the data and modifications to the 
standard linear model will be considered.

\subsection{Overview of methods}

A brief overview of the statistical methods used is presented here to allow for understanding of results, 
but for further details see Section refstatmethods Statistical Methods. For each of the outcome variables of interest 
we build a set of linear models using the appropriate subset of predictors. Each relationship is modeled 
using Best Subset, Forward Selection and ElasticNet regression. Best Subset considers all variable combinations to find the 
linear model of best fit. Forward Selection fits a linear model by progressively adding variables to the model until a best fit is 
found. This results in only some of the variables being included, chosen in a discrete manner. ElasticNet regression fits a linear 
model by constraining the coefficients so that they are smaller than in standard least squares, and for many variables actually 
shrunken to zero. Similar to Forward Selection this results in a sparser model, but variable selection can be carefully tuned 
as optimization is done in a more continuous way.

\section{Results}

\subsection{Farm Success}

We define farm success with three different measurements: productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock produced divided by the farm surface area. Net income is calculated as the difference between the earnings from all products sold and total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard worker's salary.

\subsubsection{Productivity}

Response variable plots and discussion
\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{production_histograms.pdf}
\caption{Histogram of Productivity...}
\end{figure}

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} & \bfseries \underline{Best~Subset}
\csvreader[head to column names]{production_top10names.csv}{}%
{\\\elasticnet & \forward & \subset}%
\\\hline
\end{tabular}
\caption{Modeling of Productivity, Top 5 features for each of the three methods}
\label{table:1}
\end{table}

\begin{figure}
\includegraphics[width = \textwidth]{elastic_cv_production.pdf}
\caption{Cross Validation plots for Elasticnet...}
\end{figure}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{forward_nvars_production.pdf}
\caption{Number of variables Forward Selection plots...}
\end{figure}

Bootstrap plots

\subsubsection{Net Income}

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{netincome_histograms.pdf}
\caption{Histogram of Net Income...}
\end{figure}

\begin{table}[h!]
\centering
\begin{tabular}{c|c|c}\hline%
\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} & \bfseries \underline{Best~Subset}
\csvreader[head to column names]{netincome_top10names.csv}{}%
{\\\elasticnet & \forward & \subset}%
\\\hline
\end{tabular}
\caption{Modeling of Net Income, Top 5 features for each of the three methods}
\label{table:1}
\end{table}

\begin{figure}
\includegraphics[width = \textwidth]{elastic_cv_netincome.pdf}
\caption{Cross Validation plots for Elasticnet...}
\end{figure}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{forward_nvars_netincome.pdf}
\caption{Number of variables Forward Selection plots...}
\end{figure}

\subsubsection{Number of Workers Supported}

\subsection{Invasive Species}

\subsection{Land use choices}

\section{Statistical Methods} \label{stat_methods}

\subsection{Generalized Linear Models}

\[\bY = \bX\beta + \epsilon \tag{1}\]
%\[Y_i =  \sum_{j=1}^{p} X_{i,j}\beta_j + \epsilon \tag{1}\]

Standard Linear Regression can be represented in matrix form as equation 1 above, where $\bY$ is a $n \times 1$ vector of 
the outcome variable, $\bX$ is a $n \times p$ matrix of the predictor variables, $\bbeta$ is a $p \times 1$ vector of variable coefficients and $\epsilon$ is the error 
term. The standard linear model assumes the outcome variable $\bY$ has a normal distribution, and therefore takes 
continuous values. When the outcome variable is continuous, but not normal shaped (perhaps skewed like the productivity 
data) it can be possible to transform the data by taking the logarithm or something similar. However, when the outcome 
variable is discrete (such as labels of 1 and 0 denoting absence/presence of agricultural land use) a further modification must be made. The outcome variable is clearly no longer normally distributed, as it is not even continuous. Without modification we could get predicted values below 0, above 1 or somewhere in between, none of which make much sense.

This calls for the use of logistic regression, in which we perform a logit transformation as seen in the equation below so that the 
$\bX\beta$ can still be mapped to a continuous scale. In some respects this is a computational concern, but it also changes the 
way coefficients can be interpreted. Instead of a one unit change in $X_1$ predicting a $\beta_1$ change in the predicted $Y$, 
in this case it predicts a $\beta_1$ change in the log odds of $Y$.

\[log \frac{\Pr(Y = 1)}{\Pr(Y = 0)} = \bX\beta \]

This equation can be rewritten as below

\[ \Pr(Y = 1) = \frac{e^{\bX\beta}}{1 + e^{\bX\beta}} \]

Multiclass Logistic Regression

\[ \Pr(Y = k) = \frac{e^{\bX\beta_k}}{\sum_{l=1}^{K-1} e^{\bX\beta_l}} \]

% include Poisson function

\subsection{Performance Measures}

There are many measures of the fit of linear models. Since there are many possible predictor variables, care must be taken 
to use appropriate measures, as some measures will favor just adding all of the variables to the model. For example, if we 
aim to minimize the mean square error adding more predictor variables to the model will always be encouraged. Since that is not desired, measurements like Bayesian Information Criterion (BIC) and Mallow's $C_p$ ($C_p$) can used here. BIC and $C_p$ are a combination of how well the model fits the data and a penalty term for the number of predictor variables included in the model. The goal is to minimize BIC and $C_p$, that is the model with the best balance of small size and goodness of fit. BIC puts a large penalty on the inclusion of additional variables, so it will be our primary metric.

Another approach is to use cross-validation. In this technique the dataset is first split into $n$ equally sized sets. Then a 
model is fit using $n - 1$ of the sets (training sets) and evaluated on the remaining 1 (test) set. This is repeated $n$ times, 
each time reserving a different 1 test set, and then results across the $n$ runs are averaged. The benefit of this is that model 
building and model evaluating are happening on different portions of the data, so we can distinguish if the model is picking up 
on generalizable patterns or just random noise. Using cross-validation the average test set mean square error is an 
appropriate measure of model fit.

\subsection{Best Subset and Forward Selection}

The essential goal of variable selection is to find the best combination of predictor variables to explain the outcome variable. As 
discussed above, when we have many possible predictors we often want to put a constraint on the problem so that all variables 
are not included. Such a constraint might be limiting the number of variables included or that the model found can generalize to 
other data. Best subset selection, the most natural, but computationally difficult way is to try all possible combinations of variables and select the best fitting combination. However, when the number of variables, $p$, is large this quickly becomes infeasible, as there are $2^p$ possible combinations. Recent advances have expanded the problem sizes that can be efficiently computed using Mixed Integer Optimization. Using this technique we find the best 5 variable models for each of the outcomes, but are not able to search all possible model sizes to find the optimal number of variables. In addition, as marked in the results above, sometimes the algorithm can verify that it has found the optimal solution, while other times it returns the best solution found in the allotted time limit (set to 60 seconds), that may or may not be the overall optimal solution.

One approach to tackle the computational complexity discussed above is to restrict the search for the optimal number of 
predictor variables, which is what Forward Selection does. In this algorithm, we start with an empty model and iteratively add a 
new variable at each stage that is most beneficial. This procedure can work well, but may not find the optimal solution. As an example, consider a case where $X_1$ is the single most predictive variable, but the combination of $X_2$ and $X_3$ is the best two variable combination. The algorithm will first add $X_1$, but then regardless whether it adds $X_2$ or $X_3$ next, it will have found a suboptimal solution. In general, we can decide to stop adding variables once we have reached an optimal performance measure like BIC or cross-validation test error.

\subsection{Regularized Regression}

% Consider taking these out of matrix form

	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} \tag{linear model}\]
	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda||\beta||_2^2 \tag{ridge regression}\]
	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda||\beta|| \tag{LASSO}\]

As shown in equation ? in the standard linear model, we try to find the $\beta$ that minimizes the squared difference between 
the true $\bY$ and the predicted $\hat{\bY}$ (which is $\beta\bX$). In regularized regression we do the same thing, but also 
add a second term that we look to simultaneously minimize. This second term adds a penalty for increasing values of $
\beta$, so the two terms must be balanced. The optimal model will find a balance between fitting the outcome variable 
closely, but not having too large of coefficient values. The difference between Ridge regression and LASSO is how we add up 
the coefficients. In Ridge Regression the coefficients are squared and then summed, in LASSO we take the absolute value of 
the coefficients and then sum them. LASSO will encourage most of the coefficients to go to zero, thus only including a small 
number of terms in the model. Ridge regression will encourage the coefficient values to be spread out among predictor 
variables, leaving all of the variables in the model, but helping to offset negative effects of correlated predictor variables. Further explanations of this can be found in ???

\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[(1 - \alpha)||\beta||_2^2 + \alpha||\beta||] \tag{ElasticNet}\]

The technique that is used in this analysis is a combination of the Ridge and LASSO penalties, called ElasticNet. As can be seen in the equation above both the square of the coefficients and the absolute value of the coefficients is included, with the contribution of each controlled by the size of $\alpha$ which takes a value between 0 and 1. ElasticNet, thus combines the favorable properties of Ridge and LASSO in that it can achieve both sparse models and deal with correlated predictor variables. Both the $\lambda$ and the $\alpha$ can be set using cross-validation (as discussed above) to be appropriate values for the particular dataset.

\subsection{Comparison of methods}

\section{Conclusion}

\medskip

\small

\begin{thebibliography}{99}

\end{thebibliography}

\section{Appendix}

\begin{table}[h!]
\begin{tabular}{l|c|c}\hline%
\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{fullcoeflist_production.csv}{}%
{\\\variable & \elastic & \forward}%
\\\hline
\end{tabular}
\caption{Full coefficient list for Production model}
\end{table}

\begin{table}[h!]
\begin{tabular}{l|c|c}\hline%
\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{fullcoeflist_netincome.csv}{}%
{\\\variable & \elastic & \forward}%
\\\hline
\end{tabular}
\caption{Full coefficient list for Net Income model}
\end{table}

\begin{figure}[h]
\includegraphics[width = \textwidth]{resids_netincome.pdf}
\caption{Diagnostic Residual plots for Net Income...}
\end{figure}

\end{document}

### 2018-02-23 CensusExploration2.R

@@ -1,0 +1,508 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos/BASES_CENSO_UPA_spss")
+
+#Still to do: 
+  #Add categorical data from spreadsheets (such as crops)
+  #Generate physical data from DEM (This will take a while still)
+
+library(tidyverse)
+
+#Load Data (currently as SPSS files and CSV files, trying out different ones. Cleaned CSV files Use those. Omitting some )
+  library(foreign)
+  #Censo CGREG-MAG 2015
+  CGenerales <- read.csv("Cap1_CaracteristicasGenerales.csv", header = TRUE)
+  Superficie <- read.csv("Cap2_Superficie.csv", header = TRUE)
+  Riego <- read.csv("Cap3_Riego.csv", header = TRUE)
+  UsoSuelo <- read.csv("Cap3_2_UsoSuelo_Hoy.csv", header = TRUE)
+  CPermanentes <- read.csv("Cap4_CultivosPermanentes.csv", header = TRUE)
+  CTransitorios <- read.csv("Cap5_Cultivos_Transitorios.csv", header = TRUE)
+  Arboles <- read.csv("Cap6_ArbolesDispersos.csv", header = TRUE)
+  Pastos <- read.csv("Cap7_PastosCultivados.csv", header = TRUE)
+  Forestal <- read.csv("Cap8_ActividadForestal.csv", header = TRUE)
+  Vacuno <- read.csv("Cap9_GanadoVacuno.csv", header = TRUE)
+  #Porcino <- read.spss("Cap10_Ganado Porcino.sav", to.data.frame=TRUE)
+  #OtroGanado <- read.spss("Cap11_Otras Especies de Ganado.sav", to.data.frame=TRUE)
+  #Aves <- read.spss("Cap12_Aves.sav", to.data.frame=TRUE)
+  Gastos <- read.csv("Cap13_Gastos_Inversiones_Financiamiento.csv", header = TRUE)
+  Equipo <- read.csv("Cap14_Equipo_maquinaria_instalaciones.csv", header = TRUE)
+  Trabajadores <- read.csv("Cap15_Trabajadores.csv", header = TRUE)
+  AfiliacionDeshechos <- read.csv("Cap16_AsistenciaTecnica_Afiliacion_ManejoDesechos.csv", header = TRUE)
+  
+  #INEC 2015
[…]

### 2018-02-23 bootstrap_code.R

@@ -1,0 +1,24 @@
+set.seed(1)
+bootstrap_coefs <- data.frame(matrix(ncol = ncol(x.matrix)+1, nrow = 0))
+colnames(bootstrap_coefs) <- c("(Intercept)", colnames(x.matrix))
+
+for (i in 1:1000){
+  bootstrap_samples <- sample(1:nrow(x.matrix), size = nrow(x.matrix), replace = TRUE)
+  weights <- table(factor(bootstrap_samples, levels = 1:nrow(x.matrix)))
+  bs.elasticnet = glmnet(x.matrix,y.response,weights = weights,alpha=best_alpha, lambda = best.lambda)
+  
+  bs.coef <- t(data.frame(predict(bs.elasticnet,type="coefficients",s=best.lambda)[1:(ncol(x.matrix)+1),]))
+  bootstrap_coefs <- rbind(bootstrap_coefs, bs.coef)
+}
+
+row.names(bootstrap_coefs) <- c(seq(1:1000))
+top.names <- unique(c(elastic10.names, fwd.coefs10))
+top.boostrap <- bootstrap_coefs[,top.names]
+
+ggplot(data = melt(top.boostrap), aes(x=variable, y=value)) + geom_boxplot() + coord_flip()
+
+pct_zeroes <- data.frame(colSums(top.boostrap != 0)/nrow(top.boostrap))
+pct_zeroes <- rownames_to_column(pct_zeroes)
+colnames(pct_zeroes) <- c("variable", "pct_zero")
+
+ggplot(pct_zeroes) + geom_bar(stat='identity', aes(x=variable, y=pct_zero)) + coord_flip()

### 2018-02-23 data_cleaning.R

@@ -1,0 +1,41 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+
+load("BASES_CENSO_UPA_spss/prelim_data.RData")
+
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+var_list <- vars$`Variable Name`
+for (i in 1:nrow(vars)){
+  result <- tryCatch(reduced_data <-  data[,var_list[i]],
+                     error = function(e) {print(paste(i,var_list[i]))})
+}
+reduced_data <-  data[,var_list]
+
+binary_vars <- vars[which(vars$Binary==1),1]$`Variable Name`
+reduced_data[,binary_vars][is.na(reduced_data[,binary_vars])] <- 0
+
+colnames(reduced_data)[colSums(is.na(reduced_data)) > 0]
+
+fill_na_zero <- c("r3","p7","p8","p9","p26","CosechaLibras","VentaLibras","num_cultivo","t28","libras_cosechada",
+                  "libras_vendida","t8","t10","pc6","ad6","ad4","ad11","produccion_en_libras_producto_cosechado_autoconsumo",
+                  "produccion_en_libras_producto_vendido","v30_a","v44","v49","v53_a","v53_b","v54","af6","e29",
+                  "e30_f","to56_e","to57_e","to57_a","to57_b","to57_c","to57_d","to57_f")
+
+reduced_data[,fill_na_zero][is.na(reduced_data[,fill_na_zero])] <- 0
+
+# Add None level for Farms without Pasture
+levels(reduced_data$pc4) <- c(levels(reduced_data$pc4), "None")
+reduced_data[,"pc4"][is.na(reduced_data[,"pc4"])] <- "None"
+
+# Convert mistaken factor variables to numeric variables
+reduced_data$e29_f <- as.numeric(reduced_data$e29_f)
+reduced_data$v45 <- as.numeric(reduced_data$v45)
+
+# Check to make sure there are no more mistaken factor variables
+reduced_data[1,sapply(reduced_data, is.factor)]
+
+
+save(reduced_data, file='BASES_CENSO_UPA_spss/clean_data.RData')

### 2018-02-23 data_creation.R

@@ -1,0 +1,392 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos/BASES_CENSO_UPA_spss")
+
+#Still to do: 
+  #Add categorical data from spreadsheets (such as crops)
+  #Generate physical data from DEM (This will take a while still)
+
+library(tidyverse)
+
+#Load Data (currently as SPSS files and CSV files, trying out different ones. Cleaned CSV files Use those. Omitting some )
+  library(foreign)
+  #Censo CGREG-MAG 2015
+  CGenerales <- read.csv("Cap1_CaracteristicasGenerales.csv", header = TRUE)
+  Superficie <- read.csv("Cap2_Superficie.csv", header = TRUE)
+  Riego <- read.csv("Cap3_Riego.csv", header = TRUE)
+  UsoSuelo <- read.csv("Cap3_2_UsoSuelo_Hoy.csv", header = TRUE)
+  CPermanentes <- read.csv("Cap4_CultivosPermanentes.csv", header = TRUE)
+  CTransitorios <- read.csv("Cap5_Cultivos_Transitorios.csv", header = TRUE)
+  Arboles <- read.csv("Cap6_ArbolesDispersos.csv", header = TRUE)
+  Pastos <- read.csv("Cap7_PastosCultivados.csv", header = TRUE)
+  Forestal <- read.csv("Cap8_ActividadForestal.csv", header = TRUE)
+  Vacuno <- read.csv("Cap9_GanadoVacuno.csv", header = TRUE)
+  Porcino <- read.csv("Cap10_GanadoPorcino.csv",header = TRUE)
+  OtroGanado <- read.csv("Cap11_OtrasEspeciesGanado.csv", header = TRUE)
+  Aves <- read.csv("Cap12_Aves.csv", header = TRUE)
+  Gastos <- read.csv("Cap13_Gastos_Inversiones_Financiamiento.csv", header = TRUE)
+  Equipo <- read.csv("Cap14_Equipo_maquinaria_instalaciones.csv", header = TRUE)
+  Trabajadores <- read.csv("Cap15_Trabajadores.csv", header = TRUE)
+  AfiliacionDeshechos <- read.csv("Cap16_AsistenciaTecnica_Afiliacion_ManejoDesechos.csv", header = TRUE)
+  
+  #INEC 2015
+  #Poblacion2015 <- read.spss("Poblacion_CPVG15_AT.sav", to.data.frame=TRUE)
[…]

### 2018-02-23 invasive_model.R

@@ -1,0 +1,36 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+library(reshape2)
+library(gridExtra)
+library(leaps)
+library(bestsubset)
+
+source("modeling.function.R")
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+
+invasive_include <- vars[which(is.na(vars$`Surface area with invasive species`)),1]$`Variable Name`
+
+
+invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = percinv))
+nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
+
+pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
+grid.arrange(all_workers_hist, nonzero_hist, nrow = 1)
+dev.off()
+
+
+workers_nonzero <- reduced_data[reduced_data$fulltimework > 0, 'fulltimework']
+reduced_data$workers_binary <- (reduced_data$fulltimework > 0)*1
+
+workers_x.df <- subset(reduced_data[,-workers_binary], select = workers_include)
+
+
+workers.models <- fit.models(model.name = "workers", x.data = workers_x.df, 
+                                y.response = reduced_data$workers_binary, response.family = "binomial")

### 2018-02-23 modeling.R

@@ -1,0 +1,61 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I242")
+
+production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
+production_x <- subset(reduced_data, select = production_include)
+#production_x_num <- as.matrix(production_x[sapply(production_x, function(x) !is.factor(x))])
+production_x <- model.matrix(~., production_x)[,-1]
+
+fit <- lm(productivity ~ . , data = subset(reduced_data, select = production_include))
+
+fit = glmnet(production_x, reduced_data$productivity)
+plot(fit)  
+
+reduced_data$log_productivity <- log(reduced_data$productivity+1)
+
+foldid=sample(1:10,size=length(reduced_data$log_productivity),replace=TRUE)
+cv1=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=1)
+cv.5=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=.5)
+cv0=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=0)
+
+
+par(mfrow=c(2,2))
+plot(cv1,main="Alpha = 1");plot(cv.5,main="Alpha = 0.5");plot(cv0,main="Alpha = 0");plot()
+#plot(log(cv1$lambda),cv1$cvm,pch=19,col="red",xlab="log(Lambda)",ylab=cv1$name,  main="Alpha = 1")
+#points(log(cv.5$lambda),cv.5$cvm,pch=19,col="grey", main="Alpha = 0.5")
+#points(log(cv0$lambda),cv0$cvm,pch=19,col="blue", main="Alpha = 0")
+#legend("bottomleft",legend=c("alpha= 1","alpha= .5","alpha 0"),pch=19,col=c("red","grey","blue"))
+par(mfrow=c(1,1))
+
+lasso.coef=predict(cv1,type="coefficients",s=cv1$lambda.1se)[1:240,]
+lasso.coef[lasso.coef!=0]
+
+#model fulltime workers as poisson? overdispersion? negative binomial?
+hist(reduced_data$fulltimework)
[…]

### 2018-02-23 modeling.function.R

@@ -1,0 +1,182 @@
+fit.models <- function(model.name, x.data, y.response, response.family){
+  
+  # For testing
+  x.data <- workers_x.df
+  y.response <- reduced_data$workers_binary
+  response.family <- "binomial"
+  
+  x.matrix <- model.matrix(~., x.data)[,-1]
+  
+  # Elastic net modeling
+  set.seed(1)
+  foldid=sample(1:10,size=length(y.response),replace=TRUE)
+  if (response.family == "gaussian"){
+    cv1=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=1)
+    cv.75=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=0.75)
+    cv.5=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=.5)
+    cv.25=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=0.25)
+  } else {
+    cv1=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=1)
+    cv.75=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.75)
+    cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
+    cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
+  }
+  models <- vector(mode="list", length=4)
+  
+  models[[1]] <- cv.25
+  models[[2]] <- cv.5 
+  models[[3]] <- cv.75 
+  models[[4]] <- cv1
+
+  alphas <- c(0.25,0.5,0.75,1)
+
+  cv1.mse <- cv1$cvm[cv1$lambda == cv1$lambda.1se]
+  cv.75.mse <- cv.75$cvm[cv.75$lambda == cv.75$lambda.1se]
+  cv.5.mse <- cv.5$cvm[cv.5$lambda == cv.5$lambda.1se]
+  cv.25.mse <- cv.25$cvm[cv.25$lambda == cv.25$lambda.1se]
+  
+  mses <- c(cv.25.mse, cv.5.mse, cv.75.mse, cv1.mse)
+  best_model <- models[[which.min(mses)]]
+  best_alpha <- alphas[which.min(mses)]
+  
[…]

### 2018-02-23 netincome_model.R

@@ -1,0 +1,34 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+library(reshape2)
+library(gridExtra)
+library(leaps)
+library(bestsubset)
+
+source("modeling.function.R")
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+
+netincome_include <- vars[which(is.na(vars$`Net Income`)),1]$`Variable Name`
+
+
+
+full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome))
+reduced_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + 
+  scale_x_continuous(limits = c(-5e+04, 5e+04))
+pdf("Paper/images/netincome_histograms.pdf",width=12,height=8)
+grid.arrange(full_hist, reduced_hist, nrow = 1)
+dev.off()
+
+
+netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < 5e+04,], select = netincome_include)
+netincome <- reduced_data[abs(reduced_data$netincome) < 5e+04, 'netincome']
+
+netincome.models <- fit.models(model.name = "netincome", x.data = netincome_x.df, 
+                                y.response = netincome, response.family = "gaussian")

### 2018-02-23 production_model.R

@@ -1,0 +1,35 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+library(reshape2)
+library(gridExtra)
+library(leaps)
+library(bestsubset)
+
+source("modeling.function.R")
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+
+production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
+production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
+
+log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])
+
+zero_prod <- reduced_data[reduced_data$productivity ==0,]
+par(mfrow = c(2,1))
+raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity))
+log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity)))
+lower_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) +
+  xlim(0, 10000) + xlab("productivity (0-10,000)")
+
+pdf("Paper/images/production_histograms.pdf",width=12,height=8)
+grid.arrange(raw_prod_hist, lower_prod_hist, log_prod_hist, nrow = 1)
+dev.off()
+
+production.models <- fit.models(model.name = "production", x.data = production_x.df, 
+                                y.response = log_productivity, response.family = "gaussian")

### 2018-02-23 workers_model.R

@@ -1,0 +1,34 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+library(reshape2)
+library(gridExtra)
+library(leaps)
+library(bestsubset)
+library(bestglm)
+
+source("modeling.function.R")
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+
+workers_include <- vars[which(is.na(vars$`Number workers supported`)),1]$`Variable Name`
+
+workers_nonzero <- reduced_data[reduced_data$fulltimework > 0, 'fulltimework']
+reduced_data$workers_binary <- (reduced_data$fulltimework > 0)*1
+
+workers_x.df <- subset(reduced_data, select = workers_include)
+
+all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = factor(workers_binary)))
+nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
+
+pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
+grid.arrange(all_workers_hist, nonzero_hist, nrow = 1)
+dev.off()
+
+workers.models <- fit.models(model.name = "workers", x.data = workers_x.df, 
+                                y.response = reduced_data$workers_binary, response.family = "binomial")

### 2018-02-27 data_creation.R
More modeling

@@ -352,7 +352,7 @@ library(tidyverse)
   summary(data$permworkerspay)
   
 #NUMBER OF FULL TIME WORKERS SUPPORTED
-  data$fulltimework<-as.integer(data$laborcost/659)
+  data$fulltimework<-data$laborcost/659
 
 #PRODUCTION COST
   data$productioncost <-rowSums(data[,c('GastosAgricolas','GastosPecuarios')], na.rm=TRUE) #Expenses for all agricultural activities and labor

### 2018-02-27 invasive_model.R

@@ -16,21 +16,45 @@ load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
 
 invasive_include <- vars[which(is.na(vars$`Surface area with invasive species`)),1]$`Variable Name`
+invasive_x.df <- subset(reduced_data, select = invasive_include)
 
+reduced_data$invasive <- 1*(reduced_data$percinv > 0)
 
-invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = percinv))
-nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
+invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = factor(invasive)))
+nonzero_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = percinv)) + 
+  geom_histogram() + stat_bin(binwidth = 10)
+log_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = log10(percinv))) + 
+  geom_histogram() + stat_bin(binwidth = 0.25)
 
-pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
-grid.arrange(all_workers_hist, nonzero_hist, nrow = 1)
+pdf("Paper/images/invasive_histograms.pdf",width=12,height=8)
+grid.arrange(invasive_hist, nonzero_hist, log_hist, nrow = 1)
 dev.off()
 
-
-workers_nonzero <- reduced_data[reduced_data$fulltimework > 0, 'fulltimework']
-reduced_data$workers_binary <- (reduced_data$fulltimework > 0)*1
-
-workers_x.df <- subset(reduced_data[,-workers_binary], select = workers_include)
+invasive_binary.models <- fit.models(model.name = "invasive_binary", x.data = invasive_x.df, 
+                                   y.response = reduced_data$invasive, response.family = "binomial")
 
 
-workers.models <- fit.models(model.name = "workers", x.data = workers_x.df, 
-                                y.response = reduced_data$workers_binary, response.family = "binomial")
+log_invasive_nonzero <- log10(reduced_data[reduced_data$percinv > 0, 'percinv'])
+
[…]

### 2018-02-27 landuse_model.R

@@ -1,0 +1,40 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(tidyverse)
+library(ggplot2)
+library(readxl)
+library(glmnet)
+library(mpath)
+library(reshape2)
+library(gridExtra)
+library(leaps)
+library(bestsubset)
+
+source("modeling.function.R")
+
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+
+landuse_include <- vars[which(is.na(vars$`Land use`)),1]$`Variable Name`
+landuse_x.df <- subset(reduced_data, select = landuse_include)
+
+perc_vars <- c(landuse_x.df$percperm, landuse_x.df$perctemp)
+
+survey_vars <- c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush')
+reclass_vars <- c('ReclassAGRICOLA','ReclassCONSERVACION', 'ReclassFORESTAL', 'ReclassHABITACIONAL',
+                  'ReclassPECUARIO', 'ReclassSIN APROVECHAMIENTO')
+
+
+perc_long1 <- melt(landuse_x.df[,survey_vars])
+pdf("Paper/images/landuse_boxplots.pdf",width=12,height=8)
+ggplot(perc_long1, aes(x=factor(variable),y=value)) + geom_boxplot() + xlab("Land Use Category") + ylab("Percent")
+dev.off()
+
+perc_long2 <- melt(landuse_x.df[,reclass_vars])
+ggplot(perc_long2, aes(x=factor(variable),y=value)) + geom_boxplot()
+
+primary_use <- factor(apply(landuse_x.df[,c('percperm','perctemp','perctill','percpasture','percbrush')], 1, which.max))
+table(primary_use)
+
+landuse.models <- fit.models(model.name = "landuse", x.data = landuse_x.df, 
+                                     y.response = primary_use, response.family = "multinomial")

### 2018-02-27 modeling.function.R

@@ -1,9 +1,11 @@
 fit.models <- function(model.name, x.data, y.response, response.family){
   
   # For testing
-  x.data <- workers_x.df
-  y.response <- reduced_data$workers_binary
-  response.family <- "binomial"
+  
+  model.name <- "landuse"
+  x.data <- landuse_x.df
+  y.response <- primary_use
+  response.family <- "multinomial"
   
   x.matrix <- model.matrix(~., x.data)[,-1]
   
@@ -15,12 +17,21 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     cv.75=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=0.75)
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=.5)
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,foldid=foldid,alpha=0.25)
-  } else {
+  }
+  if (response.family == "binomial") {
+    cv1=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=1)
+    cv.75=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.75)
+    cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
+    cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
+  } 
+  if (response.family == "multinomial") {
     cv1=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=1)
     cv.75=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.75)
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
   }
+  
+ 
   models <- vector(mode="list", length=4)
   
   models[[1]] <- cv.25
[…]

### 2018-02-27 netincome_model.R

@@ -21,14 +21,34 @@ netincome_include <- vars[which(is.na(vars$`Net Income`)),1]$`Variable Name`
 
 full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome))
 reduced_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + 
-  scale_x_continuous(limits = c(-5e+04, 5e+04))
+  scale_x_continuous(limits = c(-2e+04, 2e+04))
 pdf("Paper/images/netincome_histograms.pdf",width=12,height=8)
 grid.arrange(full_hist, reduced_hist, nrow = 1)
 dev.off()
 
+netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < 2e+04,], select = netincome_include)
+netincome <- reduced_data[abs(reduced_data$netincome) < 2e+04, 'netincome']
 
-netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < 5e+04,], select = netincome_include)
-netincome <- reduced_data[abs(reduced_data$netincome) < 5e+04, 'netincome']
 
-netincome.models <- fit.models(model.name = "netincome", x.data = netincome_x.df, 
+netincome.matrix <- model.matrix(~., netincome_x.df)[,-1]
+col_vars <- apply(netincome.matrix, 2, var)
+zero_variance <- names(col_vars[col_vars == 0])
+
+netincome_x.nonzero <- netincome.matrix[,!colnames(netincome.matrix) %in% zero_variance]
+
+
+### Diagnostics to find linear dependencies and correlations between variables
+your.matrix <- netincome_x.nonzero
+rankifremoved <- sapply(1:ncol(your.matrix), function (x) qr(your.matrix[,-x])$rank)
+which(rankifremoved == max(rankifremoved))
+
+corrs <-cor(netincome_x.nonzero)
+corrs[upper.tri(corrs)] <- NA
+diag(corrs) <- NA
+
+corrs_reshape <- melt(corrs, na.rm = TRUE)
+(high_corrs <- corrs_reshape[abs(corrs_reshape$value) > 0.9,])
+
+
+netincome.models <- fit.models(model.name = "netincome", x.data = as.data.frame(netincome_x.nonzero), 
                                 y.response = netincome, response.family = "gaussian")

### 2018-02-27 workers_model.R

@@ -25,10 +25,46 @@ workers_x.df <- subset(reduced_data, select = workers_include)
 
 all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = factor(workers_binary)))
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
+log_nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = log10(fulltimework)))
 
 pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
-grid.arrange(all_workers_hist, nonzero_hist, nrow = 1)
+grid.arrange(all_workers_hist, nonzero_hist, log_nonzero_hist, nrow = 1)
 dev.off()
 
-workers.models <- fit.models(model.name = "workers", x.data = workers_x.df, 
+worker_binary.models <- fit.models(model.name = "workers_binary", x.data = workers_x.df, 
                                 y.response = reduced_data$workers_binary, response.family = "binomial")
+
+
+workers_x.nonzero.pre <- subset(reduced_data[reduced_data$fulltimework > 0,], select = workers_include)
+num_workers_log <- log10(reduced_data[reduced_data$fulltimework > 0, 'fulltimework'])
+
+workers.matrix <- model.matrix(~., workers_x.nonzero.pre)[,-1]
+col_vars <- apply(workers.matrix, 2, var)
+zero_variance <- names(col_vars[col_vars == 0])
+
+workers_x.nonzero <- workers.matrix[,!colnames(workers.matrix) %in% zero_variance]
+remove_vars <- c("ga15_cualLLANTAS", "`ENERGIA_ELENERGIA SOLAR PUBLICA`", "`ENERGIA_ELGENERADOR PRIVADO`")
+workers_x.nonzero <- as.data.frame(workers_x.nonzero[,!colnames(workers_x.nonzero) %in% remove_vars])
+
+worker_nonzero.models <- fit.models(model.name = "workers_nonzero", x.data = workers_x.nonzero, 
+                             y.response = num_workers_log, response.family = "gaussian")
+
+
+
+which(duplicated(workers_x.nonzero, MARGIN = 0))
+
+
+
+### Diagnostics to find linear dependencies and correlations between variables
[…]

### 2018-03-02 Paper/paper_draft.tex
More updates

@@
 \usepackage{csvsimple}
+\usepackage{subcaption}
 
@@
   University of North Carolina at Chapel Hill\\
-  \texttt{mgb2188@live.unc.edu} \\
+	\\
+  \bf{Client: Francisco Laso} \\
+  Department of Geography \\
+  University of North Carolina at Chapel Hill\\
+  \\
+  March 8th, 2018 \\
 }
-
-\begin{filecontents*}{scientists.csv}
-name,surname,age
-Albert,Einstein,133
-Marie,Curie,145
-Thomas,Edison,165
-\end{filecontents*}
 
@@
 The data available to study the dynamics between agricultural measures and related factors primarily 
-comes from the Census of Agricultural Production Units (UPA) of Galapagos (Censo de las Unidades 
-de Producci\'on Agropecuaria (upa) de Gal\'apagos). This is a self-reported survey with data from 755 
+comes from the Censo de las Unidades de Producci\'on Agropecuaria (upa) de Gal\'apagos (Census of Agricultural Production 
+Units (UPA) of Galapagos) ref census. This is a self-reported survey with data from 755 
 farms detailing the production and sale of crops and livestock, agricultural expenses, and land use 
-decisions.
+decisions. The response rate for this was ??. In addition to the census, data is also available from ... including information on water, energy and road access.
 
-In addition to the census, data is also available from ... including information on water, energy and road access.
+categories...
 
-In total, under the direction of the client 240 variables were selected for consideration in modeling 
-relationships between predictors and the outcome variables of interest.
+In total, under the direction of the client 239 variables were selected for consideration in modeling 
+relationships between predictors and five outcome variables of interest. Some of the five outcome variables of interest come 
+directly from survey responses, while others were derived from a combination of multiple variables. When modeling derived 
+outcome variables all variables used in its calculation were removed from consideration. The client also denoted specific 
+predictor variables to exclude from particular models when their inclusion would not be beneficial. For example, while the 
+amount of crops sold in pounds was not directly used to calculate net income, the obvious relationship existent precluded it 
+from inclusion. In addition, predictor variables that met one or more of the following criteria were removed prior to modeling: 
+zero variance, extremely high (>0.99) or perfect correlation with other predictor variables, or linear dependence with other 
+predictor variables (that is, two or more predictor variables could be linearly combined to create another predictor variable). The 
+exact number of predictor variables included in each model varied slightly, but there were approximately 200 predictors variables examined for each model after the preceding steps were taken.
+
+\subsection{Organization of Report}
+
+The remainder of this report is divided into four sections, Section 2: Modeling, Section 3: Results, Section 4: Statistical 
+Methods, and Section 5: Limitations and Future Work.
+Section 2 provides an brief overview of the analysis so that the results can be understood.
+In Section 3 results for each of the 5 outcome variables is provided, where standard tables and graphs are repeated for each.
+A more in-depth explanation of the statistical methods used is contained in Section 4, but this section can be referenced as needed.
+Section 5 details important considerations when interpreting the results and suggests possible avenues for future work.
+Lastly, References and the Appendix, including additional tables and figures, can be found at the end.
 
@@
 
-The primary challenge in this analysis is the vast amount of potential predictor variables. This challenge is twofold; a.) when 
-the number of predictors is large the calculation of a reliable model is difficult and b.) interpreting the coefficients of many 
+The primary challenge in this analysis is the vast amount of potential predictor variables. This challenge is twofold; 1.) when 
+the number of predictors is large the calculation of a reliable model is difficult and 2.) interpreting the coefficients of many 
 predictors simultaneously is not an easy task for humans (and will make resulting simulations overly complicated). For this 
-reason, the analysis focuses on the use of several variable selection techniques and a comparison of their performance. 
-These variable selection techniques aim to build a linear model with a subset of the available variables that still maintains 
-a strong predictive performance.
+reason, the analysis focuses on the use of two variable selection techniques that aim to build a linear model with a subset of 
+the available variables that still maintains a strong explanatory/predictive performance.
 
-Secondarily, when performing standard linear regression the outcome variable is assumed to be roughly normally 
-distributed. When this is not true, as is the case for several outcome variables in this study, a poorly fitting model will be 
-found with unreliable coefficients. In order to address this issue transformations to the data and modifications to the 
-standard linear model will be considered.
+Secondarily, when performing standard linear regression the error is assumed to be normally 
+distributed, which means the outcome variable should be roughly normally distributed. When this is not true, as is the case for 
+several outcome variables in this study, a poorly fitting model will be found with unreliable coefficients. In order to address this 
+issue transformations to the data and modifications to the standard linear model will be considered.
 
@@
 we build a set of linear models using the appropriate subset of predictors. Each relationship is modeled 
-using Best Subset, Forward Selection and ElasticNet regression. Best Subset considers all variable combinations to find the 
-linear model of best fit. Forward Selection fits a linear model by progressively adding variables to the model until a best fit is 
-found. This results in only some of the variables being included, chosen in a discrete manner. ElasticNet regression fits a linear 
-model by constraining the coefficients so that they are smaller than in standard least squares, and for many variables actually 
-shrunken to zero. Similar to Forward Selection this results in a sparser model, but variable selection can be carefully tuned 
-as optimization is done in a more continuous way.
+using Forward Selection and ElasticNet regression. Forward Selection fits a linear model by progressively adding variables to 
+the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner. 
+ElasticNet regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
+squares, and for many variables actually shrunken to zero. Similar to Forward Selection this results in a smaller model, but 
+variable selection can be carefully tuned as optimization is done in a more continuous way. In general, these techniques have 
+slightly different aims. Forward Selection chooses a model that best explains the variance in the dataset at hand. Elasticnet 
+chooses a model that can best make predictions on new data. Depending on the goals of analysis, one technique is not 
+necessarily better so we do not compare the two quantitatively, but instead offer both results as varying perspectives on 
+variable selection.
+While a variable being chosen by both methods provides stronger evidence that an important relationship exists, disagreement should suggest exploring both possibilities instead of one method necessarily being incorrect.
 
@@
 
-We define farm success with three different measurements: productivity, net income and number of workers supported.
-Productivity is calculated as the total pounds of crops and livestock produced divided by the farm surface area.
-Net income is calculated as the difference between the earnings from all products sold and total expenses.
-Number of workers supported is calculated as the total labor expenditures divided by a standard worker's salary.
+The first three outcome variables of interest can be grouped together under the category of farm success; labeled as 
+productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock 
+produced divided by the farm surface area. Net income is calculated as the the earnings from all products 
+sold minus total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard 
+full-time worker's salary.
 
@@
 
-Response variable plots and discussion
 \begin{figure}[h]
 \centering
-\includegraphics[width = 0.75\textwidth]{production_histograms.pdf}
+\includegraphics[width = 0.8\textwidth]{production_histograms.pdf}
 \caption{Histogram of Productivity...}
+\label{figure:prod_hist}
 \end{figure}
+
+The histograms of the Productivity variable (Figure \ref{figure:prod_hist}) show a strong skewness, both when looking at all 
+observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
+(bell-curved), which will benefit the linear model, we take the $log_{10}$ transformation with resulting data shown in the last 
+plot. Since the log transformation cannot be performed on zero values, we remove the ?? occasions of this from the dataset. 
+Beyond the mathematical constraint, farms with zero production perhaps are not farms as typically defined.
 
@@
 \centering
-\begin{tabular}{c|c|c}\hline%
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} & \bfseries \underline{Best~Subset}
-\csvreader[head to column names]{production_top10names.csv}{}%
-{\\\elasticnet & \forward & \subset}%
+\begin{tabular}{lc|cl}\hline%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} 
+\csvreader[head to column names]{production_top5names.csv}{}%
+{\\\elasticnet & \forward}%
 \\\hline
@@
 \caption{Modeling of Productivity, Top 5 features for each of the three methods}
-\label{table:1}
+\label{table:prod_top5}
 \end{table}
 
-\begin{figure}
-\includegraphics[width = \textwidth]{elastic_cv_production.pdf}
-\caption{Cross Validation plots for Elasticnet...}
+We build linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
+productivity variable, recording an optimal model of any size and the best 5 variable model for each. The size of 5 variables is 
+chosen for its interpretability and not for any specific statistical property. The results of the best 5 variable model are shown in 
+Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated with 
+a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made by 
+each method. The Root Mean-Squared Error (RMSE) for the 5 variable model Elasticnet model is 0.78 and the $R^2$ for 
+Forward Selection is 0.48. The RMSE is on average how far the predicted value is from the actual value and $R^2$ is the 
+percentage of variability in the outcome variable is explained by the model.
+
+Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-validation plot for Elasticnet can be understood as follows: the 
+horizontal axis shows the number of variables included in the model (on top) and the corresponding lambda ($\lambda$) value 
+(on the bottom), the vertical axis shows the Mean-Squared Error (MSE) represented as the red dots and surrounded by bars 
+showing the standard deviation. The vertical dashed line to the left, $\lambda_{min}$, is found at the minimum MSE and the 
+vertical dashed line to the right, $\lambda_{1se}$ is at the largest lambda within one standard error of the minimum. The idea 
+behind $\lambda_{1se}$ is that similar error performance can be achieved with a smaller model, in this case a model with 40 
+less variables. Since our goal is to select a small amount of variables, we will generally use the model found at $\lambda_{1se}
+$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus the 
+information criterion that we would like to minimize and mark the optimal point in red.
+
+The convex shape of the plots highlights a common trend in variable selection; not including enough variables does not provide 
+enough information, but beyond a certain point adding more variables may not be worth the added complexity. The optimal 
+model chosen for Elasticnet includes 45 variables (not counting the intercept term) and using Forward Selection we choose a 
+model of 26 variables, almost all of which are also included in the Elasticnet model. The coefficients estimated for both models 
+can be found in Table \ref{table:prod_full} in the Appendix. For the full models the RMSE and $R^2$ are 0.61 and 0.63 
+respectively. These numbers suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
+gained by adding more variables.
+Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.
+
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
+\caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Production Variable Selection}
+\label{figure:prod_opt}
 \end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width = 0.8\textwidth]{forward_nvars_production.pdf}
-\caption{Number of variables Forward Selection plots...}
-\end{figure}
-
-Bootstrap plots
 
@@
 \caption{Histogram of Net Income...}
+\label{figure:netincome_hist}
+\end{figure}
+
+The histograms for Net Income (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and a few 
+observations wide in the tails. We attempt to fit a model on the full dataset, but find the observations with large absolute 
+values are obscuring other possible information in the model. We then remove the 28 observations that are beyond one 
+standard deviation (30,478) from the mean in either direction. The following models are fit on this reduced dataset of 727 
+observations.
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{c|c|c}\hline%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{netincome_top5names.csv}{}%
+{\\\elasticnet & \forward}%
+\\\hline
+\end{tabular}
+\caption{Modeling of Net Income, Top 5 features for each of the three methods}
+\label{table:netincome_top5}
+\end{table}
+
+The results of the best 5 variable model are shown in \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection.
+For interpretation of the RMSE it is important to keep in mind the scale for Net Income is much larger than that of log productivity.
+
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_netincome.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_netincome.pdf}
+\caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Net Income Variable Selection}
+\label{figure:netincome_opt}
+\end{figure}
+
+The cross-validation plot for Net Income show wider error bars throughout the range of model sizes and just using the average 
+Net Income would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
+choose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
+include 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
+found in Table \ref{table:netincome_full} in the Appendix. The full models had an RMSE and $R^2$ of 5768.30 and 0.19, 
+respectively.
+Since the full models are not much larger than the 5 variable models, the small increases are not surprising.
+Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure
+\ref{figure:resids_netincome} in the Appendix) do not follow assumptions as closely as for the Production models, but are not so 
+concerning as to disqualify the results.
+On the whole, the results from the various plots and diagnostics suggest that the relationships found for Net Income are worth investigating, but that a linear relationship is not very strong.
+
+\subsubsection{Number of Workers Supported}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width = 0.75\textwidth]{worker_histograms.pdf}
+\caption{Histogram of Workers...}
+\label{figure:workers_hist}
+\end{figure}
+
+The first plot in Figure \ref{figure:workers_hist} shows that a large percentage of the farms are not able to support any workers. 
+For this reason, we divide up the modeling task for this outcome variable. First, we use logistic regression to model the binary 
+variable of whether a farm supports more than zero or zero workers. Secondly, we use linear regression to model the quantity 
+of workers for just those 264 farms with a positive number of workers supported.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{c|c|c}\hline%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{workers_binary_top5names.csv}{}%
+{\\\elasticnet & \forward}%
+\\\hline
+\end{tabular}
+\caption{Modeling of Binary Workers , Top 5 features for each method}
+\label{table:workers_binary_top5}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\begin{tabular}{c|c|c}\hline%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
+{\\\elasticnet & \forward}%
+\\\hline
+\end{tabular}
+\caption{Modeling of Nonzero Workers , Top 5 features for each method}
+\label{table:workers_nonzero_top5}
+\end{table}
+
+Sometimes Elasticnet will simultaneously choose to include multiple variables at the same time, in this case there is not a 5 
+variable model so we show the 6 variable model for Elasticnet. For logistic regression we can measure performance with the 
+misclassification rate, which on average is 0.31 on the test set for Elasticnet and 0.26 on the full dataset for Forward Selection. 
+For the linear model 5 variable model the RMSE is 0.46 and the $R^2$  0.30.
+Several variables are found to be most helpful for both the logistic regression and linear regression models, but there is still a fair bit of difference.
+
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_workers_binary.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_workers_binary.pdf}
+\caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Workers Binary Variable Selection}
+\label{figure:workers_binary_opt}
+\end{figure}
+
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_workers_nonzero.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_workers_nonzero.pdf}
+\caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Workers Positive Variable Selection}
+\label{figure:workers_nonzero_opt}
+\end{figure}
+
+For logistic regression Elasticnet chooses a model of size 6 and Forward Selection chooses a model of size 7. Since these 
+models are not much larger the misclassification rates remain at 0.31 and 0.26, respectively. For linear regression, the full 
+models for Elasticnet and Forward Selection have a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
+has 5 variables and the Forward Selection full model has 8 variables.  In the Appendix, the coefficients estimated for logistic 
+regression models can be found in Table \ref{table:workers_binary_full} and the coefficients estimated for linear regression 
+models can be found in Table \ref{table:workers_nonzero_full}.
+
+\subsection{Invasive Species}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
+\caption{Histogram of Invasive...}
+\label{figure:invasive_hist}
 \end{figure}
@@
 \begin{tabular}{c|c|c}\hline%
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} & \bfseries \underline{Best~Subset}
-\csvreader[head to column names]{netincome_top10names.csv}{}%
-{\\\elasticnet & \forward & \subset}%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{invasive_binary_top5names.csv}{}%
+{\\\elasticnet & \forward}%
 \\\hline
 \end{tabular}
-\caption{Modeling of Net Income, Top 5 features for each of the three methods}
-\label{table:1}
+\caption{Modeling of Binary Invasive , Top 5 features for each method}
+\label{table:invasive_binary_top5}
 \end{table}
 
-\begin{figure}
-\includegraphics[width = \textwidth]{elastic_cv_netincome.pdf}
-\caption{Cross Validation plots for Elasticnet...}
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_invasive_binary.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_invasive_binary.pdf}
+\caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Invasive Binary Variable Selection}
+\label{figure:invasive_binary_opt}
 \end{figure}
 
-\begin{figure}
+\begin{table}[h!]
 \centering
-\includegraphics[width = 0.8\textwidth]{forward_nvars_netincome.pdf}
+\begin{tabular}{c|c|c}\hline%
+\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{invasive_nonzero_top5names.csv}{}%
+{\\\elasticnet & \forward}%
+\\\hline
+\end{tabular}
+\caption{Modeling of Nonzero Invasive , Top 5 features for each method}
+\label{table:invasive_nonzero_top5}
+\end{table}
+
+\begin{figure}[!tbp]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_invasive_nonzero.pdf}
+\caption{Cross-validation plots for Elasticnet...}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_invasive_nonzero.pdf}
 \caption{Number of variables Forward Selection plots...}
+\end{subfigure}
+\caption{Invasive Positive Variable Selection}
+\label{figure:invasive_nonzero_opt}
 \end{figure}
-
-\subsubsection{Number of Workers Supported}
-
-\subsection{Invasive Species}
 
@@
 
-Standard Linear Regression can be represented in matrix form as equation 1 above, where $\bY$ is a $n \times 1$ vector of 
-the outcome variable, $\bX$ is a $n \times p$ matrix of the predictor variables, $\bbeta$ is a $p \times 1$ vector of variable coefficients and $\epsilon$ is the error 
-term. The standard linear model assumes the outcome variable $\bY$ has a normal distribution, and therefore takes 
-continuous values. When the outcome variable is continuous, but not normal shaped (perhaps skewed like the productivity 
+Standard Linear Regression can be represented in matrix form as seen in equation 1 above, When there are $n$ observations and $p$ predictor variables,  $\bY$ is a $n \times 1$ vector of the outcome variable, $\bX$ is a $n \times p$ matrix of predictor variables, $\bbeta$ is a $p \times 1$ vector of variable coefficients and $\epsilon$ is the error
+term. The standard linear model works best when the outcome variable $\bY$ has a normal distribution, and therefore takes 
+continuous values. When the outcome variable is continuous, but not normal shaped (e.g., skewed like the productivity 
 data) it can be possible to transform the data by taking the logarithm or something similar. However, when the outcome 
-variable is discrete (such as labels of 1 and 0 denoting absence/presence of agricultural land use) a further modification must be made.
+variable is discrete (such as binary labels of 1 and 0 denoting absence/presence of a feature) a further modification must be made.
 The outcome variable is clearly no longer normally distributed, as it is not even continuous.
@@
 $\bX\beta$ can still be mapped to a continuous scale. In some respects this is a computational concern, but it also changes the 
-way coefficients can be interpreted. Instead of a one unit change in $X_1$ predicting a $\beta_1$ change in the predicted $Y$, 
+way coefficients can be interpreted. For example, instead of a one unit change in $X_1$ predicting a $\beta_1$ change in the predicted $Y$, 
 in this case it predicts a $\beta_1$ change in the log odds of $Y$.
@@
 
-There are many measures of the fit of linear models. Since there are many possible predictor variables, care must be taken 
+There are many measures of fit for linear models. When there are many possible predictor variables, care must be taken 
 to use appropriate measures, as some measures will favor just adding all of the variables to the model. For example, if we 
-aim to minimize the mean square error adding more predictor variables to the model will always be encouraged.
-Since that is not desired, measurements like Bayesian Information Criterion (BIC) and Mallow's $C_p$ ($C_p$) can used here.
-BIC and $C_p$ are a combination of how well the model fits the data and a penalty term for the number of predictor variables included in the model.
-The goal is to minimize BIC and $C_p$, that is the model with the best balance of small size and goodness of fit.
-BIC puts a large penalty on the inclusion of additional variables, so it will be our primary metric.
+aim to minimize the mean square error adding more predictor variables to the model will always be encouraged. Since that is 
+not desired, measurements like Bayesian Information Criterion (BIC) can be used. BIC is a combination of how well the model 
+fits the data and a penalty term for the number of predictor variables included in the model. The goal is to minimize BIC, that is 
+the model with the best balance of small size and goodness of fit. BIC is chosen over other potential measurements because it 
+puts a large penalty on the inclusion of additional variables.
 
-Another approach is to use cross-validation. In this technique the dataset is first split into $n$ equally sized sets. Then a 
-model is fit using $n - 1$ of the sets (training sets) and evaluated on the remaining 1 (test) set. This is repeated $n$ times, 
-each time reserving a different 1 test set, and then results across the $n$ runs are averaged. The benefit of this is that model 
+Another approach is to use cross-validation. In this technique the dataset is first split into $k$ equally sized sets. Then a 
+model is fit using $k - 1$ of the sets (training sets) and evaluated on the remaining 1 (test) set. This is repeated $k$ times, 
+each time reserving a different 1 test set, and then results across the $k$ runs are averaged. The benefit of this is that model 
 building and model evaluating are happening on different portions of the data, so we can distinguish if the model is picking up 
 on generalizable patterns or just random noise. Using cross-validation the average test set mean square error is an 
-appropriate measure of model fit.
+appropriate measure of model fit. We can also capture the standard deviation across the $k$ runs to measure variability, which is shown in the error bars of the Elasticnet cross-validation plots.
 
@@
 are not included. Such a constraint might be limiting the number of variables included or that the model found can generalize to 
-other data.
-Best subset selection, the most natural, but computationally difficult way is to try all possible combinations of variables and select the best fitting combination.
-However, when the number of variables, $p$, is large this quickly becomes infeasible, as there are $2^p$ possible combinations.
-Recent advances have expanded the problem sizes that can be efficiently computed using Mixed Integer Optimization.
-Using this technique we find the best 5 variable models for each of the outcomes, but are not able to search all possible model sizes to find the optimal number of variables.
-In addition, as marked in the results above, sometimes the algorithm can verify that it has found the optimal solution, while other times it returns the best solution found in the allotted time limit (set to 60 seconds), that may or may not be the overall optimal solution.
+other data. Best subset selection, the most natural, but computationally difficult way is to try all possible combinations of 
+variables and select the best fitting combination. However, when the number of variables, $p$, is large this quickly becomes 
+infeasible, as there are $2^p$ possible combinations.
 
@@
 predictor variables, which is what Forward Selection does. In this algorithm, we start with an empty model and iteratively add a 
-new variable at each stage that is most beneficial.
-This procedure can work well, but may not find the optimal solution.
-As an example, consider a case where $X_1$ is the single most predictive variable, but the combination of $X_2$ and $X_3$ is the best two variable combination.
-The algorithm will first add $X_1$, but then regardless whether it adds $X_2$ or $X_3$ next, it will have found a suboptimal solution.
-In general, we can decide to stop adding variables once we have reached an optimal performance measure like BIC or cross-validation test error.
+new variable at each stage that most increases the fit. This procedure can work well, but may not find the optimal solution. As 
+an example, consider a case where $X_1$ is the single most predictive variable, but the combination of $X_2$ and $X_3$ is 
+the best two variable combination. The algorithm will first add $X_1$, but then regardless whether it adds $X_2$ or $X_3$ next, 
+it will have found a suboptimal solution. In general, we can decide to stop adding variables once we have reached an optimal 
+performance measure like BIC or cross-validation test error.
 
 \subsection{Regularized Regression}
-
-% Consider taking these out of matrix form
 
@@
 
-As shown in equation ? in the standard linear model, we try to find the $\beta$ that minimizes the squared difference between 
-the true $\bY$ and the predicted $\hat{\bY}$ (which is $\beta\bX$). In regularized regression we do the same thing, but also 
-add a second term that we look to simultaneously minimize. This second term adds a penalty for increasing values of $
-\beta$, so the two terms must be balanced. The optimal model will find a balance between fitting the outcome variable 
-closely, but not having too large of coefficient values. The difference between Ridge regression and LASSO is how we add up 
-the coefficients. In Ridge Regression the coefficients are squared and then summed, in LASSO we take the absolute value of 
+The above notation of $||\cdot||_{2}^{2}$ and  $||\cdot||$ are defined in general as: $||\bX||_{2}^{2} = x_1^2 + x_2^2 + ... + 
+x_n^2$ and $||\bX|| = |x_1| + |x_2| + ... + |x_n|$. As shown in equation ? in the standard linear model, we try to find the $\beta$, that is a vector of coefficients,  
+that minimizes the squared difference between the true $\bY$ and the predicted $\hat{\bY}$ (which is $\beta\bX$). In 
+regularized regression we do the same thing, but also add a second term that we look to simultaneously minimize. This second 
+term adds a penalty for increasing values of $\beta$, so the two terms must be balanced. The optimal model will find a balance 
+between fitting the outcome variable closely, but not having too large of coefficient values. The difference between Ridge 
+regression and LASSO is how we add up the coefficients. In Ridge Regression the coefficients are squared and then summed, 
+in LASSO we take the absolute value of 
 the coefficients and then sum them. LASSO will encourage most of the coefficients to go to zero, thus only including a small 
 number of terms in the model. Ridge regression will encourage the coefficient values to be spread out among predictor 
-variables, leaving all of the variables in the model, but helping to offset negative effects of correlated predictor variables. Further explanations of this can be found in ???
+variables, leaving all of the variables in the model, but helping to offset negative effects of correlated predictor variables.
 
@@
 As can be seen in the equation above both the square of the coefficients and the absolute value of the coefficients is included, with the contribution of each controlled by the size of $\alpha$ which takes a value between 0 and 1.
-ElasticNet, thus combines the favorable properties of Ridge and LASSO in that it can achieve both sparse models and deal with correlated predictor variables.
+ElasticNet, thus combines the favorable properties of Ridge and LASSO, in that it can achieve both sparse models and can handle correlated predictor variables.
 Both the $\lambda$ and the $\alpha$ can be set using cross-validation (as discussed above) to be appropriate values for the particular dataset.
 
-\subsection{Comparison of methods}
+\section{Limitations and Future Work}
+There are a few key considerations that should be kept in mind when interpreting this analysis.
+First, is that relationships discovered in this analysis are correlational nature and cannot be assumed to be causal.
+Just because farms with a higher coverage of invasive species have lower productivity does not necessarily mean the invasive species causes lower productivity.
+It could be that lower productivity causes higher invasive species coverage.
+Or there could other factors not captured in the model that influence both productivity and invasive species.
+In order to determine causality, relationships of interest should be tested in a designed experiment.
 
-\section{Conclusion}
+Secondly, p-values and confidence intervals for coefficients were intentionally not included in the analysis. In standard 
+regression analysis we pre-specify the model and then test which variables are found to be significant. However, when using 
+Elasticnet and Forward Selection like we have done here, we do not specify the model ahead of time, but instead let the data 
+decide the model.
+This violates the significance test assumption and can lead to misleadingly small p-values.
+While there are some advanced techniques to try to adjust for this, it is recommended to view the results in this report as an exploratory analysis rather than definitive evidence.
+
+Future analysis might look to explore better fitting relationships, particularly for the outcome variables that had poor RMSE and $R^2$ values.
+The relationships modeled in this report only considered linear combinations of predictor variables to predict/explain the outcome variables.
+Modifications could include adding interaction terms (i.e., $x_1x_2$) or nonlinear terms (i.e., $x_1^2$ or binary transformations $x_1 > 10$).
+Exploring all possible modifications of this type is not computationally feasible, but with domain knowledge a subset of theorized relationships could be tested.
+Lastly, the variables used here primarily covered socioeconomic dimensions.
+The addition of physical and biotic variables may help better predict/explain the outcome variables or may change the importance of previously highlighted socioeconomic variables.
 
@@
 \caption{Full coefficient list for Production model}
+\label{table:prod_full}
 \end{table}
+
+\begin{figure}[h]
+\includegraphics[width = \textwidth]{resids_production.pdf}
+\caption{Diagnostic Residual plots for Production.
+The plots on the left show the predicted values vs.
+the residuals and both demonstrate a desirable lack of pattern.
+The plots of the right examine the normality of the residuals, both staying close to the desired straight diagonal pattern.
+}
+\label{figure:resids_prod}
+\end{figure}
 
@@
 \caption{Full coefficient list for Net Income model}
+\label{table:netincome_full}
 \end{table}
@@
 \includegraphics[width = \textwidth]{resids_netincome.pdf}
-\caption{Diagnostic Residual plots for Net Income...}
+\caption{Diagnostic Residual plots for Net Income.
+The plots of the predicted values vs.
+the residuals both demonstrate a desirable lack of pattern.
+The plots examining the normality of the residuals do not follow a diagonal line as closely as would be hoped, but not an extreme departure.}
+\label{figure:resids_netincome}
+\end{figure}
+
+\begin{table}[h!]
+\begin{tabular}{l|c|c}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{fullcoeflist_workers_binary.csv}{}%
+{\\\variable & \elastic & \forward}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Binary Workers model}
+\label{table:workers_binary_full}
+\end{table}
+
+\begin{table}[h!]
+\begin{tabular}{l|c|c}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{fullcoeflist_workers_nonzero.csv}{}%
+{\\\variable & \elastic & \forward}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Nonzero Workers model}
+\label{table:workers_nonzero_full}
+\end{table}
+
+\begin{figure}[h]
+\includegraphics[width = \textwidth]{resids_workers_nonzero.pdf}
+\caption{Diagnostic Residual plots for Nonzero workers.}
+\label{figure:resids_workers_nonzero}
+\end{figure}
+
+\begin{table}[h!]
+\begin{tabular}{l|c|c}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{fullcoeflist_invasive_binary.csv}{}%
+{\\\variable & \elastic & \forward}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Binary Invasive model}
+\label{table:workers_invasive_full}
+\end{table}
+
+\begin{table}[h!]
+\begin{tabular}{l|c|c}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\csvreader[head to column names]{fullcoeflist_invasive_nonzero.csv}{}%
+{\\\variable & \elastic & \forward}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Nonzero Invasive model}
+\label{table:invasive_nonzero_full}
+\end{table}
+
+\begin{figure}[h]
+\includegraphics[width = \textwidth]{resids_invasive_nonzero.pdf}
+\caption{Diagnostic Residual plots for Nonzero Invasive.}
+\label{figure:resids_invasive_nonzero}
 \end{figure}

### 2018-03-02 landuse_model.R

@@ -34,7 +34,14 @@ perc_long2 <- melt(landuse_x.df[,reclass_vars])
 ggplot(perc_long2, aes(x=factor(variable),y=value)) + geom_boxplot()
 
 primary_use <- factor(apply(landuse_x.df[,c('percperm','perctemp','perctill','percpasture','percbrush')], 1, which.max))
+primary_use <- factor(apply(landuse_x.df[,survey_vars], 1, which.max))
 table(primary_use)
+
+perc_vars <- as.matrix(reduced_data[,survey_vars])
 
 landuse.models <- fit.models(model.name = "landuse", x.data = landuse_x.df, 
                                      y.response = primary_use, response.family = "multinomial")
+
+summary(landuse_x.df$ReclassPECUARIO)
+summary(landuse_x.df$`ReclassSIN APROVECHAMIENTO`)
+summary(landuse_x.df$percpasture)

### 2018-03-02 modeling.function.R

@@ -2,10 +2,10 @@ fit.models <- function(model.name, x.data, y.response, response.family){
   
   # For testing
   
-  model.name <- "landuse"
-  x.data <- landuse_x.df
-  y.response <- primary_use
-  response.family <- "multinomial"
+  #model.name <- "workers_binary"
+  #x.data <- workers_x.df
+  #y.response <- reduced_data$workers_binary
+  #response.family <- "binomial"
   
   x.matrix <- model.matrix(~., x.data)[,-1]
   
@@ -24,14 +24,14 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
   } 
-  if (response.family == "multinomial") {
+  if (response.family == "mgaussian") {
     cv1=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=1)
     cv.75=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.75)
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
   }
   
- 
+  
   models <- vector(mode="list", length=4)
   
   models[[1]] <- cv.25
@@ -55,11 +55,21 @@ fit.models <- function(model.name, x.data, y.response, response.family){
   #elastic.nvars <- glmnet(x.matrix,y.response,family=response.family,alpha=best_alpha, lambda = lambda.seq)
 
   found_df = FALSE
-  df = 25
+  df = 5
   while (found_df == FALSE){
     if (length(elastic.full$lambda[elastic.full$df == df]) > 0){
       lambda.5var <- elastic.full$lambda[elastic.full$df == df]  
-      elastic5.error <- best_model$cvm[best_model$glmnet.fit$df == df][1]
[…]

### 2018-03-02 netincome_model.R

@@ -21,13 +21,13 @@ netincome_include <- vars[which(is.na(vars$`Net Income`)),1]$`Variable Name`
 
 full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome))
 reduced_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + 
-  scale_x_continuous(limits = c(-2e+04, 2e+04))
+  scale_x_continuous(limits = c(-3e+04, 3e+04))
 pdf("Paper/images/netincome_histograms.pdf",width=12,height=8)
 grid.arrange(full_hist, reduced_hist, nrow = 1)
 dev.off()
 
-netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < 2e+04,], select = netincome_include)
-netincome <- reduced_data[abs(reduced_data$netincome) < 2e+04, 'netincome']
+netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome),], select = netincome_include)
+netincome <- reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome), 'netincome']
 
 
 netincome.matrix <- model.matrix(~., netincome_x.df)[,-1]

### 2018-03-02 production_model.R

@@ -19,9 +19,6 @@ production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name
 production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
 
 log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])
-
-zero_prod <- reduced_data[reduced_data$productivity ==0,]
-par(mfrow = c(2,1))
 raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity))
 log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity)))
 lower_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) +

### 2018-03-02 workers_model.R

@@ -19,13 +19,17 @@ vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
 workers_include <- vars[which(is.na(vars$`Number workers supported`)),1]$`Variable Name`
 
 workers_nonzero <- reduced_data[reduced_data$fulltimework > 0, 'fulltimework']
-reduced_data$workers_binary <- (reduced_data$fulltimework > 0)*1
+reduced_data$workers_binary <- factor(reduced_data$fulltimework > 0)
 
 workers_x.df <- subset(reduced_data, select = workers_include)
 
-all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = factor(workers_binary)))
+levels(reduced_data$workers_binary) = c("Zero", "Positive")
+
+all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = workers_binary)) + xlab("fulltimework")
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
 log_nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = log10(fulltimework)))
+
+#log_nonzero_hist2 <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(fulltimework+0.1)))
 
 pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
 grid.arrange(all_workers_hist, nonzero_hist, log_nonzero_hist, nrow = 1)

### 2018-03-08 Paper/paper_draft.tex
completion of paper draft

@@
 \usepackage{subcaption}
+\usepackage{numprint}
 
@@
 
-\title{Survival of the Fittest: Variable Selection on Galapagos Agricultural Data}
+\title{Survival of the Fittest: Variable Selection on Agricultural Data from the Gal\'apagos Islands}
 
@@
 \begin{abstract}
-  This paper...
-
+  Variable selection is an important first step when analyzing datasets with a large number of potential predictor variables. We 
+  apply two techniques, Forward Selection and Elasticnet,  to find the most important factors in a dataset detailing over 200 
+  socioeconomic measurements for 755 farms on the Gal\'apagos Islands. Modeling five different outcome variables, we find 
+  the data available has the strongest linear relationships with Productivity and Land Use Choices. For each of the outcome 
+  variables we present the top five predictor variables as well as a full set of coefficients for the optimally predictive model.
   \end{abstract}
@@
 
-The Galapagos Islands make for a feasible and significant case study of complex systems.
-Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Galapagos Islands than other systems.
-Yet the Galapagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment.
-Prior work has created agent-based models of the Galapagos, but with limited interaction parameters between agents, particularly in regards to farm success walsh.
-In order to create a more detailed, and perhaps more accurate, simulation, the relationships between different factors on the island must be better understood.
-This work aims to sort through a large number of possible relationships and clarify the empirically most significant ones for future study and incorporation into simulations.
+The Gal\'apagos Islands make for a feasible and significant case study of complex systems.
+Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Gal\'apagos Islands than other systems.
+Yet the Gal\'apagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment.
+Prior work has created agent-based models of the Gal\'apagos, but with limited interaction parameters between agents, particularly in regards to farm success (\cite{miller}, \cite{valdivia},\cite{walsh}).
+In order to create a more detailed and perhaps more accurate simulation, the relationships between different factors on the island must be better understood.
+This work aims to search through a large number of possible relationships and identify the empirically most significant ones for future study and incorporation into simulation models.
 
@@
 comes from the Censo de las Unidades de Producci\'on Agropecuaria (upa) de Gal\'apagos (Census of Agricultural Production 
-Units (UPA) of Galapagos) ref census. This is a self-reported survey with data from 755 
-farms detailing the production and sale of crops and livestock, agricultural expenses, and land use 
-decisions. The response rate for this was ??. In addition to the census, data is also available from ... including information on water, energy and road access.
+Units (UPA) of Galapagos) (\cite{census}). This is a self-reported survey with data from 755 farms (UPAs) detailing many characteristics. The questions covered fall into the following categories:
+\begin{itemize}
+\item{General characteristics (land area, age of landowner, etc.)}
+\item{Permanent crops (specific types and quantity)}
+\item{Temporary crops (specific types and quantity)}
+\item{Pastures (specific types and quantity)}
+\item{Tree crops (specific types and quantity)}
+\item{Livestock production (detailed by animal)}
+\item{Expenses (detailed by category)}
+\item{Workers (detailed by role)}
+\item{Land use (8 different categories)}
 
-categories...
+\end{itemize}
 
-In total, under the direction of the client 239 variables were selected for consideration in modeling 
-relationships between predictors and five outcome variables of interest. Some of the five outcome variables of interest come 
-directly from survey responses, while others were derived from a combination of multiple variables. When modeling derived 
-outcome variables all variables used in its calculation were removed from consideration. The client also denoted specific 
-predictor variables to exclude from particular models when their inclusion would not be beneficial. For example, while the 
+In addition to this census, data is also available from satellite image classification including 
+information on water, energy and road access. In total, under the direction of the client 239 variables were selected for 
+consideration in modeling relationships between potential predictors and five outcome variables of interest.
+
+Some of the five outcome variables of interest came directly from survey responses, while others were derived from a 
+combination of multiple variables. When analyzing a derived outcome variable all variables used in its original calculation were 
+removed from consideration in the model . The client also denoted specific predictor variables to exclude from particular 
+models when their inclusion would not be beneficial. For example, while the 
 amount of crops sold in pounds was not directly used to calculate net income, the obvious relationship existent precluded it 
@@
 Methods, and Section 5: Limitations and Future Work.
-Section 2 provides an brief overview of the analysis so that the results can be understood.
-In Section 3 results for each of the 5 outcome variables is provided, where standard tables and graphs are repeated for each.
+Section 2 provides a brief overview of the analysis so that the results can be understood.
+In Section 3 results for each of the 5 outcome variables are provided, where standard tables and graphs are repeated for each.
 A more in-depth explanation of the statistical methods used is contained in Section 4, but this section can be referenced as needed.
@@
 
-The primary challenge in this analysis is the vast amount of potential predictor variables. This challenge is twofold; 1.) when 
-the number of predictors is large the calculation of a reliable model is difficult and 2.) interpreting the coefficients of many 
+The primary challenge in this analysis is the vast number of potential predictor variables. This challenge is twofold; 1.) when 
+the number of predictors is large the determination of a reliable model is difficult and 2.) interpreting the coefficients of many 
 predictors simultaneously is not an easy task for humans (and will make resulting simulations overly complicated). For this 
@@
 
-Secondarily, when performing standard linear regression the error is assumed to be normally 
-distributed, which means the outcome variable should be roughly normally distributed. When this is not true, as is the case for 
-several outcome variables in this study, a poorly fitting model will be found with unreliable coefficients. In order to address this 
-issue transformations to the data and modifications to the standard linear model will be considered.
+Secondarily, when performing standard linear regression the error is assumed to be normally distributed.  While this does not 
+mean the outcome variable necessarily needs to be normally distributed, large deviations from normality can cause issues. 
+Several outcome variables in this study show strong non-normality, which can contribute to a poorly fitting model and unreliable estimates of the coefficients.
+In order to address this issue transformations to the data and modifications to the standard linear model will be considered.
 
@@
 
-A brief overview of the statistical methods used is presented here to allow for understanding of results, 
-but for further details see Section refstatmethods Statistical Methods. For each of the outcome variables of interest 
-we build a set of linear models using the appropriate subset of predictors. Each relationship is modeled 
+A summary of the statistical methods used is presented here to allow for understanding of results. For further details see 
+Section 4: Statistical Methods. For each of the outcome variables of interest a set of linear models using the appropriate subset 
+of predictors was built. Each relationship was modeled 
 using Forward Selection and ElasticNet regression. Forward Selection fits a linear model by progressively adding variables to 
-the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner. 
+the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner (computed using the R package `leaps' \cite{lumley}). 
 ElasticNet regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
 squares, and for many variables actually shrunken to zero. Similar to Forward Selection this results in a smaller model, but 
-variable selection can be carefully tuned as optimization is done in a more continuous way. In general, these techniques have 
-slightly different aims. Forward Selection chooses a model that best explains the variance in the dataset at hand. Elasticnet 
-chooses a model that can best make predictions on new data. Depending on the goals of analysis, one technique is not 
-necessarily better so we do not compare the two quantitatively, but instead offer both results as varying perspectives on 
-variable selection.
-While a variable being chosen by both methods provides stronger evidence that an important relationship exists, disagreement should suggest exploring both possibilities instead of one method necessarily being incorrect.
+variable selection can be carefully tuned as optimization is done in a more continuous way (computed using the R package `glmnet' \cite{friedman}).
+
+In general, these techniques have slightly different aims. Forward Selection chooses a model that best explains the variance in 
+the dataset at hand. Elasticnet chooses a model that can best make predictions on new data. Depending on the goals of 
+analysis, one technique is not necessarily better so we do not compare the two quantitatively, but instead offer both results as 
+varying perspectives on variable selection. While a variable being chosen by both methods provides stronger evidence that an 
+important relationship exists, disagreement suggests exploring both possibilities instead of one method necessarily being 
+incorrect.
 
@@
 productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock 
-produced divided by the farm surface area. Net income is calculated as the the earnings from all products 
+produced divided by the farm surface area. Net income is calculated as the the revenue from all products 
 sold minus total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard 
@@
 \includegraphics[width = 0.8\textwidth]{production_histograms.pdf}
-\caption{Histogram of Productivity...}
+\caption{Histogram of Productivity, from left to right showing the full range, between 0-10,000 lbs/hectare, and the log transformed nonzero values.
+The log transformed values exhibit the desired normal shape.}
 \label{figure:prod_hist}
@@
 observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
-(bell-curved), which will benefit the linear model, we take the $log_{10}$ transformation with resulting data shown in the last 
-plot. Since the log transformation cannot be performed on zero values, we remove the ?? occasions of this from the dataset. 
+(bell-curved), which will benefit the linear model, we took the $log_{10}$ transformation with resulting data shown in the last 
+plot. Since the log transformation cannot be performed on zero values, we removes the 38 occasions of this from the dataset. 
 Beyond the mathematical constraint, farms with zero production perhaps are not farms as typically defined.
@@
 \centering
-\begin{tabular}{lc|cl}\hline%
+\begin{tabular}{cc}
 \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} 
@@
 {\\\elasticnet & \forward}%
-\\\hline
 \end{tabular}
-\caption{Modeling of Productivity, Top 5 features for each of the three methods}
+\caption{Modeling of Productivity, Top 5 features for both methods}
 \label{table:prod_top5}
@@
 
-We build linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
+We built linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
 productivity variable, recording an optimal model of any size and the best 5 variable model for each. The size of 5 variables is 
-chosen for its interpretability and not for any specific statistical property. The results of the best 5 variable model are shown in 
-Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated with 
-a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made by 
-each method. The Root Mean-Squared Error (RMSE) for the 5 variable model Elasticnet model is 0.78 and the $R^2$ for 
+chosen to provide quick insight and not for any specific statistical property. The results of the best 5 variable model are shown 
+in Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated 
+with a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made 
+by each method. The Root Mean-Squared Error (RMSE) for the 5 variable Elasticnet model is 0.78 and the $R^2$ for 
 Forward Selection is 0.48. The RMSE is on average how far the predicted value is from the actual value and $R^2$ is the 
-percentage of variability in the outcome variable is explained by the model.
+percentage of variability in the outcome variable that is explained by the model. To provide context for the RMSE, we can look 
+at the point furthest to the right on the cross-validation plot and see how the model would perform when no variables are 
+included in the model, that is just predicting the average outcome value.
 
-Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-validation plot for Elasticnet can be understood as follows: the 
+Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-
+validation plot for Elasticnet can be understood as follows: the 
 horizontal axis shows the number of variables included in the model (on top) and the corresponding lambda ($\lambda$) value 
@@
 behind $\lambda_{1se}$ is that similar error performance can be achieved with a smaller model, in this case a model with 40 
-less variables. Since our goal is to select a small amount of variables, we will generally use the model found at $\lambda_{1se}
-$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus the 
-information criterion that we would like to minimize and mark the optimal point in red.
+fewer variables. Since our goal is to select a small amount of variables, we will generally use the model found at $
+\lambda_{1se}$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus 
+the information criterion that we choose to minimize and mark the optimal point in red.
+
+\begin{figure}[h]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
+\caption{Cross-validation plot for Elasticnet}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
+\caption{Forward Selection plot}
+\end{subfigure}
+\caption{Productivity Variable Selection.
+The Elasticnet plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error.
+The optimal Forward Selection model includes 26 variables.}
+\label{figure:prod_opt}
+\end{figure}
 
@@
 model of 26 variables, almost all of which are also included in the Elasticnet model. The coefficients estimated for both models 
-can be found in Table \ref{table:prod_full} in the Appendix. For the full models the RMSE and $R^2$ are 0.61 and 0.63 
+can be found in Table \ref{table:prod_full} in the Appendix. For the full models, the RMSE and $R^2$ are 0.61 and 0.63, 
 respectively. These numbers suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
@@
 Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.
-
-\begin{figure}[!tbp]
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
-\caption{Cross-validation plots for Elasticnet...}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
-\caption{Number of variables Forward Selection plots...}
-\end{subfigure}
-\caption{Production Variable Selection}
-\label{figure:prod_opt}
-\end{figure}
 
@@
 \includegraphics[width = 0.75\textwidth]{netincome_histograms.pdf}
-\caption{Histogram of Net Income...}
+\caption{Histograms of Net Income, with full data and data within one standard deviation displayed. The values very far from the center are of concern when modeling.}
 \label{figure:netincome_hist}
@@
 The histograms for Net Income (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and a few 
-observations wide in the tails. We attempt to fit a model on the full dataset, but find the observations with large absolute 
-values are obscuring other possible information in the model. We then remove the 28 observations that are beyond one 
-standard deviation (30,478) from the mean in either direction. The following models are fit on this reduced dataset of 727 
-observations.
+observations wide in the tails. We attempted to fit a model on the full dataset, but find the observations with large absolute 
+values are obscuring other possible information in the model.
+After trying several cutoff thresholds and examining model fit, we  removed the 28 observations that are beyond one standard deviation (30,478) from the mean in either direction.
+The
+following models are fit on this reduced dataset of 727 observations.
 
@@
 \centering
-\begin{tabular}{c|c|c}\hline%
+\begin{tabular}{cc}
 \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
@@
 {\\\elasticnet & \forward}%
-\\\hline
 \end{tabular}
-\caption{Modeling of Net Income, Top 5 features for each of the three methods}
+\caption{Modeling of Net Income, Top 5 features for both methods}
 \label{table:netincome_top5}
@@
 
-The results of the best 5 variable model are shown in \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection.
+The results of the best 5 variable model are shown in Table \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection.
 For interpretation of the RMSE it is important to keep in mind the scale for Net Income is much larger than that of log productivity.
-
+However, in this case adding variables has caused a minimal decrease in the error.
 \begin{figure}[!tbp]
@@
 \includegraphics[width=\textwidth]{elastic_cv_netincome.pdf}
-\caption{Cross-validation plots for Elasticnet...}
+\caption{Cross-validation plot for Elasticnet...}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_netincome.pdf}
-\caption{Number of variables Forward Selection plots...}
+\caption{Forward Selection plot}
 \end{subfigure}
-\caption{Net Income Variable Selection}
+\caption{Net Income Variable Selection. The Elasticnet plot shows a much smaller improvement in prediction error and Forward Selection includes a small number of variables.}
 \label{figure:netincome_opt}
@@
 Net Income would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
-choose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
-include 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
+chose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
+included 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
 found in Table \ref{table:netincome_full} in the Appendix. The full models had an RMSE and $R^2$ of 5768.30 and 0.19, 
 respectively.
-Since the full models are not much larger than the 5 variable models, the small increases are not surprising.
+Since the full models are not much larger than the 5 variable models, the small improvements are not surprising.
 Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure
@@
 \includegraphics[width = 0.75\textwidth]{worker_histograms.pdf}
-\caption{Histogram of Workers...}
+\caption{Histograms of Workers, showing the proportion of zeroes, farms with nonzero workers, and the log transformed number of workers.
+The log transformation helps remove the skewness, but some bi-modality still remains.}
 \label{figure:workers_hist}
@@
 The first plot in Figure \ref{figure:workers_hist} shows that a large percentage of the farms are not able to support any workers. 
-For this reason, we divide up the modeling task for this outcome variable. First, we use logistic regression to model the binary 
-variable of whether a farm supports more than zero or zero workers. Secondly, we use linear regression to model the quantity 
-of workers for just those 264 farms with a positive number of workers supported.
+For this reason, we divided up the modeling task for this outcome variable. First, we used logistic regression to model the 
+binary variable of whether a farm supports zero or more than zero workers. Secondly, we used linear regression to model the 
+quantity of workers for just those 264 farms with a positive number of workers supported.
 
@@
 {\\\elasticnet & \forward}%
-\\\hline
 \end{tabular}
@@
 
-\begin{table}[h!]
+\begin{table}[h]
 \centering
-\begin{tabular}{c|c|c}\hline%
+\begin{tabular}{cc}
 \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
@@
 {\\\elasticnet & \forward}%
-\\\hline
+\newline
 \end{tabular}
-\caption{Modeling of Nonzero Workers , Top 5 features for each method}
+\caption{Nonzero Workers}
+\caption{Modeling of Nonzero Workers, Top 5 features for both methods}
 \label{table:workers_nonzero_top5}
@@
 
-Sometimes Elasticnet will simultaneously choose to include multiple variables at the same time, in this case there is not a 5 
+Sometimes Elasticnet will simultaneously choose to eliminate multiple variables at the same time, in this case there is not a 5 
 variable model so we show the 6 variable model for Elasticnet. For logistic regression we can measure performance with the 
 misclassification rate, which on average is 0.31 on the test set for Elasticnet and 0.26 on the full dataset for Forward Selection. 
+These can be compared to the 0.35 misclassification rate we would achieve if we simply predicted the majority class, zero 
+workers, every time.
 For the linear model 5 variable model the RMSE is 0.46 and the $R^2$  0.30.
+The low misclassification rate coupled with low $R^2$ show that there is a fairly clear divide between farms that can or cannot support workers, but the specific number of workers is much harder to predict.
 Several variables are found to be most helpful for both the logistic regression and linear regression models, but there is still a fair bit of difference.
 
-\begin{figure}[!tbp]
+\begin{figure}[h]
 \begin{subfigure}[b]{0.5\textwidth}
 \includegraphics[width=\textwidth]{elastic_cv_workers_binary.pdf}
-\caption{Cross-validation plots for Elasticnet...}
+\caption{Cross-validation plot for Elasticnet}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_workers_binary.pdf}
-\caption{Number of variables Forward Selection plots...}
+\caption{Forward Selection plot}
 \end{subfigure}
-\caption{Workers Binary Variable Selection}
+\caption{Workers Binary Variable Selection.
+The Elasticnet plot shows ideal predictive performance when including 5-25 variables.
+The computation for logistic regression Forward Selection stops at the optimal value so the rest of the plot is not included as before.}
 \label{figure:workers_binary_opt}
@@
 
-\begin{figure}[!tbp]
+\begin{figure}[h]
 \begin{subfigure}[b]{0.5\textwidth}
 \includegraphics[width=\textwidth]{elastic_cv_workers_nonzero.pdf}
-\caption{Cross-validation plots for Elasticnet...}
+\caption{Cross-validation plot for Elasticnet}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_workers_nonzero.pdf}
-\caption{Number of variables Forward Selection plots...}
+\caption{Forward Selection plot}
 \end{subfigure}
-\caption{Workers Positive Variable Selection}
+\caption{Nonzero Workers Variable Selection. Both plots show that adding variables does little to improve performance, highlighting a lack of linear relationship.}
 \label{figure:workers_nonzero_opt}
@@
 
-For logistic regression Elasticnet chooses a model of size 6 and Forward Selection chooses a model of size 7. Since these 
-models are not much larger the misclassification rates remain at 0.31 and 0.26, respectively. For linear regression, the full 
-models for Elasticnet and Forward Selection have a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
-has 5 variables and the Forward Selection full model has 8 variables.  In the Appendix, the coefficients estimated for logistic 
+For logistic regression Elasticnet chose a model of size 6 and Forward Selection chose a model of size 7. Since these 
+models are not much larger the misclassification rates remained at 0.31 and 0.26, respectively. For linear regression, the full 
+models for Elasticnet and Forward Selection had a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
+had 5 variables and the Forward Selection full model had 8 variables.  In the Appendix, the coefficients estimated for logistic 
 regression models can be found in Table \ref{table:workers_binary_full} and the coefficients estimated for linear regression 
@@
 
-\begin{figure}[h]
+The analysis of Invasive Species follows much the same pattern as for Number of Workers Supported. Again, we have a large 
+number of zeros in the outcome variable, farms with zero percent of their land covered by invasive species. We first modeled 
+this binary variable using logistic regression and then performed linear regression on the log transformed values for the farms 
+that have a percent coverage greater than zero. As can be seen in the far right plot of Figure \ref{figure:invasive_hist} there are 
+two very small values (0.005{\%} and 0.08{\%} prior to taking the log). They effect the fit of the model and are negligibly above 
+zero so we removed them from the linear model, leaving 155 farms for modeling.
+
+\begin{figure}[h!]
 \centering
 \includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
-\caption{Histogram of Invasive...}
+\caption{Histogram of Invasive Species.
+To the left: the proportion of zeroes, in the middle: farms with greater than zero, and to the right: the log transformed percent of invasive species.
+The two very small values on the right plot are removed.}
 \label{figure:invasive_hist}
@@
 
-\begin{figure}[!tbp]
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{elastic_cv_invasive_binary.pdf}
-\caption{Cross-validation plots for Elasticnet...}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{forward_nvars_invasive_binary.pdf}
-\caption{Number of variables Forward Selection plots...}
-\end{subfigure}
-\caption{Invasive Binary Variable Selection}
-\label{figure:invasive_binary_opt}
-\end{figure}
-
 \begin{table}[h!]
 \centering
-\begin{tabular}{c|c|c}\hline%
+\begin{tabular}{cc}
 \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
@@
 
-\begin{figure}[!tbp]
+The top 5 variable models for logistic regression and linear regression can be found in Table \ref{table:invasive_binary_top5} 
+and Table \ref{table:invasive_nonzero_top5}, respectively. For logistic regression the misclassification rate was 0.21 on the test 
+set for Elasticnet and 0.20 on the full dataset for Forward Selection. For the linear model 5 variable model the RMSE was 0.25 and the $R^2$ was 0.30.
+
+\begin{figure}[h!]
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{elastic_cv_invasive_binary.pdf}
+\caption{Cross-validation plot for Elasticnet}
+\end{subfigure}
+\hfill
+\begin{subfigure}[b]{0.5\textwidth}
+\includegraphics[width=\textwidth]{forward_nvars_invasive_binary.pdf}
+\caption{Forward Selection plots}
+\end{subfigure}
+\caption{Invasive Binary Variable Selection. The Elasticnet plot shows adding variables does not improve just predicting the majority class. In Forward Selection the BIC very quickly levels off.}
+\label{figure:invasive_binary_opt}
+\end{figure}
+
+\begin{figure}[h!]
 \begin{subfigure}[b]{0.5\textwidth}
 \includegraphics[width=\textwidth]{elastic_cv_invasive_nonzero.pdf}
-\caption{Cross-validation plots for Elasticnet...}
+\caption{Cross-validation plot for Elasticnet}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_invasive_nonzero.pdf}
-\caption{Number of variables Forward Selection plots...}
+\caption{Forward Selection plots}
 \end{subfigure}
-\caption{Invasive Positive Variable Selection}
+\caption{Nonzero Invasive Variable Selection. Similar to Figure \ref{figure:invasive_binary_opt} the lack of benefit from adding additional variables demonstrates a weak linear relationship.}
 \label{figure:invasive_nonzero_opt}
@@
 
+The cross-validation and forward selection plots in Figure \ref{figure:invasive_binary_opt} show that there was little gained from 
+adding variables to the logistic regression model. The optimal Elasticnet model only included the intercept and the optimal 
+Forward Selection model still only included 5 variables. For this reason the misclassification rates are not changed from above.
+
+For linear regression the optimal models for Elasticnet included 4 variables with a RMSE still of 0.25 and for Forward Selection 
+included 7 variables with a $R^2$ slightly increasing to 0.35.  Since the percentage of farms with no invasive species is 20{\%} 
+the logistic regression was not able to do much more than just predict zero for all farms. This combined with the low 
+performance of the linear model suggests that invasive species coverage cannot be well explained by the variables considered 
+here.
+
 \subsection{Land use choices}
+
+The analysis of Land Use choice requires a slightly different approach than used previously since the outcome variable is a 
+categorical variable, with six different classifications of land use (each measured as a percentage of total land area):
+\begin{itemize}
+	\item{Permanent Crops}
+	\item{Temporary Crops}
+	\item{Fallow Land}
+	\item{Tilled Land}
+	\item{Pasture}
+	\item{Brush}
+\end{itemize}
+
+While not a perfect ordering, the classifications can generally thought of progressing from land that has been most worked by 
+the farmer to least worked. A primary concern is that unworked land may be most susceptible to invasive species growth. For 
+this reason we sort the farms in ascending order first by percent invasive species, and then by percent brush, percent pasture, 
+etc. and graph the stacked bar plots in Figure \ref{figure:landuse_plot}. Since percent brush is given precedence in sorting the 
+purple area stands out more than the rest. Even so, it can be seen that there are farms primarily made up of each of the 
+categories as well as farms fairly evenly split between the categories.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width = 0.75\textwidth]{landuse_barplot.pdf}
+\caption{Stacked Barplot of Land Use.
+Farms are sorted in ascending order by percent invasive and then by percbrush, percpasture, etc.
+Some bars do not total 100\% because percent invasive and percent other categories are not included.
+There appear to be different classes of farms, some primarily permanent crops, some primarily temporary crops, etc.
+}
+\label{figure:landuse_plot}
+\end{figure}
+
+There are multiple ways that this outcome variable can be formulated for modeling, but we decide to label each farm with its 
+highest percentage land use category. There are only 9 such farms labeled as fallow, which is too small for modeling so we 
+remove them from the consideration. The remaining categories and the number of farms are shown in Table \ref{table:landuse_cat}.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{lc}
+Majority Category    & Number of Farms \\
+\hline
+Percperm    & 209             \\
+Perctemp    & 50              \\
+Perctill    & 36              \\
+Percpasture & 271             \\
+Percbrush   & 189            
+\end{tabular}
+\caption{Landuse Category Count}
+\label{table:landuse_cat}
+\end{table}
+
+Using this derived categorical variable we used multinomial logistic regression with Elasticnet to build a model, which found an 
+optimal set of coefficients for each category.
+The results from the 5 variable model, which had a misclassification rate of 0.39, are shown in Table \ref{table:landuse_top5}.
+Different categories may have varying number of coefficients and because Perctill had no variables included at this lambda level, it is left out of the table.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{cc}
+\bfseries \underline{Percperm} & \bfseries \underline{Perctemp}%
+\csvreader[head to column names]{landuse_top5names.csv}{}%
+{\\\percperm & \perctemp}%
+\end{tabular}
+\begin{tabular}{cc}
+\bfseries \underline{Percpasture} & \bfseries \underline{Percbrush}%
+\csvreader[head to column names]{landuse_top5names.csv}{}%
+{\\\percpasture & \percbrush}%
+\end{tabular}
+
+\caption{Modeling of Landuse , Top features for each category}
+\label{table:landuse_top5}
+\end{table}
+
+The plot of the Elasticnet cross-validation is shown in Figure \ref{figure:landuse_opt}.
+The optimal model achieves a 0.36 misclassification rate against the test set on average, which given 5 categories to choose from shows decent predictive strength.
+In the Appendix, the coefficients estimated for the optimal model can be found in Tables \ref{table:landuse_full_a} and
+\ref{table:landuse_full_b}.
+
+\begin{figure}[!tbp]
+\includegraphics[width=\textwidth]{elastic_cv_landuse.pdf}
+\caption{Land Use Cross-validation plot for Elasticnet.
+The Elasticnet model shows a large improvement over just predicting the majority class.
+Since there are now 5 categories, such a strategy would result in a 0.65 misclassification rate.}
+\hfill
+\label{figure:landuse_opt}
+\end{figure}
 
@@
 
-This calls for the use of logistic regression, in which we perform a logit transformation as seen in the equation below so that the 
+This calls for the use of logistic regression, in which we perform a logit transformation as seen in equation 2 below so that the 
 $\bX\beta$ can still be mapped to a continuous scale. In some respects this is a computational concern, but it also changes the 
@@
 
-\[log \frac{\Pr(Y = 1)}{\Pr(Y = 0)} = \bX\beta \]
+\[log \frac{\Pr(Y = 1)}{\Pr(Y = 0)} = \bX\beta \tag{2}\]
 
-This equation can be rewritten as below
+Equation can be rewritten as below in equation 3, which gives the predicted probability of an observation being class 1.
 
-\[ \Pr(Y = 1) = \frac{e^{\bX\beta}}{1 + e^{\bX\beta}} \]
+\[ \Pr(Y = 1) = \frac{e^{\bX\beta}}{1 + e^{\bX\beta}} \tag{3}\]
 
-Multiclass Logistic Regression
+This can be extended beyond binary variables to categorical variables when there are $K$ classes, referred to as Multinomial 
+Logistic Regression. Now to calculate the probability of class $k$, we use equation 4 below.
 
-\[ \Pr(Y = k) = \frac{e^{\bX\beta_k}}{\sum_{l=1}^{K-1} e^{\bX\beta_l}} \]
-
-% include Poisson function
+\[ \Pr(Y = k) = \frac{e^{\bX\beta_k}}{\sum_{l=1}^{K-1} e^{\bX\beta_l}} \tag{4}\]
 
@@
 The above notation of $||\cdot||_{2}^{2}$ and  $||\cdot||$ are defined in general as: $||\bX||_{2}^{2} = x_1^2 + x_2^2 + ... + 
-x_n^2$ and $||\bX|| = |x_1| + |x_2| + ... + |x_n|$. As shown in equation ? in the standard linear model, we try to find the $\beta$, that is a vector of coefficients,  
+x_n^2$ and $||\bX|| = |x_1| + |x_2| + ... + |x_n|$. As shown in the top equation of the standard linear model, we try to find the $\beta$, that is a vector of coefficients,  
 that minimizes the squared difference between the true $\bY$ and the predicted $\hat{\bY}$ (which is $\beta\bX$). In 
@@
 The technique that is used in this analysis is a combination of the Ridge and LASSO penalties, called ElasticNet.
-As can be seen in the equation above both the square of the coefficients and the absolute value of the coefficients is included, with the contribution of each controlled by the size of $\alpha$ which takes a value between 0 and 1.
+As can be seen in the equation above both the square of the coefficients and the absolute value of the coefficients is included, with the contribution of each controlled by the size of alpha ($\alpha$) which takes a value between 0 and 1.
 ElasticNet, thus combines the favorable properties of Ridge and LASSO, in that it can achieve both sparse models and can handle correlated predictor variables.
-Both the $\lambda$ and the $\alpha$ can be set using cross-validation (as discussed above) to be appropriate values for the particular dataset.
+Both the $\lambda$ and the $\alpha$ can be set using cross-validation (as discussed above) to appropriate values for the particular dataset.
+In each Elasticnet cross-validation plot found in the Results section, the specific $\alpha$ used is labeled at the top of the graph.
 
@@
 decide the model.
-This violates the significance test assumption and can lead to misleadingly small p-values.
+This violates the significance test assumption and can lead to misleadingly small p-values (see Chapter 6 of \cite{hastie}).
 While there are some advanced techniques to try to adjust for this, it is recommended to view the results in this report as an exploratory analysis rather than definitive evidence.
@@
 
+\bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.
+
+\bibitem{friedman} Friedman, J., Hastie, T., \& Tibshirani, R. (2009). glmnet: Lasso and elastic-net regularized generalized linear models. R package version, 1(4).
+
+\bibitem{hastie} Hastie, T., Tibshirani, R., \& Wainwright, M. (2015). Statistical learning with sparsity: the lasso and 
+generalizations. CRC press.
+
+\bibitem{lumley} Lumley, T., \& Miller, A. (2009). Leaps: regression subset selection. R package version 2.9. Online at http://CRAN. R-project.org/package= leaps.
+
+\bibitem{miller} Miller, B.
+W., Breckheimer, I., McCleary, A.
+L., Guzm\'an-Ramirez, L., Caplow, S.
+C., Jones-Smith, J.
+C., \& Walsh, S.
+J.
+(2010).
+Using stylized agent-based models for population environment research: a case study from the Gal\'apagos Islands.
+Population and environment, 31(6), 401-426.
+
+\bibitem{valdivia}Valdivia, G., Wolford, W., \& Lu, F.
+(2014).
+Border crossings: New geographies of protection and production in the Gal\'apagos Islands.
+Annals of the Association of American Geographers, 104(3), 686-701.
+
+\bibitem{walsh} Walsh S.J., Mena C.F.
+(2013) Perspectives for the Study of the Gal\'apagos Islands: Complex Systems and Human-Environment Interactions.
+In: Walsh S., Mena C.
+(eds) Science and Conservation in the Gal\'apagos Islands.
+Social and Ecological Interactions in the Gal\'apagos Islands, vol 1.
+Springer, New York, NY
+
+\bibitem{zou} Zou, H., \& Hastie, T. (2005). Regularization and variable selection via the elastic net. Journal of the Royal Statistical Society: Series B (Statistical Methodology), 67(2), 301-320.
+
 \end{thebibliography}
@@
 \begin{table}[h!]
-\begin{tabular}{l|c|c}\hline%
+\npdecimalsign{.}
+\nprounddigits{4}
+\begin{tabular}{ln{1}{4}n{1}{4}}\hline%
 \bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
@@
 \begin{table}[h!]
-\begin{tabular}{l|c|c}\hline%
+\npdecimalsign{.}
+\nprounddigits{4}
+\begin{tabular}{ln{4}{4}n{4}{4}}\hline%
 \bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
@@
 the residuals both demonstrate a desirable lack of pattern.
-The plots examining the normality of the residuals do not follow a diagonal line as closely as would be hoped, but not an extreme departure.}
+The plots examining the normality of the residuals do not follow a diagonal line as closely as would be hoped, but are not an extreme departure.}
 \label{figure:resids_netincome}
@@
 \includegraphics[width = \textwidth]{resids_workers_nonzero.pdf}
-\caption{Diagnostic Residual plots for Nonzero workers.}
+\caption{Diagnostic Residual plots for Nonzero workers. The plots of predictions versus residuals shows points bunch at the left end, but the QQ plot closely follows a diagonal line.}
 \label{figure:resids_workers_nonzero}
@@
 \includegraphics[width = \textwidth]{resids_invasive_nonzero.pdf}
-\caption{Diagnostic Residual plots for Nonzero Invasive.}
+\caption{Diagnostic Residual plots for Nonzero Invasive.
+Due to only a few variables being included in these models and some of them being discrete, the predictions occur at a limited number of distinct points.}
 \label{figure:resids_invasive_nonzero}
@@
 
+\begin{table}[h!]
+\npdecimalsign{.}
+\nprounddigits{4}
+\begin{tabular}{ln{1}{4}n{1}{4}n{1}{4}}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Percperm} & \bfseries \underline{Perctemp} & \bfseries \underline{Perctill}%
+\csvreader[head to column names]{fullcoeflist_landuse_a.csv}{}%
+{\\\variable & \percperm & \perctemp & \perctill}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Landuse model}
+\label{table:landuse_full_a}
+\end{table}
+
+\begin{table}[h!]
+\npdecimalsign{.}
+\nprounddigits{4}
+\begin{tabular}{ln{1}{4}n{1}{4}}\hline%
+\bfseries \underline{Variable} & \bfseries \underline{Percpasture} & \bfseries \underline{Percbrush}%
+\csvreader[head to column names]{fullcoeflist_landuse_b.csv}{}%
+{\\\variable & \percpasture & \percbrush}%
+\\\hline
+\end{tabular}
+\caption{Full coefficient list for Landuse model (cont.)}
+\label{table:landuse_full_b}
+\end{table}
+
 \end{document}

### 2018-03-08 invasive_model.R

@@ -18,9 +18,10 @@ vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
 invasive_include <- vars[which(is.na(vars$`Surface area with invasive species`)),1]$`Variable Name`
 invasive_x.df <- subset(reduced_data, select = invasive_include)
 
-reduced_data$invasive <- 1*(reduced_data$percinv > 0)
+reduced_data$invasive <- factor(reduced_data$percinv > 0)
+levels(reduced_data$invasive) = c("Zero", "Positive")
 
-invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = factor(invasive)))
+invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = invasive))
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = percinv)) + 
   geom_histogram() + stat_bin(binwidth = 10)
 log_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = log10(percinv))) + 
@@ -34,9 +35,9 @@ invasive_binary.models <- fit.models(model.name = "invasive_binary", x.data = in
                                    y.response = reduced_data$invasive, response.family = "binomial")
 
 
-log_invasive_nonzero <- log10(reduced_data[reduced_data$percinv > 0, 'percinv'])
+log_invasive_nonzero <- log10(reduced_data[reduced_data$percinv > 0.1, 'percinv'])
 
-invasive.matrix <- model.matrix(~., invasive_x.df[reduced_data$invasive == 1,])[,-1]
+invasive.matrix <- model.matrix(~., invasive_x.df[reduced_data$percinv > 0.1,])[,-1]
 col_vars <- apply(invasive.matrix, 2, var)
 zero_variance <- names(col_vars[col_vars == 0])
 

### 2018-03-08 landuse_model.R

@@ -20,7 +20,7 @@ landuse_x.df <- subset(reduced_data, select = landuse_include)
 
 perc_vars <- c(landuse_x.df$percperm, landuse_x.df$perctemp)
 
-survey_vars <- c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush')
+survey_vars <- c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush',"percinv")
 reclass_vars <- c('ReclassAGRICOLA','ReclassCONSERVACION', 'ReclassFORESTAL', 'ReclassHABITACIONAL',
                   'ReclassPECUARIO', 'ReclassSIN APROVECHAMIENTO')
 
@@ -30,11 +30,20 @@ pdf("Paper/images/landuse_boxplots.pdf",width=12,height=8)
 ggplot(perc_long1, aes(x=factor(variable),y=value)) + geom_boxplot() + xlab("Land Use Category") + ylab("Percent")
 dev.off()
 
+use_data <- reduced_data[,survey_vars]
+use_data <- use_data[order(use_data[,7],use_data[,6],use_data[,5],use_data[,4],use_data[,3],use_data[,2],use_data[,1]),][,-7]
+use_data$rownum <- 1:nrow(use_data)
+use_data_melt <- melt(use_data, id="rownum")
+pdf("Paper/images/landuse_barplot.pdf",width=12,height=8)
+g <- ggplot(use_data_melt, aes(x = rownum, y= value))
+g + geom_bar(aes(fill=variable), stat="identity")+xlab("Farms")+ylab("Percent")
+dev.off()
+
 perc_long2 <- melt(landuse_x.df[,reclass_vars])
 ggplot(perc_long2, aes(x=factor(variable),y=value)) + geom_boxplot()
 
-primary_use <- factor(apply(landuse_x.df[,c('percperm','perctemp','perctill','percpasture','percbrush')], 1, which.max))
-primary_use <- factor(apply(landuse_x.df[,survey_vars], 1, which.max))
+primary_use <- factor(apply(reduced_data[,c('percperm','perctemp','perctill','percpasture','percbrush')], 1, which.max))
+#primary_use <- factor(apply(landuse_x.df[,survey_vars], 1, which.max))
 table(primary_use)
 
 perc_vars <- as.matrix(reduced_data[,survey_vars])

### 2018-03-08 modeling.function.R

@@ -2,10 +2,10 @@ fit.models <- function(model.name, x.data, y.response, response.family){
   
   # For testing
   
-  #model.name <- "workers_binary"
-  #x.data <- workers_x.df
-  #y.response <- reduced_data$workers_binary
-  #response.family <- "binomial"
+  #model.name <- "landuse"
+  #x.data <- landuse_x.df
+  #y.response <- primary_use
+  #response.family <- "multinomial"
   
   x.matrix <- model.matrix(~., x.data)[,-1]
   
@@ -24,7 +24,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
   } 
-  if (response.family == "mgaussian") {
+  if (response.family == "multinomial") {
     cv1=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=1)
     cv.75=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.75)
     cv.5=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=.5)
@@ -94,20 +94,97 @@ fit.models <- function(model.name, x.data, y.response, response.family){
   elastic5.signs <- lapply(elastic.coef5.nonzero, sign_check)
   
   ############ Figure out what to do with matrix of coefficients #############
-  if (response.family == "multinomial"){
-    elastic.coef5 <- coef(elastic.full,s=lambda.5var[1])
-    elastic.coef5.nonzero1 <- as.data.frame(as.matrix(elastic.coef5$`1`))
-    elastic5.names1 <- names(elastic.coef5.nonzero1[elastic.coef5.nonzero != 0,])
-    elastic.coef5.nonzero2 <- as.data.frame(as.matrix(elastic.coef5$`2`))
[…]

### 2018-03-27 Paper/paper_draft.tex
Added readme

@@
 \usepackage{numprint}
+\usepackage{wrapfig}
 
@@
   \\
-  March 8th, 2018 \\
+  March 27th, 2018 \\
 }
@@
   Variable selection is an important first step when analyzing datasets with a large number of potential predictor variables. We 
-  apply two techniques, Forward Selection and Elasticnet,  to find the most important factors in a dataset detailing over 200 
+  apply two techniques, Forward Selection and Elastic Net,  to find the most important variables in a dataset detailing over 200 
   socioeconomic measurements for 755 farms on the Gal\'apagos Islands. Modeling five different outcome variables, we find 
-  the data available has the strongest linear relationships with Productivity and Land Use Choices. For each of the outcome 
-  variables we present the top five predictor variables as well as a full set of coefficients for the optimally predictive model.
+  the data available has the strongest linear relationships with the outcomes \textit{productivity} and \textit{land use choices}. For each of 
+  the outcome variables we present the top five predictor variables as well as a full set of coefficients for the optimally predictive 
+  model.
   \end{abstract}
@@
 The Gal\'apagos Islands make for a feasible and significant case study of complex systems.
-Due to its relative isolation and smaller size the interaction of factors can more realistically be modeled for the Gal\'apagos Islands than other systems.
-Yet the Gal\'apagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment.
-Prior work has created agent-based models of the Gal\'apagos, but with limited interaction parameters between agents, particularly in regards to farm success (\cite{miller}, \cite{valdivia},\cite{walsh}).
+Due to their relative isolation and smaller size, the interaction of factors can more realistically be modeled for the Gal\'apagos Islands than other systems.
+Yet, the Gal\'apagos Islands also represents an important example of the competing forces of resource conservation and economic development in a rapidly changing environment.
+Prior work has created agent-based simulation models of the Gal\'apagos, but with limited interaction parameters between agents, particularly in regards to farm success (\cite{miller}, \cite{valdivia},\cite{walsh}).
 In order to create a more detailed and perhaps more accurate simulation, the relationships between different factors on the island must be better understood.
@@
 The data available to study the dynamics between agricultural measures and related factors primarily 
-comes from the Censo de las Unidades de Producci\'on Agropecuaria (upa) de Gal\'apagos (Census of Agricultural Production 
+comes from the Censo de las Unidades de Producci\'on Agropecuaria (UPA) de Gal\'apagos (Census of Agricultural Production 
 Units (UPA) of Galapagos) (\cite{census}). This is a self-reported survey with data from 755 farms (UPAs) detailing many characteristics. The questions covered fall into the following categories:
@@
 In addition to this census, data is also available from satellite image classification including 
-information on water, energy and road access. In total, under the direction of the client 239 variables were selected for 
+information on water, energy and road access. In total, under the direction of the client, 239 variables were selected for 
 consideration in modeling relationships between potential predictors and five outcome variables of interest.
@@
 combination of multiple variables. When analyzing a derived outcome variable all variables used in its original calculation were 
-removed from consideration in the model . The client also denoted specific predictor variables to exclude from particular 
+removed from consideration in the model. The client also denoted specific predictor variables to exclude from particular 
 models when their inclusion would not be beneficial. For example, while the 
-amount of crops sold in pounds was not directly used to calculate net income, the obvious relationship existent precluded it 
-from inclusion. In addition, predictor variables that met one or more of the following criteria were removed prior to modeling: 
-zero variance, extremely high (>0.99) or perfect correlation with other predictor variables, or linear dependence with other 
-predictor variables (that is, two or more predictor variables could be linearly combined to create another predictor variable). The 
+amount of crops sold in pounds was not directly used to calculate \textit{net income}, the obvious relationship existent 
+precluded it from inclusion. In addition, predictor variables that met one or more of the following criteria were removed prior to 
+modeling: zero variance, extremely high (>0.99) or perfect correlation with other predictor variables, or linear dependence with 
+other predictor variables (that is, two or more predictor variables could be linearly combined to create another predictor variable). The 
 exact number of predictor variables included in each model varied slightly, but there were approximately 200 predictors variables examined for each model after the preceding steps were taken.
@@
 
-The remainder of this report is divided into four sections, Section 2: Modeling, Section 3: Results, Section 4: Statistical 
-Methods, and Section 5: Limitations and Future Work.
+The remainder of this report is divided into four sections.
 Section 2 provides a brief overview of the analysis so that the results can be understood.
@@
 Section 5 details important considerations when interpreting the results and suggests possible avenues for future work.
-Lastly, References and the Appendix, including additional tables and figures, can be found at the end.
+Lastly, References and the Appendix, including additional tables and figures, can be found following Section 5.
 
@@
 Several outcome variables in this study show strong non-normality, which can contribute to a poorly fitting model and unreliable estimates of the coefficients.
-In order to address this issue transformations to the data and modifications to the standard linear model will be considered.
+In order to address this issue transformations to the data and modifications to the standard linear model were considered.
 
@@
 of predictors was built. Each relationship was modeled 
-using Forward Selection and ElasticNet regression. Forward Selection fits a linear model by progressively adding variables to 
-the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner (computed using the R package `leaps' \cite{lumley}). 
-ElasticNet regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
+using Forward Selection and Elastic Net regression. Forward Selection fits a linear model by progressively adding variables to 
+the model until a best fit is found. This results in only some of the variables being included, chosen in a discrete manner 
+(computed using the R package `leaps' \cite{lumley}). 
+Elastic Net regression fits a linear model by limiting the size of the coefficients so that they are smaller than in standard least 
 squares, and for many variables actually shrunken to zero. Similar to Forward Selection this results in a smaller model, but 
@@
 In general, these techniques have slightly different aims. Forward Selection chooses a model that best explains the variance in 
-the dataset at hand. Elasticnet chooses a model that can best make predictions on new data. Depending on the goals of 
+the dataset at hand. Elastic Net chooses a model that can best make predictions on new data. Depending on the goals of 
 analysis, one technique is not necessarily better so we do not compare the two quantitatively, but instead offer both results as 
@@
 
+The Results section is broken into three subsections covering the outcome categories of interest: farm success, invasive species and land use choices.
+Within farm success we further divide into three specific measurements, resulting in a total of 5 different outcome variables.
+We explore each outcome variable in turn, presenting a visualization of the data, the top five predictor variables and analysis of optimal linear models.
+
 \subsection{Farm Success}
 
-The first three outcome variables of interest can be grouped together under the category of farm success; labeled as 
-productivity, net income and number of workers supported. Productivity is calculated as the total pounds of crops and livestock 
-produced divided by the farm surface area. Net income is calculated as the the revenue from all products 
-sold minus total expenses. Number of workers supported is calculated as the total labor expenditures divided by a standard 
+The first three outcome variables of interest can be grouped together under the category of farm success:
+\begin{itemize}
+	\item \textit{Productivity}: total pounds of crops and livestock produced divided by the farm surface area. 	
+	\item \textit{Net income}: revenue from all products sold minus total expenses.
+	\item \textit{Number of workers supported}: total labor expenditures divided by a standard 
 	full-time worker's salary.
+\end{itemize}
 
@@
 \includegraphics[width = 0.8\textwidth]{production_histograms.pdf}
-\caption{Histogram of Productivity, from left to right showing the full range, between 0-10,000 lbs/hectare, and the log transformed nonzero values.
-The log transformed values exhibit the desired normal shape.}
+\setlength{\belowcaptionskip}{-10pt}
+\caption{\textsl{\small Histogram of Productivity, from left to right showing the full range, zoomed into 0-10,000 lbs/hectare, and the log transformed nonzero values.
+The log transformed values exhibit the desired normal shape.}}
 \label{figure:prod_hist}
@@
 
-The histograms of the Productivity variable (Figure \ref{figure:prod_hist}) show a strong skewness, both when looking at all 
-observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
+The histograms of the \textit{Productivity} variable (Figure \ref{figure:prod_hist}) show a strong skewness, both when looking at 
+all observations, and when zooming into observations between 0-10,000 lbs/hectare. To achieve a distribution closer to normal 
 (bell-curved), which will benefit the linear model, we took the $log_{10}$ transformation with resulting data shown in the last 
-plot. Since the log transformation cannot be performed on zero values, we removes the 38 occasions of this from the dataset. 
+plot. Since the log transformation cannot be performed on zero values, we removed the 38 occasions of this from the dataset. 
 Beyond the mathematical constraint, farms with zero production perhaps are not farms as typically defined.
 
-\begin{table}[h!]
+We built linear models using both methods, Elastic Net and Forward Selection, on the log-transformed 
+\textit{Productivity} variable, recording an optimal model of any size and the best 5 variable model for each.
+The size of 5 variables is chosen to provide quick insight and not based on any specific statistical property.
+The results of the best 5 variable model are shown
+in Table \ref{table:prod_top5}, listed in order of entrance into the model.  Next to variable names the direction of the relationship is indicated 
+with a (+) or (-).
+
+\begin{table}[h]
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection} 
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection} 
 \csvreader[head to column names]{production_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Modeling of Productivity, Top 5 features for both methods}
+\caption{\textsl{\small Modeling of Productivity, Top 5 features for both methods}}
 \label{table:prod_top5}
@@
 
-We built linear models using both methods, Elasticnet and Forward Selection, on the log-transformed 
-productivity variable, recording an optimal model of any size and the best 5 variable model for each. The size of 5 variables is 
-chosen to provide quick insight and not for any specific statistical property. The results of the best 5 variable model are shown 
-in Table \ref{table:prod_top5}, listed in alphabetical order.  Next to variable names the direction of the relationship is indicated 
-with a (+) or (-). There is some overlap between the variables selected by each of the methods, but also unique choices made 
-by each method. The Root Mean-Squared Error (RMSE) for the 5 variable Elasticnet model is 0.78 and the $R^2$ for 
-Forward Selection is 0.48. The RMSE is on average how far the predicted value is from the actual value and $R^2$ is the 
-percentage of variability in the outcome variable that is explained by the model. To provide context for the RMSE, we can look 
-at the point furthest to the right on the cross-validation plot and see how the model would perform when no variables are 
-included in the model, that is just predicting the average outcome value.
-
-Plots from the optimal models for Elasticnet and Forward Selection are shown in Figure \ref{figure:prod_opt}. The cross-
-validation plot for Elasticnet can be understood as follows: the 
+Plots from the optimal models for Elastic Net and Forward Selection are shown in Figures \ref{figure:prod_opt_e} and  \ref{figure:prod_opt_f}. The cross-
+validation plot for Elastic Net can be understood as follows: the 
 horizontal axis shows the number of variables included in the model (on top) and the corresponding lambda ($\lambda$) value 
@@
 fewer variables. Since our goal is to select a small amount of variables, we will generally use the model found at $
-\lambda_{1se}$. For Forward Selection the plot is much more straightforward, we plot the number of variables included versus 
-the information criterion that we choose to minimize and mark the optimal point in red.
+\lambda_{1se}$.
 
-\begin{figure}[h]
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{elastic_cv_production.pdf}
-\caption{Cross-validation plot for Elasticnet}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{forward_nvars_production.pdf}
-\caption{Forward Selection plot}
-\end{subfigure}
-\caption{Productivity Variable Selection.
-The Elasticnet plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error.
-The optimal Forward Selection model includes 26 variables.}
-\label{figure:prod_opt}
-\end{figure}
+\begin{wrapfigure}{l}{0.6\textwidth}
+\includegraphics[width=0.6\textwidth]{elastic_cv_production.pdf}
+\setlength{\belowcaptionskip}{-10pt}
+\caption{\textsl{\small Elastic Net Cross-validation plot for Productivity.
+The plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error.}}
+\label{figure:prod_opt_e}
+\end{wrapfigure}
 
-The convex shape of the plots highlights a common trend in variable selection; not including enough variables does not provide 
-enough information, but beyond a certain point adding more variables may not be worth the added complexity. The optimal 
-model chosen for Elasticnet includes 45 variables (not counting the intercept term) and using Forward Selection we choose a 
-model of 26 variables, almost all of which are also included in the Elasticnet model. The coefficients estimated for both models 
-can be found in Table \ref{table:prod_full} in the Appendix. For the full models, the RMSE and $R^2$ are 0.61 and 0.63, 
-respectively. These numbers suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
-gained by adding more variables.
-Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.
+For Elastic Net we measure performance with Root Mean-Squared Error (RMSE), which is on average how far the predicted 
+value is from the actual value across cross-validation runs. The 5 variable Elastic Net model had a RMSE of 0.78. Not counting 
+the intercept term, the optimal model chosen for Elastic Net included 45 variables and had a RMSE of 0.61. To provide context 
+for the RMSE, we can look at the point furthest to the right on the cross-validation plot in Figure \ref{figure:prod_opt_e} and see 
+how the model would perform when no variables are included in the model, that is, just predicting the average outcome value.
+
+For Forward Selection the plot, Figure \ref{figure:prod_opt_f}, is much more straightforward. We plot the number of variables 
+included versus the Bayesian Information Criterion (BIC) that we choose to minimize, and mark the optimal point in red. The 
+convex shape of the plots highlights a common trend in variable selection; not including enough variables does not provide 
+enough information, but beyond a certain point adding more variables may not be worth the added complexity.
+
+\begin{wrapfigure}{r}{0.55\textwidth}
+\vspace*{-0.5cm}
+\includegraphics[width=0.55\textwidth]{forward_nvars_production.pdf}
+\setlength{\belowcaptionskip}{-20pt}
+\caption{\textsl{\small Forward Selection plot for Productivity. The optimal model includes 26 variables.}}
+\label{figure:prod_opt_f}
+\end{wrapfigure}
+
+We evaluate Forward Selection with $R^2$, which is the percentage of variability in the outcome variable that is explained by 
+the model. The 5 variable Forward Selection model had an $R^2$ of 0.48 and the full model included 26 variables and had a 
+$R^2$ of 0.63.
+
+The coefficients estimated for both full models can be found in Table \ref{table:prod_full} in the Appendix. These numbers 
+suggest that while the 5 variable model is helpful, there is a decent amount of information to be 
+gained by adding more variables. Diagnostics of the linear fit of the optimal Elastic Net and Forward Selection models (plots 
+shown in Figure \ref{figure:resids_prod} in the Appendix) do not raise any concerns.
 
 \subsubsection{Net Income}
+
+The histograms for \textit{Net Income} (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and 
+a few observations wide in the tails.
+We attempted to fit a model on the full dataset, but find the observations with large absolute values are obscuring other possible information in the model.
+After trying several cutoff thresholds and examining model fit, we  removed the 28 observations that are outside the middle 95{\%} of the data.
+The
+following models are fit on this reduced dataset of 727 observations.
 
@@
 \includegraphics[width = 0.75\textwidth]{netincome_histograms.pdf}
-\caption{Histograms of Net Income, with full data and data within one standard deviation displayed. The values very far from the center are of concern when modeling.}
+\caption{\textsl{\small Histograms of Net Income, with full data and the central 95{\%} of data displayed. The values very far from the center are of concern when modeling.}}
 \label{figure:netincome_hist}
 \end{figure}
-
-The histograms for Net Income (Figure \ref{figure:netincome_hist}) show a symmetric shape, but a very spiky center and a few 
-observations wide in the tails. We attempted to fit a model on the full dataset, but find the observations with large absolute 
-values are obscuring other possible information in the model.
-After trying several cutoff thresholds and examining model fit, we  removed the 28 observations that are beyond one standard deviation (30,478) from the mean in either direction.
-The
-following models are fit on this reduced dataset of 727 observations.
 
@@
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{netincome_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Modeling of Net Income, Top 5 features for both methods}
+\caption{\textsl{\small Modeling of Net Income, Top 5 features for both methods}}
 \label{table:netincome_top5}
@@
 
-The results of the best 5 variable model are shown in Table \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elasticnet and a $R^2$ of 0.15 for Forward Selection.
-For interpretation of the RMSE it is important to keep in mind the scale for Net Income is much larger than that of log productivity.
-However, in this case adding variables has caused a minimal decrease in the error.
-\begin{figure}[!tbp]
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{elastic_cv_netincome.pdf}
-\caption{Cross-validation plot for Elasticnet...}
-\end{subfigure}
-\hfill
-\begin{subfigure}[b]{0.5\textwidth}
-\includegraphics[width=\textwidth]{forward_nvars_netincome.pdf}
-\caption{Forward Selection plot}
-\end{subfigure}
-\caption{Net Income Variable Selection. The Elasticnet plot shows a much smaller improvement in prediction error and Forward Selection includes a small number of variables.}
-\label{figure:netincome_opt}
+The results of the best 5 variable model are shown in Table \ref{table:netincome_top5}, with a RMSE of 5790.35 for Elastic Net. 
+For interpretation of the RMSE it is important to keep in mind the scale for \textit{Net Income} is much larger than that of log 
+productivity. However, in this case adding variables has caused a minimal decrease in the error. The cross-validation plot for 
+\textit{Net Income} show wider error bars throughout the range of model sizes and just using the average 
+\textit{Net Income} would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
+chose the optimal Elastic Net model to be at $\lambda_{min}$, which included 9 variables and had a RMSE of 5768.30.
+
+\begin{wrapfigure}{l}{0.6\textwidth}
+\vspace*{-0.5cm}
+\includegraphics[width=0.6\textwidth]{elastic_cv_netincome.pdf}
+\setlength{\belowcaptionskip}{-10pt}
+\caption{\textsl{\small Elastic Net Cross-validation plot for Net Income. The plot shows a much smaller improvement in prediction error.}}
+\label{figure:netincome_opt_e}
+\end{wrapfigure}
+
+For Forward Selection the 5 variable model and full model had $R^2$ values of 0.15 and 0.19, respectively. The full Forward 
+Selection model included 9 variables, with all but two overlapping with the Elastic Net choices. The coefficients estimated for 
+both models can be found in Table \ref{table:netincome_full} in the Appendix.
+
+Since the full models are not much larger than the 5 variable models, the small improvements are not surprising. Diagnostics of 
+the linear fit of the optimal Elastic Net and Forward Selection models (plots shown in Figure 
+\ref{figure:resids_netincome} in the Appendix) do not follow assumptions as closely as for the Production models, but are not so 
+concerning as to disqualify the results. On the whole, the results from the various plots and diagnostics suggest that the 
+relationships found for Net Income are worth investigating, but that a linear relationship provides little predictive power.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.55\textwidth]{forward_nvars_netincome.pdf}
+\setlength{\belowcaptionskip}{-10pt}
+\caption{\textsl{\small Forward Selection plot for Net Income, which includes a small number of variables.}}
+\label{figure:netincome_opt_f}
 \end{figure}
-
-The cross-validation plot for Net Income show wider error bars throughout the range of model sizes and just using the average 
-Net Income would predict nearly as well as any other model. Since here $\lambda_{1se}$ only includes the intercept, we 
-chose the optimal Elasticnet model to be at $\lambda_{min}$, which includes 9 variables. For Forward Selection, we also 
-included 9 variables, with all but two overlapping with the Elasticnet choices. The coefficents estimated for both models can be 
-found in Table \ref{table:netincome_full} in the Appendix. The full models had an RMSE and $R^2$ of 5768.30 and 0.19, 
-respectively.
-Since the full models are not much larger than the 5 variable models, the small improvements are not surprising.
-Diagnostics of the linear fit of the optimal Elasticnet and Forward Selection models (plots shown in Figure
-\ref{figure:resids_netincome} in the Appendix) do not follow assumptions as closely as for the Production models, but are not so 
-concerning as to disqualify the results.
-On the whole, the results from the various plots and diagnostics suggest that the relationships found for Net Income are worth investigating, but that a linear relationship is not very strong.
 
@@
 \includegraphics[width = 0.75\textwidth]{worker_histograms.pdf}
-\caption{Histograms of Workers, showing the proportion of zeroes, farms with nonzero workers, and the log transformed number of workers.
-The log transformation helps remove the skewness, but some bi-modality still remains.}
+\caption{\textsl{\small Histograms of Workers, showing the proportion of zeroes, farms with nonzero workers, and the log transformed number of workers.
+The log transformation helps remove the skewness, but some bi-modality still remains.}}
 \label{figure:workers_hist}
@@
 
-The first plot in Figure \ref{figure:workers_hist} shows that a large percentage of the farms are not able to support any workers. 
-For this reason, we divided up the modeling task for this outcome variable. First, we used logistic regression to model the 
-binary variable of whether a farm supports zero or more than zero workers. Secondly, we used linear regression to model the 
-quantity of workers for just those 264 farms with a positive number of workers supported.
+The first plot in Figure \ref{figure:workers_hist} shows that a large percentage (65{\%}) of the farms are not able to support any 
+workers. For this reason, we divided up the modeling task for \textit{Number of Workers Supported}. First, we used logistic regression to model 
+the binary variable of whether a farm supports zero or more than zero workers. Secondly, we used linear regression to model 
+the quantity of workers for just those 264 farms with a positive number of workers supported.
 
-\begin{table}[h]
+\begin{table}[h!]
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{workers_binary_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Modeling of Binary Workers , Top 5 features for each method}
+\caption{\textsl{\small Modeling of Binary Workers , Top features for both methods}}
 \label{table:workers_binary_top5}
@@
 
-\begin{table}[h]
+\begin{table}[h!]
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Nonzero Workers}
-\caption{Modeling of Nonzero Workers, Top 5 features for both methods}
+\caption{\textsl{\small Modeling of Nonzero Workers, Top 5 features for both methods}}
 \label{table:workers_nonzero_top5}
@@
 
-Sometimes Elasticnet will simultaneously choose to eliminate multiple variables at the same time, in this case there is not a 5 
-variable model so we show the 6 variable model for Elasticnet. For logistic regression we can measure performance with the 
-misclassification rate, which on average is 0.31 on the test set for Elasticnet and 0.26 on the full dataset for Forward Selection. 
-These can be compared to the 0.35 misclassification rate we would achieve if we simply predicted the majority class, zero 
-workers, every time.
-For the linear model 5 variable model the RMSE is 0.46 and the $R^2$  0.30.
-The low misclassification rate coupled with low $R^2$ show that there is a fairly clear divide between farms that can or cannot support workers, but the specific number of workers is much harder to predict.
-Several variables are found to be most helpful for both the logistic regression and linear regression models, but there is still a fair bit of difference.
+The top 5 variable models for logistic regression and linear regression can be found in Table \ref{table:workers_binary_top5} 
+and Table \ref{table:workers_nonzero_top5}, respectively. 
+Sometimes Elastic Net will simultaneously choose to eliminate multiple variables, in this case there is not a 5 
+variable model so we show the 6 variable model for Elastic Net. For logistic regression we can measure performance with the 
+misclassification rate, which on average was 0.31 on the test set for Elastic Net, compared to the 0.35 misclassification rate we 
+would achieve if we simply predicted the majority class, zero workers, every time.  The full Elastic Net logistic regression model 
+had 6 variables and the same 0.31 misclassification rate. For Elastic Net linear regression, the 5 variable model resulted in a 
+RMSE of 0.46 and the full model also included 5 variables, and therefore the same RMSE.
 
@@
 \includegraphics[width=\textwidth]{elastic_cv_workers_binary.pdf}
-\caption{Cross-validation plot for Elasticnet}
+\caption{\textsl{\small Cross-validation plot for Elastic Net}}
 \end{subfigure}
@@
 \end{subfigure}
-\caption{Workers Binary Variable Selection.
-The Elasticnet plot shows ideal predictive performance when including 5-25 variables.
-The computation for logistic regression Forward Selection stops at the optimal value so the rest of the plot is not included as before.}
+\caption{\textsl{\small Workers Binary Variable Selection.
+The Elastic Net plot shows ideal predictive performance when including 5-25 variables.
+The computation for logistic regression Forward Selection stops at the optimal value so the rest of the plot is not included as before.}}
 \label{figure:workers_binary_opt}
 \end{figure}
+
+For logistic regression Forward Selection the 5 variable model had a misclassification rate of 0.26, which is measured on the 
+original dataset, not a test set. The optimal Forward Selection logistic regression chose a model of size 7 also with a 
+misclassification rate of 0.26. For linear regression, the 5 variable Forward Selection model had a $R^2$ of 0.34 and the 
+optimal model had a $R^2$ of 0.34 using 8 variables.
+
+In the Appendix, the coefficients estimated for logistic regression models can be found in Table \ref{table:workers_binary_full} 
+and the coefficients estimated for linear regression models can be found in Table \ref{table:workers_nonzero_full}. The fairly 
+low misclassification rate coupled with low $R^2$ show that there is a fairly clear divide between farms that can or cannot 
+support workers, but the specific number of workers is much harder to predict. Several variables are found to be most helpful 
+for both the logistic regression and linear regression models, but there is still a fair bit of difference. One common theme of the 
+variables chosen here is that many are related to cattle production.
 
@@
 \includegraphics[width=\textwidth]{elastic_cv_workers_nonzero.pdf}
-\caption{Cross-validation plot for Elasticnet}
+\caption{\textsl{\small Cross-validation plot for Elastic Net}}
 \end{subfigure}
@@
 \end{subfigure}
-\caption{Nonzero Workers Variable Selection. Both plots show that adding variables does little to improve performance, highlighting a lack of linear relationship.}
+\caption{\textsl{\small Nonzero Workers Variable Selection. Both plots show that adding variables does little to improve performance, highlighting a lack of linear relationship.}}
 \label{figure:workers_nonzero_opt}
@@
 
-For logistic regression Elasticnet chose a model of size 6 and Forward Selection chose a model of size 7. Since these 
-models are not much larger the misclassification rates remained at 0.31 and 0.26, respectively. For linear regression, the full 
-models for Elasticnet and Forward Selection had a RMSE and $R^2$ of 0.46 and 0.34, respectively. The Elasticnet full model 
-had 5 variables and the Forward Selection full model had 8 variables.  In the Appendix, the coefficients estimated for logistic 
-regression models can be found in Table \ref{table:workers_binary_full} and the coefficients estimated for linear regression 
-models can be found in Table \ref{table:workers_nonzero_full}.
-
 \subsection{Invasive Species}
 
-The analysis of Invasive Species follows much the same pattern as for Number of Workers Supported. Again, we have a large 
+The analysis of \textit{Invasive Species} follows much the same pattern as for \textit{Number of Workers Supported}. Again, we have a large 
 number of zeros in the outcome variable, farms with zero percent of their land covered by invasive species. We first modeled 
@@
 
-\begin{figure}[h!]
+\begin{figure}[h]
 \centering
@@
 
-\begin{table}[h!]
+\begin{table}[h]
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{invasive_binary_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Modeling of Binary Invasive , Top 5 features for each method}
+\caption{\textsl{\small Modeling of Binary Invasive , Top features for each method}}
 \label{table:invasive_binary_top5}
@@
 
-\begin{table}[h!]
+\begin{table}[h]
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
-\bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{invasive_nonzero_top5names.csv}{}%
@@
 \end{tabular}
-\caption{Modeling of Nonzero Invasive , Top 5 features for each method}
+\caption{\textsl{\small Modeling of Nonzero Invasive , Top 5 features for each method}}
 \label{table:invasive_nonzero_top5}
@@
 
+\pagebreak
+
 The top 5 variable models for logistic regression and linear regression can be found in Table \ref{table:invasive_binary_top5} 
-and Table \ref{table:invasive_nonzero_top5}, respectively. For logistic regression the misclassification rate was 0.21 on the test 
-set for Elasticnet and 0.20 on the full dataset for Forward Selection. For the linear model 5 variable model the RMSE was 0.25 and the $R^2$ was 0.30.
+and Table \ref{table:invasive_nonzero_top5}, respectively. For Elastic Net logistic regression the misclassification rate was 0.21 
+on the test set for the 5 variable model. For the Elastic Net linear regression 5 variable model the RMSE was 0.25 and the optimal model only included 4 variables with the same RMSE.
 
@@
 \includegraphics[width=\textwidth]{elastic_cv_invasive_binary.pdf}
-\caption{Cross-validation plot for Elasticnet}
+\caption{\textsl{\small  Cross-validation plot for Elastic Net}}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_invasive_binary.pdf}
-\caption{Forward Selection plots}
+\caption{\textsl{\small Forward Selection plot}}
 \end{subfigure}
-\caption{Invasive Binary Variable Selection. The Elasticnet plot shows adding variables does not improve just predicting the majority class. In Forward Selection the BIC very quickly levels off.}
+\caption{\textsl{\small Invasive Binary Variable Selection.
+The Elastic Net plot shows adding variables does not improve just predicting the majority class.
+In Forward Selection the BIC very quickly levels off.}}
 \label{figure:invasive_binary_opt}
@@
 \includegraphics[width=\textwidth]{elastic_cv_invasive_nonzero.pdf}
-\caption{Cross-validation plot for Elasticnet}
+\caption{\textsl{\small Cross-validation plot for Elastic Net}}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{forward_nvars_invasive_nonzero.pdf}
-\caption{Forward Selection plots}
+\caption{\textsl{\small Forward Selection plot}}
 \end{subfigure}
-\caption{Nonzero Invasive Variable Selection. Similar to Figure \ref{figure:invasive_binary_opt} the lack of benefit from adding additional variables demonstrates a weak linear relationship.}
+\caption{\textsl{\small Nonzero Invasive Variable Selection.
+Similar to Figure \ref{figure:invasive_binary_opt} the lack of benefit from adding additional variables demonstrates a weak linear relationship.}}
 \label{figure:invasive_nonzero_opt}
@@
 
+For Forward Selection the 5 variable logistic regression model had a misclassification rate of 0.20 on the full dataset.
+The 5 variable linear regression model had an $R^2$ of 0.30 and the optimal model included 7 variables to increase the $R^2$ 0.35.
+
 The cross-validation and forward selection plots in Figure \ref{figure:invasive_binary_opt} show that there was little gained from 
-adding variables to the logistic regression model. The optimal Elasticnet model only included the intercept and the optimal 
+adding variables to the logistic regression model. The optimal Elastic Net model only included the intercept and the optimal 
 Forward Selection model still only included 5 variables. For this reason the misclassification rates are not changed from above.
 
-For linear regression the optimal models for Elasticnet included 4 variables with a RMSE still of 0.25 and for Forward Selection 
-included 7 variables with a $R^2$ slightly increasing to 0.35.  Since the percentage of farms with no invasive species is 20{\%} 
-the logistic regression was not able to do much more than just predict zero for all farms. This combined with the low 
-performance of the linear model suggests that invasive species coverage cannot be well explained by the variables considered 
-here.
+Since the percentage of farms with no invasive species is 20{\%} the logistic regression was not able to do much more than just 
+predict zero for all farms.
+This combined with the low performance of the linear model suggests that invasive species coverage cannot be well explained by the variables considered here.
+That said, more than seen in previous outcome variables, the canton variables seem to be particularly important for \textit{invasive species}.
+Similar variables were chosen for Elastic Net and Forward Selection in logistic regression, but not for linear regression.
 
@@
 
-The analysis of Land Use choice requires a slightly different approach than used previously since the outcome variable is a 
+The analysis of land use choice requires a slightly different approach than used previously since the outcome variable is a 
 categorical variable, with six different classifications of land use (each measured as a percentage of total land area):
@@
 
-While not a perfect ordering, the classifications can generally thought of progressing from land that has been most worked by 
-the farmer to least worked. A primary concern is that unworked land may be most susceptible to invasive species growth. For 
-this reason we sort the farms in ascending order first by percent invasive species, and then by percent brush, percent pasture, 
-etc. and graph the stacked bar plots in Figure \ref{figure:landuse_plot}. Since percent brush is given precedence in sorting the 
-purple area stands out more than the rest. Even so, it can be seen that there are farms primarily made up of each of the 
-categories as well as farms fairly evenly split between the categories.
+While not a perfect ordering, the classifications can generally be thought of progressing from land that has been most worked 
+by the farmer to least worked. We visualized the land use data using a parallel coordinate plot (Figure 
+\ref{figure:landuse_plot}) where each line represents a farm and the height represents the percent of land used for that 
+category. Darker or thicker line areas indicate more farms falling along that path. We can see that there are farms primarily 
+used for each of the different land use categories, but many more so for pasture, permanent crops and brush. If there were 
+many horizontal, or near horizontal, lines on the graph that would indicate even distribution of land across multiple categories, 
+but there is not much evidence of that in this plot.
 
@@
 \centering
-\includegraphics[width = 0.75\textwidth]{landuse_barplot.pdf}
-\caption{Stacked Barplot of Land Use.
-Farms are sorted in ascending order by percent invasive and then by percbrush, percpasture, etc.
-Some bars do not total 100\% because percent invasive and percent other categories are not included.
-There appear to be different classes of farms, some primarily permanent crops, some primarily temporary crops, etc.
-}
+\includegraphics[width = 0.95\textwidth]{landuse_parplot.pdf}
+\caption{\textsl{\small Parallel Coordinate Plot of Land Use.
+There appear to be different types of farms, some primarily permanent crops, some primarily temporary crops, etc.
+The land use categories pasture, permanent crops and brush appear to be the most prevalent.}}
 \label{figure:landuse_plot}
@@
 There are multiple ways that this outcome variable can be formulated for modeling, but we decide to label each farm with its 
-highest percentage land use category. There are only 9 such farms labeled as fallow, which is too small for modeling so we 
+highest percentage land use category. There are only 9 such farms labeled as \textit{fallow}, which is too small for modeling so we 
 remove them from the consideration. The remaining categories and the number of farms are shown in Table \ref{table:landuse_cat}.
@@
 \hline
-Percperm    & 209             \\
-Perctemp    & 50              \\
-Perctill    & 36              \\
-Percpasture & 271             \\
-Percbrush   & 189            
+Perm    & 209             \\
+Temp    & 50              \\
+Till    & 36              \\
+Pasture & 271             \\
+Brush   & 189            
 \end{tabular}
-\caption{Landuse Category Count}
+\caption{\textsl{\small Landuse Category Count}}
 \label{table:landuse_cat}
@@
 
-Using this derived categorical variable we used multinomial logistic regression with Elasticnet to build a model, which found an 
-optimal set of coefficients for each category.
-The results from the 5 variable model, which had a misclassification rate of 0.39, are shown in Table \ref{table:landuse_top5}.
-Different categories may have varying number of coefficients and because Perctill had no variables included at this lambda level, it is left out of the table.
+Using this derived categorical variable we used multinomial logistic regression with Elastic Net to build a model, which found an 
+optimal set of coefficients for each category. The results from the 5 variable model, which had a misclassification rate of 0.39, 
+are shown in Table \ref{table:landuse_top5}. Different categories may have varying number of coefficients and \textit{Perctill} 
+had no variables included at this lambda level. As might be expected, presence of permanent crops predicted \textit{Percperm}, 
+presence of transitory crops predicted \textit{Perctemp}, etc., but the specific crops found most predictive may be of interest.
 
@@
 \centering
+\setlength{\tabcolsep}{20pt}
 \begin{tabular}{cc}
@@
 \end{tabular}
-\begin{tabular}{cc}
-\bfseries \underline{Percpasture} & \bfseries \underline{Percbrush}%
+\begin{tabular}{ccc}
+\bfseries \underline{Percpasture} & \bfseries \underline{Percbrush} & \bfseries \underline{Perctill}%
 \csvreader[head to column names]{landuse_top5names.csv}{}%
-{\\\percpasture & \percbrush}%
+{\\\percpasture & \percbrush & \perctill}%
 \end{tabular}
-
-\caption{Modeling of Landuse , Top features for each category}
+\caption{\textsl{\small Modeling of Landuse, Top features for each category}}
 \label{table:landuse_top5}
@@
 
-The plot of the Elasticnet cross-validation is shown in Figure \ref{figure:landuse_opt}.
-The optimal model achieves a 0.36 misclassification rate against the test set on average, which given 5 categories to choose from shows decent predictive strength.
+\pagebreak
+
+The plot of the Elastic Net cross-validation is shown in Figure \ref{figure:landuse_opt}.
+The optimal model achieved a 0.36 misclassification rate against the test set on average, which given 5 categories to choose from shows decent predictive strength.
 In the Appendix, the coefficients estimated for the optimal model can be found in Tables \ref{table:landuse_full_a} and
@@
 \includegraphics[width=\textwidth]{elastic_cv_landuse.pdf}
-\caption{Land Use Cross-validation plot for Elasticnet.
-The Elasticnet model shows a large improvement over just predicting the majority class.
-Since there are now 5 categories, such a strategy would result in a 0.65 misclassification rate.}
+\caption{\textsl{\small Land Use Cross-validation plot for Elastic Net.
+The Elastic Net model shows a large improvement over just predicting the majority class.
+Since there are now 5 categories, such a strategy would result in a 0.65 misclassification rate.}}
 \hfill
@@
 \[\bY = \bX\beta + \epsilon \tag{1}\]
-%\[Y_i =  \sum_{j=1}^{p} X_{i,j}\beta_j + \epsilon \tag{1}\]
 
@@
 
-This calls for the use of logistic regression, in which we perform a logit transformation as seen in equation 2 below so that the 
+This calls for the use of logistic regression, a type of generalized linear model \cite{dobson}, in which we perform a logit transformation as seen in equation 2 below so that the 
 $\bX\beta$ can still be mapped to a continuous scale. In some respects this is a computational concern, but it also changes the 
@@
 
-Equation can be rewritten as below in equation 3, which gives the predicted probability of an observation being class 1.
+Equation 2 can be rewritten as below in equation 3, which gives the predicted probability of an observation being class 1.
 
@@
 on generalizable patterns or just random noise. Using cross-validation the average test set mean square error is an 
-appropriate measure of model fit. We can also capture the standard deviation across the $k$ runs to measure variability, which is shown in the error bars of the Elasticnet cross-validation plots.
+appropriate measure of model fit. We can also capture the standard deviation across the $k$ runs to measure variability, which is shown in the error bars of the Elastic Net cross-validation plots.
 
@@
 
+\pagebreak
+
 \subsection{Regularized Regression}
@@
 x_n^2$ and $||\bX|| = |x_1| + |x_2| + ... + |x_n|$. As shown in the top equation of the standard linear model, we try to find the $\beta$, that is a vector of coefficients,  
-that minimizes the squared difference between the true $\bY$ and the predicted $\hat{\bY}$ (which is $\beta\bX$). In 
+that minimizes the squared difference between the true $\bY$ and the predicted $\hat{\bY}$ (which is $\bX\beta$). In 
 regularized regression we do the same thing, but also add a second term that we look to simultaneously minimize. This second 
-term adds a penalty for increasing values of $\beta$, so the two terms must be balanced. The optimal model will find a balance 
+term adds a penalty scaled by $\lambda$ for increasing values of $\beta$, so the two terms must be balanced. The optimal model will find a balance 
 between fitting the outcome variable closely, but not having too large of coefficient values. The difference between Ridge 
@@
 
-\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[(1 - \alpha)||\beta||_2^2 + \alpha||\beta||] \tag{ElasticNet}\]
+\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[(1 - \alpha)||\beta||_2^2 + \alpha||\beta||] \tag{Elastic Net}\]
 
-The technique that is used in this analysis is a combination of the Ridge and LASSO penalties, called ElasticNet.
+The technique that is used in this analysis is a combination of the Ridge and LASSO penalties, called Elastic Net.
 As can be seen in the equation above both the square of the coefficients and the absolute value of the coefficients is included, with the contribution of each controlled by the size of alpha ($\alpha$) which takes a value between 0 and 1.
-ElasticNet, thus combines the favorable properties of Ridge and LASSO, in that it can achieve both sparse models and can handle correlated predictor variables.
+Elastic Net, thus combines the favorable properties of Ridge and LASSO, in that it can achieve both sparse models and can handle correlated predictor variables.
 Both the $\lambda$ and the $\alpha$ can be set using cross-validation (as discussed above) to appropriate values for the particular dataset.
-In each Elasticnet cross-validation plot found in the Results section, the specific $\alpha$ used is labeled at the top of the graph.
+In each Elastic Net cross-validation plot found in the Results section, the specific $\alpha$ used is labeled at the top of the graph.
 
@@
 There are a few key considerations that should be kept in mind when interpreting this analysis.
-First, is that relationships discovered in this analysis are correlational nature and cannot be assumed to be causal.
+The first is that relationships discovered in this analysis are correlational in nature and cannot be assumed to be causal.
 Just because farms with a higher coverage of invasive species have lower productivity does not necessarily mean the invasive species causes lower productivity.
@@
 regression analysis we pre-specify the model and then test which variables are found to be significant. However, when using 
-Elasticnet and Forward Selection like we have done here, we do not specify the model ahead of time, but instead let the data 
+Elastic Net and Forward Selection like we have done here, we do not specify the model ahead of time, but instead let the data 
 decide the model.
 This violates the significance test assumption and can lead to misleadingly small p-values (see Chapter 6 of \cite{hastie}).
-While there are some advanced techniques to try to adjust for this, it is recommended to view the results in this report as an exploratory analysis rather than definitive evidence.
+While there is some recent work (\cite{zhao}) suggesting this may be acceptable under certain assumptions, as well as some advanced techniques to try to adjust for this, it is recommended to view the results in this report as an exploratory analysis rather than definitive evidence.
 
-Future analysis might look to explore better fitting relationships, particularly for the outcome variables that had poor RMSE and $R^2$ values.
-The relationships modeled in this report only considered linear combinations of predictor variables to predict/explain the outcome variables.
-Modifications could include adding interaction terms (i.e., $x_1x_2$) or nonlinear terms (i.e., $x_1^2$ or binary transformations $x_1 > 10$).
-Exploring all possible modifications of this type is not computationally feasible, but with domain knowledge a subset of theorized relationships could be tested.
-Lastly, the variables used here primarily covered socioeconomic dimensions.
-The addition of physical and biotic variables may help better predict/explain the outcome variables or may change the importance of previously highlighted socioeconomic variables.
+Future analysis might look to explore better fitting relationships, particularly for the outcome variables that had poor RMSE and 
+$R^2$ values. The relationships modeled in this report only considered linear combinations of predictor variables to predict/
+explain the outcome variables. Modifications could include adding interaction terms (i.e., $x_1x_2$) or nonlinear terms (i.e., 
+$x_1^2$ or binary transformations $x_1 > 10$). Exploring all possible modifications of this type is not computationally feasible, 
+but with domain knowledge a subset of theorized relationships could be tested.
+
+For some of the outcome variables the island was found to be a significant varriable, but it may be more helpful to model each 
+of the islands separately. A potential limitation to such an approach would be the small sample sizes for some of the islands, so 
+this would be best carried out with a smaller set of potential predictor variables. Lastly, the variables used here primarily 
+covered socioeconomic dimensions. The addition of physical and biotic variables may help better predict/explain the outcome 
+variables or may change the importance of previously highlighted socioeconomic variables.
 
@@
 \bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.
+
+\bibitem{dobson} Dobson, A. J., \& Barnett, A. (2008). An introduction to generalized linear models. CRC press.
 
@@
 
+\bibitem{zhao} Zhao, S., Shojaie, A., \& Witten, D. (2017). In Defense of the Indefensible: A Very Naive Approach to High-Dimensional Inference. arXiv preprint arXiv:1705.05543.
+
 \bibitem{zou} Zou, H., \& Hastie, T. (2005). Regularization and variable selection via the elastic net. Journal of the Royal Statistical Society: Series B (Statistical Methodology), 67(2), 301-320.
@@
 \end{thebibliography}
+
+\newpage
 
@@
 \begin{tabular}{ln{1}{4}n{1}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_production.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Production model}
+\caption{\textsl{\small Full coefficient list for Production model}}
 \label{table:prod_full}
@@
 \includegraphics[width = \textwidth]{resids_production.pdf}
-\caption{Diagnostic Residual plots for Production.
+\caption{\textsl{\small Diagnostic Residual plots for Production.
 The plots on the left show the predicted values vs.
 the residuals and both demonstrate a desirable lack of pattern.
-The plots of the right examine the normality of the residuals, both staying close to the desired straight diagonal pattern.
-}
+The plots of the right examine the normality of the residuals, both staying close to the desired straight diagonal pattern.}}
 \label{figure:resids_prod}
@@
 \begin{tabular}{ln{4}{4}n{4}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_netincome.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Net Income model}
+\caption{\textsl{\small Full coefficient list for Net Income model}}
 \label{table:netincome_full}
@@
 \includegraphics[width = \textwidth]{resids_netincome.pdf}
-\caption{Diagnostic Residual plots for Net Income.
+\caption{\textsl{\small Diagnostic Residual plots for Net Income.
 The plots of the predicted values vs.
 the residuals both demonstrate a desirable lack of pattern.
-The plots examining the normality of the residuals do not follow a diagonal line as closely as would be hoped, but are not an extreme departure.}
+The plots examining the normality of the residuals do not follow a diagonal line as closely as would be hoped, but are not an extreme departure.}}
 \label{figure:resids_netincome}
@@
 \begin{tabular}{ln{1}{4}n{1}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_workers_binary.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Binary Workers model}
+\caption{\textsl{\small Full coefficient list for Binary Workers model}}
 \label{table:workers_binary_full}
@@
 \begin{tabular}{ln{1}{4}n{1}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_workers_nonzero.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Nonzero Workers model}
+\caption{\textsl{\small Full coefficient list for Nonzero Workers model}}
 \label{table:workers_nonzero_full}
@@
 \includegraphics[width = \textwidth]{resids_workers_nonzero.pdf}
-\caption{Diagnostic Residual plots for Nonzero workers. The plots of predictions versus residuals shows points bunch at the left end, but the QQ plot closely follows a diagonal line.}
+\caption{\textsl{\small Diagnostic Residual plots for Nonzero workers. The plots of predictions versus residuals shows points bunch at the left end, but the QQ plot closely follows a diagonal line.}}
 \label{figure:resids_workers_nonzero}
@@
 \begin{tabular}{ln{1}{4}n{1}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_invasive_binary.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Binary Invasive model}
+\caption{\textsl{\small Full coefficient list for Binary Invasive model}}
 \label{table:workers_invasive_full}
@@
 \begin{tabular}{ln{1}{4}n{1}{4}}\hline%
-\bfseries \underline{Variable} & \bfseries \underline{Elasticnet} & \bfseries \underline{Forward~Selection}
+\bfseries \underline{Variable} & \bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
 \csvreader[head to column names]{fullcoeflist_invasive_nonzero.csv}{}%
@@
 \end{tabular}
-\caption{Full coefficient list for Nonzero Invasive model}
+\caption{\textsl{\small Full coefficient list for Nonzero Invasive model}}
 \label{table:invasive_nonzero_full}
@@
 \includegraphics[width = \textwidth]{resids_invasive_nonzero.pdf}
-\caption{Diagnostic Residual plots for Nonzero Invasive.
-Due to only a few variables being included in these models and some of them being discrete, the predictions occur at a limited number of distinct points.}
+\caption{\textsl{\small Diagnostic Residual plots for Nonzero Invasive.
+Due to only a few variables being included in these models and some of them being discrete, the predictions occur at a limited number of distinct points.}}
 \label{figure:resids_invasive_nonzero}
@@
 \end{tabular}
-\caption{Full coefficient list for Landuse model}
+\caption{\textsl{\small Full coefficient list for Landuse model}}
 \label{table:landuse_full_a}
@@
 \end{tabular}
-\caption{Full coefficient list for Landuse model (cont.)}
+\caption{\textsl{\small Full coefficient list for Landuse model (cont.)}}
 \label{table:landuse_full_b}

### 2018-03-27 data_cleaning.R

@@ -6,6 +6,7 @@ library(readxl)
 
 load("BASES_CENSO_UPA_spss/prelim_data.RData")
 
+# Reduce dataset to subset of possible variables for consideration
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
 var_list <- vars$`Variable Name`
 for (i in 1:nrow(vars)){
@@ -14,6 +15,7 @@ for (i in 1:nrow(vars)){
 }
 reduced_data <-  data[,var_list]
 
+# Where appropriate fill NA values with 0
 binary_vars <- vars[which(vars$Binary==1),1]$`Variable Name`
 reduced_data[,binary_vars][is.na(reduced_data[,binary_vars])] <- 0
 
@@ -37,5 +39,5 @@ reduced_data$v45 <- as.numeric(reduced_data$v45)
 # Check to make sure there are no more mistaken factor variables
 reduced_data[1,sapply(reduced_data, is.factor)]
 
-
+# Output clean dataset
 save(reduced_data, file='BASES_CENSO_UPA_spss/clean_data.RData')

### 2018-03-27 data_creation.R

@@ -371,6 +371,7 @@ library(tidyverse)
 levels(data$canton) <- c(levels(data$canton), "Floreana")
 data$canton[data$UPA >= 745] = "Floreana"  
 
+# Create dummy variables
 library(dummies)
 shape_data <- read.csv("ShapeAttributes.csv")
 
@@ -389,4 +390,5 @@ shape_vars <- as.data.frame(colnames(shape_agg[,-1]))
 
 data<-merge(data, shape_agg, "UPA",  all.x=TRUE, all.y = TRUE)
 
+# Output preliminary data
 save(data, file='prelim_data.RData')

### 2018-03-27 invasive_model.R

@@ -21,11 +21,12 @@ invasive_x.df <- subset(reduced_data, select = invasive_include)
 reduced_data$invasive <- factor(reduced_data$percinv > 0)
 levels(reduced_data$invasive) = c("Zero", "Positive")
 
-invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = invasive))
+invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = invasive)) +
+  theme(text = element_text(size=14))
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = percinv)) + 
-  geom_histogram() + stat_bin(binwidth = 10)
+  geom_histogram() + stat_bin(binwidth = 10) + theme(text = element_text(size=14))
 log_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = log10(percinv))) + 
-  geom_histogram() + stat_bin(binwidth = 0.25)
+  geom_histogram() + stat_bin(binwidth = 0.25) + theme(text = element_text(size=14))
 
 pdf("Paper/images/invasive_histograms.pdf",width=12,height=8)
 grid.arrange(invasive_hist, nonzero_hist, log_hist, nrow = 1)
@@ -33,6 +34,10 @@ dev.off()
 
 invasive_binary.models <- fit.models(model.name = "invasive_binary", x.data = invasive_x.df, 
                                    y.response = reduced_data$invasive, response.family = "binomial")
+
+betas <-  invasive_binary.models$betas
+betas[abs(betas$s11)>0,'variable']
+invasive_binary.models$fwdorder
 
 
 log_invasive_nonzero <- log10(reduced_data[reduced_data$percinv > 0.1, 'percinv'])
@@ -47,6 +52,10 @@ invasive_x.nonzero <- as.data.frame(invasive_x.nonzero[,!colnames(invasive_x.non
 
 invasive_nonzero.models <- fit.models(model.name = "invasive_nonzero", x.data = invasive_x.nonzero, 
                                 y.response = log_invasive_nonzero, response.family = "gaussian")
+
+betas <-  invasive_nonzero.models$betas
+betas[abs(betas$s7)>0,'variable']
+invasive_nonzero.models$fwdorder
 
 ### Diagnostics to find linear dependencies and correlations between variables
 your.matrix <- invasive_x.nonzero

### 2018-03-27 landuse_model.R

@@ -25,6 +25,12 @@ reclass_vars <- c('ReclassAGRICOLA','ReclassCONSERVACION', 'ReclassFORESTAL', 'R
                   'ReclassPECUARIO', 'ReclassSIN APROVECHAMIENTO')
 
 
+df_melt <- melt(reduced_data[,c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush',"UPA")], id.vars = "UPA")
+pdf("Paper/images/landuse_parplot.pdf",width=12,height=8)
+ggplot(df_melt, aes(x = variable, y = value, group = UPA)) +
+  geom_path(alpha = 0.1, col = 'blue') + ylab("Percent") + xlab("Land Use")
+dev.off()
+
 perc_long1 <- melt(landuse_x.df[,survey_vars])
 pdf("Paper/images/landuse_boxplots.pdf",width=12,height=8)
 ggplot(perc_long1, aes(x=factor(variable),y=value)) + geom_boxplot() + xlab("Land Use Category") + ylab("Percent")
@@ -50,6 +56,8 @@ perc_vars <- as.matrix(reduced_data[,survey_vars])
 
 landuse.models <- fit.models(model.name = "landuse", x.data = landuse_x.df, 
                                      y.response = primary_use, response.family = "multinomial")
+
+betas[abs(betas$s1)>0,'variable']
 
 summary(landuse_x.df$ReclassPECUARIO)
 summary(landuse_x.df$`ReclassSIN APROVECHAMIENTO`)

### 2018-03-27 modeling.function.R

@@ -1,11 +1,13 @@
 fit.models <- function(model.name, x.data, y.response, response.family){
   
-  # For testing
+  source("plot.cv.big.R")
+  environment(plot.cv.big) <- asNamespace('glmnet')
   
-  #model.name <- "landuse"
-  #x.data <- landuse_x.df
-  #y.response <- primary_use
-  #response.family <- "multinomial"
+  # For testing
+  #model.name <- "production"
+  #x.data <- production_x.df
+  #y.response <- log_productivity
+  #response.family <- "gaussian"
   
   x.matrix <- model.matrix(~., x.data)[,-1]
   
@@ -31,7 +33,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     cv.25=cv.glmnet(x.matrix,y.response,family=response.family,type.measure = "class",foldid=foldid,alpha=0.25)
   }
   
-  
+  # Determine best alpha value
   models <- vector(mode="list", length=4)
   
   models[[1]] <- cv.25
@@ -51,9 +53,8 @@ fit.models <- function(model.name, x.data, y.response, response.family){
   best_alpha <- alphas[which.min(mses)]
   
   elastic.full <- glmnet(x.matrix,y.response,family=response.family,alpha=best_alpha)
-  #lambda.seq <- c(seq(1, 0.64, by=-0.05), elastic.nvars$lambda)
-  #elastic.nvars <- glmnet(x.matrix,y.response,family=response.family,alpha=best_alpha, lambda = lambda.seq)
-
+ 
+  # Find lambda value that includes 5 variables
   found_df = FALSE
   df = 5
   while (found_df == FALSE){
@@ -65,6 +66,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     }
   }
   
+  # Find RMSE for 5 variable model
   found_df = FALSE
   df = 5
   while (found_df == FALSE){
@@ -76,11 +78,15 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     }
   }
   
+  betas <- as.data.frame(as.matrix(elastic.full$beta))
+  betas <- rownames_to_column(betas, var = "variable")
   
+  # Find variable names of first 5 variables
[…]

### 2018-03-27 netincome_model.R

@@ -19,9 +19,10 @@ netincome_include <- vars[which(is.na(vars$`Net Income`)),1]$`Variable Name`
 
 
 
-full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome))
+full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + theme(text = element_text(size=14))
 reduced_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + 
-  scale_x_continuous(limits = c(-3e+04, 3e+04))
+  scale_x_continuous(limits = c(quantile(reduced_data$netincome, .025), quantile(reduced_data$netincome, .975))) + 
+  theme(text = element_text(size=14))
 pdf("Paper/images/netincome_histograms.pdf",width=12,height=8)
 grid.arrange(full_hist, reduced_hist, nrow = 1)
 dev.off()
@@ -52,3 +53,9 @@ corrs_reshape <- melt(corrs, na.rm = TRUE)
 
 netincome.models <- fit.models(model.name = "netincome", x.data = as.data.frame(netincome_x.nonzero), 
                                 y.response = netincome, response.family = "gaussian")
+
+betas <- netincome.models$betas  
+betas[abs(betas$s7)>0,'variable']
+
+netincome.models$fwdorder
+

### 2018-03-27 plot.cv.big.R

@@ -1,0 +1,23 @@
+plot.cv.big <- function (x, sign.lambda = 1, ...) 
+{
+  cvobj = x
+  xlab = "log(Lambda)"
+  if (sign.lambda < 0) 
+    xlab = paste("-", xlab, sep = "")
+  plot.args = list(x = sign.lambda * log(cvobj$lambda), y = cvobj$cvm, 
+                   ylim = range(cvobj$cvup, cvobj$cvlo), xlab = xlab, ylab = cvobj$name, 
+                   type = "n")
+  new.args = list(...)
+  if (length(new.args)) 
+    plot.args[names(new.args)] = new.args
+  do.call("plot", plot.args)
+  error.bars(sign.lambda * log(cvobj$lambda), cvobj$cvup, 
+             cvobj$cvlo, width = 0.01, col = "darkgrey")
+  points(sign.lambda * log(cvobj$lambda), cvobj$cvm, pch = 20, 
+         col = "red")
+  axis(side = 3, at = sign.lambda * log(cvobj$lambda), labels = paste(cvobj$nz), 
+       tick = FALSE, line = 0, cex.axis = 2)
+  abline(v = sign.lambda * log(cvobj$lambda.min), lty = 3)
+  abline(v = sign.lambda * log(cvobj$lambda.1se), lty = 3)
+  invisible()
+}

### 2018-03-27 production_model.R

@@ -17,15 +17,15 @@ vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
 
 production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
 production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
-
 log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])
-raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity))
-log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity)))
+
+raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) + theme(text = element_text(size=14))
+log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity))) + theme(text = element_text(size=14))
 lower_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) +
-  xlim(0, 10000) + xlab("productivity (0-10,000)")
+  xlim(0, 10000) + xlab("productivity (0-10,000)") + theme(text = element_text(size=14))
 
 pdf("Paper/images/production_histograms.pdf",width=12,height=8)
-grid.arrange(raw_prod_hist, lower_prod_hist, log_prod_hist, nrow = 1)
+grid.arrange(raw_prod_hist, lower_prod_hist, log_prod_hist, nrow = 1) 
 dev.off()
 
 production.models <- fit.models(model.name = "production", x.data = production_x.df, 

### 2018-03-27 workers_model.R

@@ -25,9 +25,12 @@ workers_x.df <- subset(reduced_data, select = workers_include)
 
 levels(reduced_data$workers_binary) = c("Zero", "Positive")
 
-all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = workers_binary)) + xlab("fulltimework")
-nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = fulltimework))
-log_nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + geom_histogram(mapping = aes(x = log10(fulltimework)))
+all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = workers_binary)) +
+  xlab("fulltimework") + theme(text = element_text(size=14))
+nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + 
+  geom_histogram(mapping = aes(x = fulltimework)) + theme(text = element_text(size=14))
+log_nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + 
+  geom_histogram(mapping = aes(x = log10(fulltimework))) + theme(text = element_text(size=14))
 
 #log_nonzero_hist2 <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(fulltimework+0.1)))
 
@@ -38,6 +41,9 @@ dev.off()
 worker_binary.models <- fit.models(model.name = "workers_binary", x.data = workers_x.df, 
                                 y.response = reduced_data$workers_binary, response.family = "binomial")
 
+betas <-  worker_binary.models$betas
+betas[abs(betas$s9)>0,'variable']
+worker_binary.models$fwdorder
 
 workers_x.nonzero.pre <- subset(reduced_data[reduced_data$fulltimework > 0,], select = workers_include)
 num_workers_log <- log10(reduced_data[reduced_data$fulltimework > 0, 'fulltimework'])
@@ -53,11 +59,14 @@ workers_x.nonzero <- as.data.frame(workers_x.nonzero[,!colnames(workers_x.nonzer
 worker_nonzero.models <- fit.models(model.name = "workers_nonzero", x.data = workers_x.nonzero, 
                              y.response = num_workers_log, response.family = "gaussian")
 
+betas <-  worker_nonzero.models$betas
[…]

### 2018-03-28 Paper/paper_draft.tex
Updated commenting

@@
   \\
-  March 27th, 2018 \\
+  March 28th, 2018 \\
 }
@@
 \setlength{\tabcolsep}{20pt}
+\setlength{\abovecaptionskip}{5pt}
 \begin{tabular}{cc}
@@
 \includegraphics[width=0.6\textwidth]{elastic_cv_production.pdf}
+\setlength{\abovecaptionskip}{0pt}
 \setlength{\belowcaptionskip}{-10pt}
@@
 \includegraphics[width=0.55\textwidth]{forward_nvars_production.pdf}
+\setlength{\abovecaptionskip}{0pt}
 \setlength{\belowcaptionskip}{-20pt}
@@
 \centering
+\setlength{\abovecaptionskip}{5pt}
 \begin{tabular}{cc}
@@
 \includegraphics[width=0.6\textwidth]{elastic_cv_netincome.pdf}
+\setlength{\abovecaptionskip}{0pt}
 \setlength{\belowcaptionskip}{-10pt}
@@
 \centering
+\setlength{\abovecaptionskip}{5pt}
 \setlength{\tabcolsep}{20pt}
@@
 \includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
-\caption{Histogram of Invasive Species.
+\caption{\textsl{\small Histogram of Invasive Species.
 To the left: the proportion of zeroes, in the middle: farms with greater than zero, and to the right: the log transformed percent of invasive species.
-The two very small values on the right plot are removed.}
+The two very small values on the right plot are removed.}}
 \label{figure:invasive_hist}
@@
 \centering
+\setlength{\abovecaptionskip}{5pt}
 \begin{tabular}{lc}
@@
 \end{table}
+
+\pagebreak
 
@@
 The outcome variable is clearly no longer normally distributed, as it is not even continuous.
-Without modification we could get predicted values below 0, above 1 or somewhere in between, none of which make much sense.
+Without modification we could get predicted values below 0, above 1 or somewhere in between, none of which make sense in this context.
 
@@
 
-\bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.
+\bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, \\sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.
 

### 2018-03-28 Paper/presentation.tex
Updated commenting

\documentclass{beamer}

\usepackage[utf8]{inputenc}
\usepackage{amsfonts, amssymb, amsmath}    
\usepackage{graphicx}
\graphicspath{ {images/} }
\usetheme{Berkeley}
\usecolortheme{seahorse}

\newcommand{\bX}{\mathbf{X}}
\newcommand{\bY}{\mathbf{Y}}
\newcommand{\bbeta}{\mathbf{\beta}}

\title[Survival of the Fittest]
{Survival of the Fittest: Variable Selection on Agricultural Data from the Gal\'apagos Islands}

\author
{Michael Bostwick}

\date
{April 19th, 2018}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Background}

\begin{frame}
\frametitle{Gal\'apagos Islands}
\end{frame}

\begin{frame}
\frametitle{Agent-Based Simulation}
\includegraphics[scale=0.3]{abm_chart}

From Miller et. al.

\end{frame}

\section{Data}

\begin{frame}
\frametitle{Data description}
\end{frame}

\begin{frame}
\frametitle{Data transformation}
\end{frame}

\section{Modeling}

\begin{frame}
\frametitle{Elastic Net}

\begin{block}{Formulation}
	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[ \underbrace{(1 - \alpha)||\beta||_2^2}_\text{Ridge} + \underbrace{\alpha||\beta||}_\text{LASSO}]\]
\end{block}

\begin{itemize}
	\item{Set $\alpha$ and $\lambda$ using cross-validation}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Forward Stepwise}
\end{frame}

\section{Results}

\begin{frame}
\frametitle{Productivity Model - Elastic Net}
\end{frame}

\begin{frame}
\frametitle{Productivity Model - Forward Stepwise}
\end{frame}

\begin{frame}
\frametitle{Productivity Model - Comparison}
\end{frame}

\begin{frame}
\frametitle{References}

\end{frame}
\end{document}

### 2018-03-28 CensusExploration2.R

@@ -1,508 +1,0 @@
-setwd("/Users/michaelbostwick/Documents/Galapagos/BASES_CENSO_UPA_spss")
-
-#Still to do: 
-  #Add categorical data from spreadsheets (such as crops)
-  #Generate physical data from DEM (This will take a while still)
-
-library(tidyverse)
-
-#Load Data (currently as SPSS files and CSV files, trying out different ones. Cleaned CSV files Use those. Omitting some )
-  library(foreign)
-  #Censo CGREG-MAG 2015
-  CGenerales <- read.csv("Cap1_CaracteristicasGenerales.csv", header = TRUE)
-  Superficie <- read.csv("Cap2_Superficie.csv", header = TRUE)
-  Riego <- read.csv("Cap3_Riego.csv", header = TRUE)
-  UsoSuelo <- read.csv("Cap3_2_UsoSuelo_Hoy.csv", header = TRUE)
-  CPermanentes <- read.csv("Cap4_CultivosPermanentes.csv", header = TRUE)
-  CTransitorios <- read.csv("Cap5_Cultivos_Transitorios.csv", header = TRUE)
-  Arboles <- read.csv("Cap6_ArbolesDispersos.csv", header = TRUE)
-  Pastos <- read.csv("Cap7_PastosCultivados.csv", header = TRUE)
-  Forestal <- read.csv("Cap8_ActividadForestal.csv", header = TRUE)
-  Vacuno <- read.csv("Cap9_GanadoVacuno.csv", header = TRUE)
-  #Porcino <- read.spss("Cap10_Ganado Porcino.sav", to.data.frame=TRUE)
-  #OtroGanado <- read.spss("Cap11_Otras Especies de Ganado.sav", to.data.frame=TRUE)
-  #Aves <- read.spss("Cap12_Aves.sav", to.data.frame=TRUE)
-  Gastos <- read.csv("Cap13_Gastos_Inversiones_Financiamiento.csv", header = TRUE)
-  Equipo <- read.csv("Cap14_Equipo_maquinaria_instalaciones.csv", header = TRUE)
-  Trabajadores <- read.csv("Cap15_Trabajadores.csv", header = TRUE)
-  AfiliacionDeshechos <- read.csv("Cap16_AsistenciaTecnica_Afiliacion_ManejoDesechos.csv", header = TRUE)
-  
-  #INEC 2015
[…]

### 2018-03-28 bootstrap_code.R

@@ -1,24 +1,0 @@
-set.seed(1)
-bootstrap_coefs <- data.frame(matrix(ncol = ncol(x.matrix)+1, nrow = 0))
-colnames(bootstrap_coefs) <- c("(Intercept)", colnames(x.matrix))
-
-for (i in 1:1000){
-  bootstrap_samples <- sample(1:nrow(x.matrix), size = nrow(x.matrix), replace = TRUE)
-  weights <- table(factor(bootstrap_samples, levels = 1:nrow(x.matrix)))
-  bs.elasticnet = glmnet(x.matrix,y.response,weights = weights,alpha=best_alpha, lambda = best.lambda)
-  
-  bs.coef <- t(data.frame(predict(bs.elasticnet,type="coefficients",s=best.lambda)[1:(ncol(x.matrix)+1),]))
-  bootstrap_coefs <- rbind(bootstrap_coefs, bs.coef)
-}
-
-row.names(bootstrap_coefs) <- c(seq(1:1000))
-top.names <- unique(c(elastic10.names, fwd.coefs10))
-top.boostrap <- bootstrap_coefs[,top.names]
-
-ggplot(data = melt(top.boostrap), aes(x=variable, y=value)) + geom_boxplot() + coord_flip()
-
-pct_zeroes <- data.frame(colSums(top.boostrap != 0)/nrow(top.boostrap))
-pct_zeroes <- rownames_to_column(pct_zeroes)
-colnames(pct_zeroes) <- c("variable", "pct_zero")
-
-ggplot(pct_zeroes) + geom_bar(stat='identity', aes(x=variable, y=pct_zero)) + coord_flip()

### 2018-03-28 invasive_model.R

@@ -10,49 +10,51 @@ library(gridExtra)
 library(leaps)
 library(bestsubset)
 
+# Load modeling function
 source("modeling.function.R")
 
+# Load clean data and subset to appropriate variables for this analysis
 load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
-
 invasive_include <- vars[which(is.na(vars$`Surface area with invasive species`)),1]$`Variable Name`
 invasive_x.df <- subset(reduced_data, select = invasive_include)
-
 reduced_data$invasive <- factor(reduced_data$percinv > 0)
 levels(reduced_data$invasive) = c("Zero", "Positive")
 
+# Plot histograms and output to PDF
 invasive_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = invasive)) +
   theme(text = element_text(size=14))
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = percinv)) + 
   geom_histogram() + stat_bin(binwidth = 10) + theme(text = element_text(size=14))
 log_hist <- ggplot(data = reduced_data[reduced_data$percinv > 0,], aes(x = log10(percinv))) + 
   geom_histogram() + stat_bin(binwidth = 0.25) + theme(text = element_text(size=14))
-
 pdf("Paper/images/invasive_histograms.pdf",width=12,height=8)
 grid.arrange(invasive_hist, nonzero_hist, log_hist, nrow = 1)
 dev.off()
 
+# Call modeling function for binary data
 invasive_binary.models <- fit.models(model.name = "invasive_binary", x.data = invasive_x.df, 
                                    y.response = reduced_data$invasive, response.family = "binomial")
 
+# To find top 5 variable order
 betas <-  invasive_binary.models$betas
 betas[abs(betas$s11)>0,'variable']
 invasive_binary.models$fwdorder
 
-
+# Prepare nonzero data and remove zero variance/linear dependent columns
 log_invasive_nonzero <- log10(reduced_data[reduced_data$percinv > 0.1, 'percinv'])
-
 invasive.matrix <- model.matrix(~., invasive_x.df[reduced_data$percinv > 0.1,])[,-1]
 col_vars <- apply(invasive.matrix, 2, var)
 zero_variance <- names(col_vars[col_vars == 0])
-
[…]

### 2018-03-28 landuse_model.R

@@ -10,55 +10,29 @@ library(gridExtra)
 library(leaps)
 library(bestsubset)
 
+# Load modeling function
 source("modeling.function.R")
 
+# Load clean data and subset to appropriate variables for this analysis
 load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
-
 landuse_include <- vars[which(is.na(vars$`Land use`)),1]$`Variable Name`
 landuse_x.df <- subset(reduced_data, select = landuse_include)
 
-perc_vars <- c(landuse_x.df$percperm, landuse_x.df$perctemp)
-
-survey_vars <- c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush',"percinv")
-reclass_vars <- c('ReclassAGRICOLA','ReclassCONSERVACION', 'ReclassFORESTAL', 'ReclassHABITACIONAL',
-                  'ReclassPECUARIO', 'ReclassSIN APROVECHAMIENTO')
-
-
+# Create parallel coordinate plot and output to PDF
 df_melt <- melt(reduced_data[,c('percperm', 'perctemp', 'percfallow', 'perctill', 'percpasture', 'percbrush',"UPA")], id.vars = "UPA")
 pdf("Paper/images/landuse_parplot.pdf",width=12,height=8)
 ggplot(df_melt, aes(x = variable, y = value, group = UPA)) +
   geom_path(alpha = 0.1, col = 'blue') + ylab("Percent") + xlab("Land Use")
 dev.off()
 
-perc_long1 <- melt(landuse_x.df[,survey_vars])
-pdf("Paper/images/landuse_boxplots.pdf",width=12,height=8)
-ggplot(perc_long1, aes(x=factor(variable),y=value)) + geom_boxplot() + xlab("Land Use Category") + ylab("Percent")
-dev.off()
-
-use_data <- reduced_data[,survey_vars]
-use_data <- use_data[order(use_data[,7],use_data[,6],use_data[,5],use_data[,4],use_data[,3],use_data[,2],use_data[,1]),][,-7]
-use_data$rownum <- 1:nrow(use_data)
-use_data_melt <- melt(use_data, id="rownum")
-pdf("Paper/images/landuse_barplot.pdf",width=12,height=8)
-g <- ggplot(use_data_melt, aes(x = rownum, y= value))
[…]

### 2018-03-28 modeling.R

@@ -1,61 +1,0 @@
-setwd("/Users/michaelbostwick/Documents/Galapagos")
-
-library(tidyverse)
-library(ggplot2)
-library(readxl)
-library(glmnet)
-library(mpath)
-
-load("BASES_CENSO_UPA_spss/clean_data.RData")
-
-vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I242")
-
-production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
-production_x <- subset(reduced_data, select = production_include)
-#production_x_num <- as.matrix(production_x[sapply(production_x, function(x) !is.factor(x))])
-production_x <- model.matrix(~., production_x)[,-1]
-
-fit <- lm(productivity ~ . , data = subset(reduced_data, select = production_include))
-
-fit = glmnet(production_x, reduced_data$productivity)
-plot(fit)  
-
-reduced_data$log_productivity <- log(reduced_data$productivity+1)
-
-foldid=sample(1:10,size=length(reduced_data$log_productivity),replace=TRUE)
-cv1=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=1)
-cv.5=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=.5)
-cv0=cv.glmnet(production_x,reduced_data$log_productivity,foldid=foldid,alpha=0)
-
-
-par(mfrow=c(2,2))
-plot(cv1,main="Alpha = 1");plot(cv.5,main="Alpha = 0.5");plot(cv0,main="Alpha = 0");plot()
-#plot(log(cv1$lambda),cv1$cvm,pch=19,col="red",xlab="log(Lambda)",ylab=cv1$name,  main="Alpha = 1")
-#points(log(cv.5$lambda),cv.5$cvm,pch=19,col="grey", main="Alpha = 0.5")
-#points(log(cv0$lambda),cv0$cvm,pch=19,col="blue", main="Alpha = 0")
-#legend("bottomleft",legend=c("alpha= 1","alpha= .5","alpha 0"),pch=19,col=c("red","grey","blue"))
-par(mfrow=c(1,1))
-
-lasso.coef=predict(cv1,type="coefficients",s=cv1$lambda.1se)[1:240,]
-lasso.coef[lasso.coef!=0]
-
-#model fulltime workers as poisson? overdispersion? negative binomial?
-hist(reduced_data$fulltimework)
[…]

### 2018-03-28 modeling.function.R

@@ -235,6 +235,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     fwd5.r2 <- best.fwd.summary$rsq[5]
     fwd.full.r2 <- best.fwd.summary$rsq[which.min(best.fwd.summary$bic)]
     
+    # Output forwad selection plot
     path <- sprintf("Paper/images/forward_nvars_%s.pdf", model.name)
     pdf(path,width=12,height=8)
     par(mar=c(5,5,5,2)+0.1)
@@ -253,6 +254,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     fwd.steps <- step(model.null,scope = list(upper=model.full), direction="forward", data=fwd.data,
                     k = log(nrow(fwd.data)), trace = 0)
     
+    # Output forwad selection plot
     path <- sprintf("Paper/images/forward_nvars_%s.pdf", model.name)
     pdf(path,width=12,height=8)
     par(mar=c(5,5,5,2)+0.1)
@@ -261,10 +263,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     points(which.min(fwd.steps$anova$AIC),min(fwd.steps$anova$AIC),col="red",cex=2,pch=20)
     dev.off()
     
-    # fwd.coefs5 <- fwd.steps$anova$Step[2:6]
-    # class(fwd.coefs5) <- class(fwd.coefs5)[-match("AsIs", class(fwd.coefs5))]
-    # fwd.coefs5 <- trimws(gsub("[^[:alnum:] ]", "", fwd.coefs5))
-    
+
     fwd.coefs5 <- names(fwd.steps$coefficients)[2:6]
     fwd.order <- fwd.coefs5
     fwd.coefs5.signs <- lapply(fwd.steps$coefficients[2:6], sign_check)
@@ -288,6 +287,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
     fwd.full.r2 <- mean(fwd.correct)
   }
   if (response.family == "gaussian"){
+    # Output residual plots
     path <- sprintf("Paper/images/resids_%s.pdf", model.name)
     pdf(path,width=12,height=8)
     par(mfrow=c(2,2))
@@ -303,6 +303,7 @@ fit.models <- function(model.name, x.data, y.response, response.family){
 
   }
   
+  # Combine top 5 variables from Elastic Net and Forward Selection into 1 table
[…]

### 2018-03-28 netincome_model.R

@@ -10,15 +10,17 @@ library(gridExtra)
 library(leaps)
 library(bestsubset)
 
+# Load modeling function
 source("modeling.function.R")
 
+# Load clean data and subset to appropriate variables for this analysis
 load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
-
 netincome_include <- vars[which(is.na(vars$`Net Income`)),1]$`Variable Name`
+netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome),], select = netincome_include)
+netincome <- reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome), 'netincome']
 
-
-
+# Plot histograms and output to PDF
 full_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + theme(text = element_text(size=14))
 reduced_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = netincome)) + 
   scale_x_continuous(limits = c(quantile(reduced_data$netincome, .025), quantile(reduced_data$netincome, .975))) + 
@@ -27,14 +29,11 @@ pdf("Paper/images/netincome_histograms.pdf",width=12,height=8)
 grid.arrange(full_hist, reduced_hist, nrow = 1)
 dev.off()
 
-netincome_x.df <- subset(reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome),], select = netincome_include)
-netincome <- reduced_data[abs(reduced_data$netincome) < sd(reduced_data$netincome), 'netincome']
 
-
+# Check for zero variance columns and remove
 netincome.matrix <- model.matrix(~., netincome_x.df)[,-1]
 col_vars <- apply(netincome.matrix, 2, var)
 zero_variance <- names(col_vars[col_vars == 0])
-
 netincome_x.nonzero <- netincome.matrix[,!colnames(netincome.matrix) %in% zero_variance]
 
 
@@ -43,19 +42,15 @@ your.matrix <- netincome_x.nonzero
 rankifremoved <- sapply(1:ncol(your.matrix), function (x) qr(your.matrix[,-x])$rank)
 which(rankifremoved == max(rankifremoved))
 
-corrs <-cor(netincome_x.nonzero)
-corrs[upper.tri(corrs)] <- NA
-diag(corrs) <- NA
-
[…]

### 2018-03-28 production_model.R

@@ -10,23 +10,27 @@ library(gridExtra)
 library(leaps)
 library(bestsubset)
 
+# Load modeling function
 source("modeling.function.R")
 
+# Load clean data and subset to appropriate variables for this analysis
 load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
-
 production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
 production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
+
+# Perform log transformation
 log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])
 
+# Plot histograms and output to PDF
 raw_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) + theme(text = element_text(size=14))
 log_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(productivity))) + theme(text = element_text(size=14))
 lower_prod_hist <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = productivity)) +
   xlim(0, 10000) + xlab("productivity (0-10,000)") + theme(text = element_text(size=14))
-
 pdf("Paper/images/production_histograms.pdf",width=12,height=8)
 grid.arrange(raw_prod_hist, lower_prod_hist, log_prod_hist, nrow = 1) 
 dev.off()
 
+# Call modeling function
 production.models <- fit.models(model.name = "production", x.data = production_x.df, 
                                 y.response = log_productivity, response.family = "gaussian")

### 2018-03-28 workers_model.R

@@ -11,62 +11,58 @@ library(leaps)
 library(bestsubset)
 library(bestglm)
 
+# Load modeling function
 source("modeling.function.R")
 
+# Load clean data and subset to appropriate variables for this analysis
 load("BASES_CENSO_UPA_spss/clean_data.RData")
 vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
-
 workers_include <- vars[which(is.na(vars$`Number workers supported`)),1]$`Variable Name`
+workers_x.df <- subset(reduced_data, select = workers_include)
 
 workers_nonzero <- reduced_data[reduced_data$fulltimework > 0, 'fulltimework']
 reduced_data$workers_binary <- factor(reduced_data$fulltimework > 0)
-
-workers_x.df <- subset(reduced_data, select = workers_include)
-
 levels(reduced_data$workers_binary) = c("Zero", "Positive")
 
+# Plot histograms and output to PDF
 all_workers_hist <- ggplot(data = reduced_data) + geom_bar(mapping = aes(x = workers_binary)) +
   xlab("fulltimework") + theme(text = element_text(size=14))
 nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + 
   geom_histogram(mapping = aes(x = fulltimework)) + theme(text = element_text(size=14))
 log_nonzero_hist <- ggplot(data = reduced_data[reduced_data$fulltimework > 0,]) + 
   geom_histogram(mapping = aes(x = log10(fulltimework))) + theme(text = element_text(size=14))
-
-#log_nonzero_hist2 <- ggplot(data = reduced_data) + geom_histogram(mapping = aes(x = log10(fulltimework+0.1)))
-
 pdf("Paper/images/worker_histograms.pdf",width=12,height=8)
 grid.arrange(all_workers_hist, nonzero_hist, log_nonzero_hist, nrow = 1)
 dev.off()
 
+# Call modeling function on binary data
 worker_binary.models <- fit.models(model.name = "workers_binary", x.data = workers_x.df, 
                                 y.response = reduced_data$workers_binary, response.family = "binomial")
 
+# To find top 5 variable order
 betas <-  worker_binary.models$betas
-betas[abs(betas$s9)>0,'variable']
+betas[abs(betas$s0)>0,'variable']
 worker_binary.models$fwdorder
 
[…]

### 2018-04-20 Paper/presentation.tex
Presentation updates

@@
 \graphicspath{ {images/} }
-\usetheme{Berkeley}
-\usecolortheme{seahorse}
+\usetheme{Metropolis}
 
@@
 \frametitle{Gal\'apagos Islands}
+\begin{figure}
+\begin{tabular}{cc}
+ \includegraphics[scale=0.09]{galapagosmap} \footnote{https://canconf.com/map-of-south-america-including-galapagos-islands/} &   \includegraphics[scale=0.2]{turtle} \footnote{https://www.flickr.com/photos/pancholp/albums} \\
+ \includegraphics[scale=0.2]{cow1} &   \includegraphics[scale=0.2]{invasive}
+\end{tabular}
+\end{figure}
+
 \end{frame}
@@
 \frametitle{Agent-Based Simulation}
-\includegraphics[scale=0.3]{abm_chart}
+\centering
+\includegraphics[scale=0.3]{abm_chart} \footnote{Miller, B.
+W., Breckheimer, I., McCleary, A.
+L., Guzm\'an-Ramirez, L., Caplow, S.
+C., Jones-Smith, J.
+C., \& Walsh, S.
+J.
+(2010).
+Using stylized agent-based models for population environment research: a case study from the Gal\'apagos Islands.
+Population and environment, 31(6), 401-426.}
 
-From Miller et. al.
+\end{frame}
+
+\begin{frame}
+\frametitle{Census of Farms}
+\centering
+\includegraphics[scale=0.5]{survey}
 
@@
 \frametitle{Data description}
+\centering
+\includegraphics[scale=0.4]{755_farms}
+
+Model 5 different response variables with 200+ possible predictors
 \end{frame}
@@
 \frametitle{Data transformation}
+\centering
+%\[ \bY = \bX\beta + \epsilon, \epsilon \sim N(0,\sigma^2)  \]
+
+\includegraphics[scale=0.35]{production_histograms}
 \end{frame}
@@
 \begin{block}{Formulation}
-	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[ \underbrace{(1 - \alpha)||\beta||_2^2}_\text{Ridge} + \underbrace{\alpha||\beta||}_\text{LASSO}]\]
+	\[\min_{\beta}  \|\bY - \bX\beta\|_{2}^{2} + \lambda[ \underbrace{(1 - \alpha)||\beta||_2^2}_\text{Ridge} + \underbrace{\alpha||\beta||}_\text{Lasso}]\]
 \end{block}
 
-\begin{itemize}
-	\item{Set $\alpha$ and $\lambda$ using cross-validation}
-\end{itemize}
+%\begin{itemize}
+%	\item{Can set $\alpha$ and $\lambda$ using cross-validation}
+%	\item Compromise between Ridge and Lasso can help handle correlated variables
+%\end{itemize}
+
+\centering
+\includegraphics[width=8cm]{elastic3d} \footnote{Hastie, T., Tibshirani, R., \& Wainwright, M. (2015). Statistical learning with sparsity: the lasso and 
+generalizations. CRC press.}
 
@@
 \begin{frame}
+\frametitle{Elastic Net Coefficients}
+\centering
+
+\includegraphics[scale=0.26]{elasticplot1}
+Alpha = 1
+
+\includegraphics[scale=0.26]{elasticplot2}
+Alpha = 0.1
+\end{frame}
+
+\begin{frame}
 \frametitle{Forward Stepwise}
+
+\begin{enumerate}
+	\item Start with null model
+	\item Fit $p$ simple linear regression models and pick one with lowest residual sum of squares (RSS)
+	\item Search through remaining $p - 1$ variables and add one that best improves RSS
+	\item Repeat Step 3 until all variables have been added to the model
+	\item Choose optimal model using information criterion, Bayesian Information Criterion (BIC) in this case
+\end{enumerate}
+
+\[ BIC =  \frac{1}{n}(RSS + log(n)d\hat{\sigma}^2)\]
+
 \end{frame}
@@
 \begin{frame}
+\frametitle{Response Variables}
+\begin{itemize}
+	\item{Productivity}
+	\item{Net Income}
+	\item{Number of Workers Supported}
+	\item{Invasive Species}
+	\item{Land Use Choices}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
 \frametitle{Productivity Model - Elastic Net}
+
+\begin{columns}
+
+\column{0.5\textwidth}
+First 5 variables to enter model
+\begin{enumerate}
+	\item Percent pasture land (-)
+	\item No pasture land (+) 
+	\item Percent of feed from grass (-) 
+	\item Percent permanent crops (+) 
+	\item Surface area of pastures (-)
+\end{enumerate}
+
+\column{0.5\textwidth}
+\includegraphics[scale=0.2]{elastic_cv_production}
+
+\end{columns}
 \end{frame}
@@
 \frametitle{Productivity Model - Forward Stepwise}
+\begin{columns}
+
+\column{0.5\textwidth}
+First 5 variables to enter model
+\begin{enumerate}
+	\item San Cristobal Island (+)
+	\item Percent invasive species (-)
+	\item 2 year growth in permanent crops (+)
+	\item Percent brush (-)
+	\item Growing papaya (+)
+\end{enumerate}
+
+\column{0.5\textwidth}
+\includegraphics[scale=0.2]{forward_nvars_production}
+
+\end{columns}
+
 \end{frame}
@@
 \frametitle{Productivity Model - Comparison}
+\centering
+\includegraphics[scale=0.4]{corplot}
 \end{frame}
@@
 \begin{frame}
-\frametitle{References}
+\frametitle{Productivity Model - Comparison}
+\centering
+\includegraphics[scale=0.25]{plsplot1}
+
+\includegraphics[scale=0.25]{plsplot2}
 
 \end{frame}
+
+\begin{frame}
+\centering
+{\huge \textbf{Questions?}}
+
+\end{frame}
+
 \end{document}

### 2018-04-20 for_presentation.R

@@ -1,0 +1,103 @@
+setwd("/Users/michaelbostwick/Documents/Galapagos")
+
+library(readxl)
+
+# Load clean data and subset to appropriate variables for this analysis
+load("BASES_CENSO_UPA_spss/clean_data.RData")
+vars <-  read_excel("Variables.xlsx", sheet = "vars", range = "C2:I241")
+production_include <- vars[which(is.na(vars$`UPA Production`)),1]$`Variable Name`
+production_x.df <- subset(reduced_data[reduced_data$productivity > 0,], select = production_include)
+
+# Perform log transformation
+log_productivity <- log10(reduced_data[reduced_data$productivity > 0, 'productivity'])
+
+elastic5 <-  production_x.df[,c("pc4", "percpasture", "percperm","v30_a","pc6")]
+elastic5.mat <- model.matrix(~.,data=elastic5)[,-1]
+cor(elastic5.mat[,-1])  
+
+forward5 <- production_x.df[,c("canton", "percinv", "percperm2","percbrush","CPermanentes_PAPAYA")]
+forward5.mat <- model.matrix(~.,data=forward5)[,-1]
+cor(forward5.mat[,-c(2,3)])  
+
+
+var.names <- c("PastureNo","PercPasture","PercPerm","GrassFeed","PastureSurface",
+               "SanCristobal","PercInv","PercPerm2","PercBrush","Papaya")
+keep_vars1 <- c("pc4None","percperm","pc6", "cantonSan Cristobal",  
+                "percinv","percperm2","percbrush","CPermanentes_PAPAYA")
+var.names1 <- c("PastureNo","PercPerm","PastureSurface",
+                "SanCristobal","PercInv","PercPerm2","PercBrush","Papaya")
+keep_vars2 <- c("percpasture", "v30_a")
+var.names2 <- c("PercPasture","GrassFeed")
+
+all5 <- production_x.df[,c("pc4", "percpasture", "percperm","v30_a","pc6",
+                           "canton", "percinv", "percperm2","percbrush","CPermanentes_PAPAYA")]
+all5.mat <- model.matrix(~.,data=all5)[,-1]
+remove_vars <- c("pc4Solo","cantonSanta Cruz", "cantonFloreana")
+keep_vars <- c("pc4None","percpasture", "percperm", "v30_a","pc6", "cantonSan Cristobal",  
[…]
